Run Alexes notebook for 9 MO full surface AB panel to create dataframe of expression values
a

```{r,include=FALSE}
# setup the enviroment
# multiple packages may need to be installed

require("flowCore") #Used for reading the data
require("ggplot2")
require("ggridges") #visualization
require("stringr") #set of functions to manipulate strings type in order to have nice titles
require("rlist") #set of functions that allows to easily manipulate lists
require("reshape2") #visualization
require("flowStats") #Alignment functions
require("scales") #scale colour intensity for visualization
require("dplyr")

#libraries
library("flowCore")

# installations
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("flowCore")

if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("flowStats")



```


Functions to be called later - run code chunk to have the function 'callable' in your working enviromnent

```{r}
# functions Alex wrote
# will need to make these into a separate function with documentation

# function plotdensity_flow set

plotdensity_flowset <- function(flowset){ ggplot(melt(lapply(as.list(flowset@frames),function(x){x=as.data.frame(x@exprs)})), aes(x=value,y=L1,fill=L1)) + geom_density_ridges(alpha=.4,verbose=FALSE) +facet_wrap(~variable)+theme_light()} 

#defines a function for visualizing flowset with densityplots

# function renmame markers
# fcs files will have the marker names input during acquistion using flowjo

rename_markers<-function(flowset){#Defines a function to use marker names 
  copy_flowset=flowset[seq(along=flowset)]
  for (i in 1:length(copy_flowset)){
    marker.names=copy_flowset[[i]]@parameters@data$desc
    marker.names=lapply(marker.names,function(x){str_replace_all(x,"-","_")})
    colnames(copy_flowset[[i]]@exprs)<-unlist(lapply(marker.names, function(x){sapply(str_split(x,"_"),head,1)})) 
  }
  return(copy_flowset)
}



```



```{r}

input_path2 <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/FlowDataFiles/9MBO"
input_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/GatingPlanExperiment/May10_gating/"

output_path2 <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outs2/"
output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/GatingPlanExperiment/Output/"

write_fcs_files <- TRUE #Set to true to write fcs files at each step (subsampled, transformed, aligned and scaled) - recommanded 

#Create output folder if it's not already created
if(dir.exists(output_path)==FALSE){ #check
  dir.create(output_path) #create directory
}


```


#Read in the data - This chunck reads in the data and selects the area values that are used for analysis. It creates a flowSet object (containing multiple flowFrames, corresponding to fcs files)

```{r}


# all the files in the input folder will be added to this data object flowset
flowset.new <- read.flowSet(path=input_path,transformation = FALSE ,emptyValue = FALSE,truncate_max_range = FALSE, package="flowCore") #Create a flowset object
flowset <- flowset.new

# flowset1 is the new data
flowset1 <- fsApply(flowset,
                   function(x){x[ ,grepl("^[{'FJComp'}|{'FCS'}|{'SSC'}].*A$",colnames(flowset))]}) #Apply a function to flowSet (fsApply) that selects only the columns corresponding to areas values using "regular expression".

# check the 9000MBO data
flowset <- read.flowSet(path=input_path2,transformation = FALSE ,emptyValue = FALSE,truncate_max_range = FALSE, package="flowCore") #Create a flowset object

flowset2 <- fsApply(flowset,
                   function(x){x[ ,grepl("^[{'FJComp'}|{'FCS'}|{'SSC'}].*A$",colnames(flowset))]}) #Apply a function to flowSet (fsApply) that selects only the columns corresponding to areas values using "regular expression".


### @Shuming I can see in the old fsc files (9000 cells from 9 MBO we used before) the FSC information is present in the two data slots.  In the new data it is onlyin one spot

flowset@frames$`2020-03-06- export_bioinfo_3450c_live cells.fcs`@parameters@data$name
# here under 'name' we see the FSC-A in the old data


```

We see the FSC-A etc in the desc slot in the old data

```{r}
flowset@frames$`2020-03-06- export_bioinfo_3450c_live cells.fcs`@parameters@data$desc
```

In the new data we see the FSC labels only under 'name' not under 'desc' (which I think is description).

```{r}
flowset.new@frames$export_MBO_AIW_2_AUG_2021_PBS_livecells.fcs@parameters@data$name
```
```{r}
flowset.new@frames$export_MBO_AIW_2_AUG_2021_PBS_livecells.fcs@parameters@data$desc
```



```{r}
# see number of cells per sample
print(fsApply(flowset1,function(x){dim(x@exprs)})[,1])


sampleNames(flowset1) #Prints the name of each flowframe inside the flowset. User can modify it in the following chunck
#write.flowSet(flowset,outdir=paste0(output_path,"input_flowset"))#Writes the input flowset
#flowset=read.flowSet(path=paste0(output_path,"input_flowset"),phenoData = "annotation.txt")#Read the written flowset
```


#This chunck can be used for renaming the fcs files inside the flowset
```{r}
# create a vector with the new file names and assign them
# this must be a flowCore function
# note there are two ages - I'll look up the date and calulate ages later for old I'll write A and young B 0306 and 0317 are march 6 and march 17 the experiment dates of dissociation and acquistion

#sampleNames(flowset) <- c("3450_0306","AIW002_0306","AJG001C_0306","3450_0317A","AIW002_0317A","AJG001C_0317A","3450_0317B","AIW002_0317B","AJG001C_0317B") #directly modify the names inside the flowset

sampleNames(flowset1) <- c("S2-B","S2-A","S1-B","S1-A")
sampleNames(flowset1)
# now the sample or file names are changed under the 'frame' data slot


# this appears to be again reading in the data an getting the expression values
# I'll make this a function for the package



```


#Subsets - Allows to subset each fcs with a desired number of cells selected randomly (set the seed), go directly to transformation step for not subsetting

To determine if I want to subset cells or how many to subset I need to know the number of cells in each sample. This can be seen in the table above

For this analysis I'll take 9000 events for each sample except the final sample I'll take all 1578


Set an arbitrary number of cells for subsetting samples (if a fcs files has less cells it will be completely selected)
```{r,include=FALSE}
desired_size <- 12500
# I've run this chunk setting the cell number to 9000
```
Set the number to the smallest fcs files (containing the smallest amount of cells)
```{r,include=FALSE}
desired_size <- min(fsApply(flowset,function(x){nrow(x@exprs)}))
# not run
```
Subsets the data
```{r,echo=FALSE}
set.seed(42) #Set a seed for reproducibility 
sf <- sampleFilter(filterId = "SizeFilter", size =desired_size) #Creates a "filter" object to subset
flowset.sub <- fsApply(flowset1,function(x){Subset(x,sf,truncate_max_range = FALSE)}) #apply the filter on every flowframe

```


#Writing csv files with informations

```{r}
# watchout -I'm renaming all the columns to match the first dataframe in the list to avoid the R bind error
# they were all the same by visual instruction but it's possible I'm renaming with the wrong AB which will be a huge problem

# saving function
flowset_to_csv=function(flowset){  
  list_of_flowframes=fsApply(rename_markers(flowset),function(x){as.data.frame(x@exprs)})#Makes a list of dataframes
  list_names=names(list_of_flowframes)#extract names for not calling names() function at each loop
  for (index in seq_along(list_of_flowframes)){ #Iterates along the index for adding sample names
    list_of_flowframes[[index]]=list_of_flowframes[[index]] %>% #Using tidyverse package for adding features (name of batch)
      mutate(Sample=list_names[index])#Using tidyverse package for adding features (name of batch)
  #rename the columns to fix the rbind error 
      colnames(list_of_flowframes[[index]]) = colnames(list_of_flowframes[[1]])  
  }
  # this is wehre the error occurs but all the df have the same column names???
  ds=list.rbind(list_of_flowframes)#binds every fcs file in a single dataframe
  ds$cell=as.factor(unlist(lapply(as.list(c(1:length(flowset))),function(x){c(1:nrow(flowset[[x]]@exprs))})))#add cell IDs - cell count per sample 
  write.csv(ds,file=paste0(output_path,deparse(substitute(flowset)),".csv"))#save the R data for further usage
}

# there is an error in the appove function or the below application
# this should make a csv with the antibody names - I haven't renamed the antibodies and I don't see that code chunk 



flowset_to_csv(flowset1)#new files - error
flowset_to_csv(flowset2)#old files 





```


Try debugging - why can't we save the new files?  Is it because the descriptions are wrong?

```{r}
# saving function
flowset = flowset1

  list_of_flowframes=fsApply(rename_markers(flowset),function(x){as.data.frame(x@exprs)})#Makes a list of dataframes
  # fine
# check what it is
  list_of_flowframes$`S2-B`
  
  
  list_names=names(list_of_flowframes)#extract names for not calling names() function at each loop
  # fine

  for (index in seq_along(list_of_flowframes)){ #Iterates along the index for adding sample names
    list_of_flowframes[[index]]$Sample = list_names[index] # add the sample name column
     
     # colnames(list_of_flowframes[[index]]) = colnames(list_of_flowframes[[1]])  
  }
  # error 
  
  
  # this is wehre the error occurs but all the df have the same column names???
  ds=list.rbind(list_of_flowframes)#binds every fcs file in a single dataframe
  ds$cell=as.factor(unlist(lapply(as.list(c(1:length(flowset))),function(x){c(1:nrow(flowset[[x]]@exprs))})))#add cell IDs - cell count per sample 
  write.csv(ds,file=paste0(output_path,deparse(substitute(flowset)),".csv"))#save the R data for further usage


  
  # whats inside the flowframe list
  
  list_of_flowframes
  
  colnames(list_of_flowframes$export_MBO_AIW_2_AUG_2021_PBS_livecells.fcs)
  
  #maybe we can add a conditions - if column name is NA rename 'NA.1", "NA.2"... and so on
  # first check for missing names for duplicates output
  # then rename in all the dataframes
  
  
colnames(list_of_flowframes$export_MBO_AIW_2_AUG_2021_PBS_livecells.fcs)[1] <- "Rename1"
colnames(list_of_flowframes$export_MBO_AIW_2_AUG_2021_PBS_livecells.fcs)[2] <- "Rename2"
  
colnames(list_of_flowframes$export_MBO_AIW_2_AUG_2021_PBS_livecells.fcs)

# yes we can rename

# add use rename in all

for(df in list_of_flowframes){
   colnames(df[1])="rename1"
   colnames(df[1])="rename2"
 }



  
```

```{r}
# break apart the for loop

index = 1
  
    list_of_flowframes[[index]]$Sample = list_names[index]
  # this fixes the adding the sample column problem.   
    
    #Using tidyverse package for adding features (name of sample/file name)
  #rename the columns to fix the rbind error 
      colnames(list_of_flowframes[[index]]) = colnames(list_of_flowframes[[1]])  
  
  # error 


```


```{r}
# saveing function updated
# saving function
flowset_to_csv=function(flowset){  
  list_of_flowframes=fsApply(rename_markers(flowset),function(x){as.data.frame(x@exprs)})#Makes a list of dataframes
  list_names=names(list_of_flowframes)#extract names for not calling names() function at each loop
  for (index in seq_along(list_of_flowframes)){ #Iterates along the index for adding sample names
    list_of_flowframes[[index]]$Sample = list_names[index]
    colnames(list_of_flowframes[[index]]) = colnames(list_of_flowframes[[1]]) 
  }
  ds=list.rbind(list_of_flowframes)#binds every fcs file in a single dataframe
  ds$cell=as.factor(unlist(lapply(as.list(c(1:length(flowset))),function(x){c(1:nrow(flowset[[x]]@exprs))})))#add cell IDs - cell count per sample 
  write.csv(ds,file=paste0(output_path,deparse(substitute(flowset)),".csv"))#save the R data for further usage
}

flowset_to_csv(flowset1)

# no error
flowset_to_csv(flowset2)

```







With the old files

```{r}

# saving function
flowset = flowset2

  list_of_flowframes2=fsApply(rename_markers(flowset),function(x){as.data.frame(x@exprs)})#Makes a list of dataframes
list_of_flowframes2$`2020-03-06- export_bioinfo_3450c_live cells.fcs`




  
```



