---
title: "R Notebook Create Reference matrix"
output: html_notebook
---

Here I will read in each of the datasets
1. 2D Flow
2. BenBarres bulk RNA from human isolated cells
3. scRNAseq from in house MBO 47 days
4. scRNAseq public data from human brain tissues - multiple datasets to be merged
5. scRNAseq from public data from cerebral organoids

Each dataset will normalized from 0-1

```{r}
# load libraries
library(dplyr)
library(ggplot2)

```



```{r}
# I'll start with the scRNAseq from brain tissue because there are multiple datasets. From human cell landscape
# values manually copied by cluster
# all neuron averaged, all astrocytes averaged, all radial glia averaged. 

fb3 <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/fetalbrain3.csv")
fb5 <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/fetalbrain5.csv")
lake <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/Lake.csv")
lamanno <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/lamanno.csv")
zhang <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/zhang.csv")

# column names each of the 5 HCL datasheets
colnames(fb3)
colnames(fb5)
colnames(lake)
colnames(lamanno)
colnames(zhang)

#cell_types

```

Transform values to be between 0 and 1 

```{r}

# select only numeric values  
df <- fb3
df.value <- df[0:13,3:13]


zero.one <- apply(df.value, MARGIN = 2, FUN = function(X) (X - min(X))/diff(range(X)))

# check that the function works 



```








```{r}

print(colnames(fb3))
print(colnames(fb5))
# column names were not the same
fb5 <- fb5 %>% rename(Epithelial = Edpendymal.Epithealial)

# rbind worked to join two df but it doesn't add anything automatically to indicate where the data came from these are just turning into replicates 
# add source column
fb5$Source <- "fb5"
head(fb5)


fb3 <- fb3 %>% rename(Epithelial = Edpendymal)
# add source column for fb3
fb3$Source <- "fb3"

```
Z-score each matrix get a mean of 0 and std of 1 

```{r}

sdat <- scale(fb3)


rm.cat <- fb3 %>% select(-"Annotation",-"Marker",-"Source")
scale.data <- scale(rm.cat)

rowMeans(scale.data, na.rm = TRUE)
colMeans(scale.data, na.rm = TRUE)

tryrow <- apply(rm.cat,1,scale)
rowMeans(tryrow, na.rm = TRUE)

df <-rm.cat

# margin is the dimension - 1 is rows and 2 is columns 

# example of apply
zscore <- apply(df, 1, max, na.rm = TRUE)

zscore  
# example of a function

# z = y- mean(y) /std 

# z = x - mean/ sd

# this might work more easily if I transpose the table so I z score the columns
tdf <- t(df)


tdf$z1 <- (tdf[1] - mean(tdf[1], na.rm = TRUE))/ sd(tdf[1], na.rm = TRUE)


zscorefun <- function(df) {
  new.df <- 
  return(new.df)
}

df.scale2 <- (apply(rm.cat, 1, x - mean/sd))

```



```{r}

twodf <- rbind(fb3,fb5)
head(twodf)

# perfect this works
# now I realize I need to z score first


```

