---
title: "R Notebook Create Reference matrix"
output: html_notebook
---

Here I will read in each of the datasets
1. 2D Flow
2. BenBarres bulk RNA from human isolated cells
3. scRNAseq from in house MBO 47 days
4. scRNAseq public data from human brain tissues - multiple datasets to be merged
5. scRNAseq from public data from cerebral organoids

Each dataset will normalized from 0-1

```{r}
# load libraries
library(dplyr)
library(ggplot2)

```



```{r}
# I'll start with the scRNAseq from brain tissue because there are multiple datasets. From human cell landscape
# values manually copied by cluster
# all neuron averaged, all astrocytes averaged, all radial glia averaged. 

fb3 <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/fetalbrain3.csv")
fb5 <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/fetalbrain5.csv")
lake <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/Lake.csv")
lamanno <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/lamanno.csv")
zhang <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/zhang.csv")

# column names each of the 5 HCL datasheets
colnames(fb3)
colnames(fb5)
colnames(lake)
colnames(lamanno)
colnames(zhang)

#cell_types

```

Get the markers names - Need to be sure that are all in the same order

```{r}

# check antibodies
head(fb3)
head(fb5)
head(lake)
head(lamanno)
head(zhang)

# everything looks correct

```

Transform each matrix to be between 0 and 1 across rows for the antibodies

```{r}
# put the df into a list
df.list <- list(fb3,fb5,lake, lamanno, zhang)

# normalize each df

markers <- fb3$Marker
#print(markers)

col.names.markers <- c("AQP4","CD24","CD44","CD184","CD15", "HepaCam", "CD29","CD56","O4","CD140a", "CD133", "Glast",   "CD71")   

norm.list <- lapply(df.list, function(x) {
  #source.name <- df.names[df]
  # get marker names
  #markers <- as.vector(x['Marker'])
  # make a df with no character to normalize
  df.value <- x[0:13,3:13]
  # create normalized matrix
  zero.one <- apply(df.value, MARGIN = 1, FUN = function(X) (X - min(X, na.rm=TRUE))/diff(range(X, na.rm=TRUE)))
  colnames(zero.one) <-col.names.markers
  x <- zero.one

  })

#I now have a list of normized df 
# I need to extract each dataframe, I'm not sure how to unlist and still get the name of the dataframe
#I now have a list of normized df 
fb3.norm <- norm.list[[1]]
fb5.norm <- norm.list[[2]]
lake.norm <- norm.list[[3]]
lamanno.norm <- norm.list[[4]]
zhang.norm <- norm.list[[5]]

fb3.norm[1,2] # row, column  The row names are outside of the matrix and won't be part of the index
fb5.norm[1,2]
lake.norm[1,2]
lamanno.norm[1,2]
zhang.norm[1,2]

# see the datatabels
head(fb3.norm)
head(fb5.norm)
head(lake.norm)
head(lamanno.norm)
head(zhang.norm)


# the tabels are now transposed
# cell types are rows and Antibodies are columns
class(fb3.norm)
# the outputs are not a [1] "matrix" "array" 

```


Now we want to take the means - 

```{r}

# this was to test the means
# we don't have the means of the matrix yet

aqp4 <- c(fb3.norm[1,1],fb5.norm[1,1],lake.norm[1,1],lamanno.norm[1,1],zhang.norm[1,1])

mean.aqp4 = mean(na.omit(aqp4))
mean.aqp4
markers <- as.vector(lake.norm['Marker'])



```

Next I will get the mean values for each celltype/marker from across the 5 scRNAseq dataset (public)

```{r}

# we need to make an array
#  It takes rows(x), columns(y), and a number of matrices of dimensions nrow * ncol(z).
dim(fb3.norm) 
# [1] 11 13


arr <- array(as.numeric(unlist(norm.list)), dim = c(11, 13, 5))


all.matrix <- abind(arr, along=3)
df.mean <- apply(all.matrix, c(1,2), mean,na.rm=TRUE)
head(df.mean)

# I have some values now - I need to check them
# yes this has worked

# add the column names back in - these are now the antibody names
colnames(df.mean) <- col.names.markers  
head(df.mean)

# add the cell type names back in 
cell.types <- rownames(fb3.norm)
rownames(df.mean) <- cell.types
head(df.mean)

sc.pub <- df.mean
write.csv(df.mean, "scRNA.pub.mean.exp_20210517.csv")




```


Now I need to prepare each of the other expression matrices in the same way


```{r}
bb <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/Inputs_for_reference_combined/BenBarres_marker_exp_input.csv")

ussc <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/Inputs_for_reference_combined/UCSC_organoid_average_by_cluster.csv")

mbo.47 <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/Inputs_for_reference_combined/MBO47days.csv")

mbo.160 <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/Inputs_for_reference_combined/MBO_160days.csv")

#look at the data make sure the row and column orders are the same

head(bb)
head(ussc)
head(mbo.47)
head(mbo.160)




```


The only other dataset is the flow cytometery data - this has a lot of negative values 


```{r}
# checkout the flow cytometery data - might need to transform

fc <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/Inputs_for_reference_combined/Flow_2D.csv")
# I'm able to use the negative values it is fine. 

# test if the NA in CD184 is the problem
#fc <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/Inputs_for_reference_combined/Flow_2D_changeCD184.csv")

# yes this is the problem but this means I might be normalizing by column and not by row. 
head(fc)

#in the min max function to get the values between zero and 1 all values are NaN but maybe it's because of the NAN 


```

Now I'll normalize the values between zero and 1 for all the rest of the df 

```{r}
# put the df into a list
df.list <- list(bb,ussc,mbo.47,mbo.160,fc)

# normalize each df

markers <- mbo.160$Marker
#print(markers)

col.names.markers <- c("AQP4","CD24","CD44","CD184","CD15", "HepaCam", "CD29","CD56","O4","CD140a", "CD133", "Glast",   "CD71")   

norm.list <- lapply(df.list, function(x) {
  #source.name <- df.names[df]
  # get marker names
  #markers <- as.vector(x['Marker'])
  # make a df with no character to normalize
  df.value <- x[0:13,3:13]
  # create normalized matrix
  zero.one <- apply(df.value, MARGIN = 1, FUN = function(X) (X - min(X, na.rm=TRUE))/diff(range(X, na.rm=TRUE)))
  colnames(zero.one) <-col.names.markers
  x <- zero.one

  })

#I now have a list of normized df 
# I need to extract each dataframe, I'm not sure how to unlist and still get the name of the dataframe
bb.norm <- norm.list[[1]]
ucsc.norm <- norm.list[[2]]
mbo.47.norm <- norm.list[[3]]
mbo.160.norm <- norm.list[[4]]
fc.norm <- norm.list[[5]]

# check if the dataframes are correct
bb.norm[1,2] # row, column
ucsc.norm[1,2]
mbo.47.norm[1,2]
mbo.160.norm[1,2]
fc.norm[1,2]

head(bb.norm)
head(ucsc.norm)
head(mbo.47.norm)
head(mbo.160.norm)
head(fc.norm)
# the tabels are now transposed
# cell types are rows and Antibodies are columns


```
Now I have the following inputs:
sc.pub this is the mean of 5 fetal brain, fetal midbrain and adult brain from HCL
ucsc lots of cerebral organoids
mbo.47 My scRNAseq data from 47 days MBO
mbo.160 My scRNAseq from 160 days MBO
bb  Brain total RNAseq from sorted brain cells Ben Barres
fc Flow cytometery data with the antibody pannel but no CD184


```{r}
# I'll list these all together these are now matrix array 
# I'll take the mean of the 6 and see if things how it looks
df.list2 <- list(sc.pub, bb.norm,ucsc.norm,mbo.47.norm,mbo.160.norm,fc.norm)

arr2 <- array(as.numeric(unlist(df.list2)), dim = c(11, 13, 6))

all.matrix <- abind(arr2, along=3)
rf.mean <- apply(all.matrix, c(1,2), mean,na.rm=TRUE)
head(rf.mean)

# I have some values now - I need to check them
# make a vector of all the AQP4 at stem cell values
aqp4 <- c(sc.pub[1,1],bb.norm[1,1],ucsc.norm[1,1],mbo.47.norm[1,1],mbo.160.norm[1,1],fc.norm[1,1])
# take the mean
mean.aqp4 = mean(na.omit(aqp4))
mean.aqp4

# add the column names back in - these are now the antibody names
colnames(rf.mean) <- col.names.markers  
# add the cell type names back in 
rownames(rf.mean) <- cell.types
head(rf.mean)

# save the reference matrix
write.csv(rf.mean, "ReferenceMatrix_meanof6_20210517_B.csv")

```


Try to plot the matrix

```{r}
# keep the original matrix 
# then plot each matrix with this script
A <- rf.mean

longData<-melt(A)
longData<-longData[longData$value!=0,]
ggplot(longData, aes(x = Var2, y = Var1)) + 
  geom_raster(aes(fill=value)) + 
  scale_fill_gradient(low="grey90", high="red") +
  labs(x="Antibodies", y="Cell Types", title="Matrix") +
  theme_bw() + theme(axis.text.x=element_text(size=9, angle=90, vjust=0.3),
                     axis.text.y=element_text(size=9),
                     plot.title=element_text(size=11))

png("ReferenceMatrix.png")
ggplot(longData, aes(x = Var2, y = Var1)) + 
  geom_raster(aes(fill=value)) + 
  scale_fill_gradient(low="grey90", high="red") +
  labs(x="Antibodies", y="Cell Types", title="Matrix") +
  theme_bw() + theme(axis.text.x=element_text(size=9, angle=90, vjust=0.3),
                     axis.text.y=element_text(size=9),
                     plot.title=element_text(size=11))
dev.off()


```

See some of the other matrix of expression

```{r}
# keep the original matrix 
# then plot each matrix with this script
# Ben Barres total seq
# A <- bb.norm
# HCL mean data
#A <- sc.pub
A<- fc.norm
# MBO inhouse scRNA seq
#A <- mbo.47.norm
#A <- mbo.160.norm
longData<-melt(A)
longData<-longData[longData$value!=0,]
ggplot(longData, aes(x = Var2, y = Var1)) + 
  geom_raster(aes(fill=value)) + 
  scale_fill_gradient(low="grey90", high="red") +
  labs(x="Antibodies", y="Cell Types", title="Matrix") +
  theme_bw() + theme(axis.text.x=element_text(size=9, angle=90, vjust=0.3),
                     axis.text.y=element_text(size=9),
                     plot.title=element_text(size=11))

png("ReferenceMatrix.png")
ggplot(longData, aes(x = Var2, y = Var1)) + 
  geom_raster(aes(fill=value)) + 
  scale_fill_gradient(low="grey90", high="red") +
  labs(x="Antibodies", y="Cell Types", title="Matrix") +
  theme_bw() + theme(axis.text.x=element_text(size=9, angle=90, vjust=0.3),
                     axis.text.y=element_text(size=9),
                     plot.title=element_text(size=11))
dev.off()



```






```{r}
# for fun try some other things from the tutorial 
# try to cluster the matrix

library(igraph)
#define a graph that represented as adjacency matrix with matrix A
g <- graph.incidence(A, weighted = TRUE)
#cluster wit Louvain algorithm
lou <- cluster_louvain(g)
df.lou <- data.frame(lou$names,lou$membership)

# After that join cluster information to longData that we want to plot.
library(dplyr)

#the same longData than earlier
longData <- left_join(longData, df.lou, by=c("Var1"="lou.names"))
colnames(longData)[4] <- "Var1_clust"
longData$Var2 <- as.factor(longData$Var2)
longData <- left_join(longData, df.lou, by=c("Var2"="lou.names"))
colnames(longData)[5] <- "Var2_clust"
longData$colour <- ifelse(longData$Var1_clust==longData$Var2_clust, longData$Var1_clust, 0)

# now plot the cluster
longData$Var1 <- factor(longData$Var1, levels=unique(arrange(longData, Var1_clust)[,1]))
longData$Var2 <- factor(longData$Var2, levels=unique(arrange(longData, Var2_clust)[,2]))
#levels must be names
longData$colour <- factor(longData$colour)
#for colours variabes must be factors (discrete scale) otherwise ggplot recognize it continous

ggplot(longData, aes(x = Var2, y = Var1, fill=colour)) + 
  geom_raster() + 
  scale_fill_manual(values=c("grey80", "#B40404", "#0B6121", "#FFBF00", "blue")) +
  labs(x="letters", y="LETTERS", title="Matrix") +
  theme_bw() + theme(axis.text.x=element_text(size=9, angle=0, vjust=0.3),
                     axis.text.y=element_text(size=9),
                     plot.title=element_text(size=11),
                     legend.text=element_text(size=7))



```






```{r}
# testing the function for above 
df.value <- fc[0:13,3:13]
new <- apply(df.value, MARGIN = 1, FUN = function(X) (X - min(X, na.rm=TRUE))/diff(range(X, na.rm=TRUE)))
head(new)

```





Now I'll take the mean values

```{r}
# we need to make an array
#  It takes rows(x), columns(y), and a number of matrices of dimensions nrow * ncol(z).

arr <- array(as.numeric(unlist(norm.list)), dim = c(13, 12, 5))


all.matrix <- abind(arr, along=3)
df.mean <- apply(all.matrix, c(1,2), mean,na.rm=TRUE)
head(df.mean)

# I have some values now - I need to check them against doing it individually in the matrices
aqp4 <- c(bb.norm[1,2],ucsc.norm[1,2],mbo.47.norm[1,2],mbo.160.norm[1,2],fc.norm[1,2])
mean.aqp4 = mean(na.omit(aqp4))
mean.aqp4

# yes it has worked but column 1 is the markers - which all turned into NaN

# add the column names back in
# get column names from an original matrix
col.names.df <- as.vector(colnames(bb))
colnames(df.mean) <- col.names.df 
# remove the NaN column  "Marker"
mean.exp <- df.mean[,2:12]
# replace the antibody names
# get the marker names
markers <- as.vector(bb.norm['Marker'])
mean.exp<- cbind(markers,mean.exp)
head(mean.exp)

```






I want to visualize the data I'll need to make a 'long table'

```{r}
library(tidyr)
df.long <- as.data.frame(pivot_longer(mean.exp, cols=2:12, names_to = "Cell Type", values_to = "Expression" ))

write.csv(df.long, "scRNA_public_data_average_long.csv")

test.df <- read.csv("scRNA_public_data_average_long.csv")

test.df['Marker'] <- as.factor(test.df['Marker'])
test.df['Cell.Type'] <- as.factor(test.df['Cell.Type'])

ggplot(data= test.df, mapping = aes(x = 'Cell.Type', y= 'Expression', color= 'Marker')) + geom_boxplot()



ggplot(test.df, aes(x = 'Marker', y= 'Cell Type')) +
  geom_raster(aes(fill='Expression')) + 
  scale_fill_gradient(low="grey90", high="red") +
  labs(x="letters", y="LETTERS", title="Matrix") +
  theme_bw() + theme(axis.text.x=element_text(size=9, angle=0, vjust=0.3),
                     axis.text.y=element_text(size=9),
                     plot.title=element_text(size=11))

  

  
```





```{r}

print(colnames(fb3))
print(colnames(fb5))
# column names were not the same
fb5 <- fb5 %>% rename(Epithelial = Edpendymal.Epithealial)

# rbind worked to join two df but it doesn't add anything automatically to indicate where the data came from these are just turning into replicates 
# add source column
fb5$Source <- "fb5"
head(fb5)


fb3 <- fb3 %>% rename(Epithelial = Edpendymal)
# add source column for fb3
fb3$Source <- "fb3"

```
Z-score each matrix get a mean of 0 and std of 1 

```{r}

sdat <- scale(fb3)


rm.cat <- fb3 %>% select(-"Annotation",-"Marker",-"Source")
scale.data <- scale(rm.cat)

rowMeans(scale.data, na.rm = TRUE)
colMeans(scale.data, na.rm = TRUE)

tryrow <- apply(rm.cat,1,scale)
rowMeans(tryrow, na.rm = TRUE)

df <-rm.cat

# margin is the dimension - 1 is rows and 2 is columns 

# example of apply
zscore <- apply(df, 1, max, na.rm = TRUE)

zscore  
# example of a function

# z = y- mean(y) /std 

# z = x - mean/ sd

# this might work more easily if I transpose the table so I z score the columns
tdf <- t(df)


tdf$z1 <- (tdf[1] - mean(tdf[1], na.rm = TRUE))/ sd(tdf[1], na.rm = TRUE)


zscorefun <- function(df) {
  new.df <- 
  return(new.df)
}

df.scale2 <- (apply(rm.cat, 1, x - mean/sd))

```



```{r}

twodf <- rbind(fb3,fb5)
head(twodf)

# perfect this works
# now I realize I need to z score first


```

