---
title: "R Notebook"
output: html_notebook
---



```{r}
# load libraries



```

Hypergate takes in 
1. an expression matrix
2. a vector of events to attempt to gate on - there are different ways to get these

interactive gating - try later
Clustering - this is what I want to use

hypergate is the function to optomize gating strategies

xp = a numberic matrix encoding expression 
gate_vector a vector with a few unique values --- this should be the cluster labels
level specifies what value of gate vector togate upon 

```{r}
library("hypergate")

input.xm = as.matrix(GetAssayData(seu.q, slot = 'scale.data'))
xm.t <- t(input.xm)
cluster.labels <- as.vector(seu.q$labels)


hg_output <- hypergate(xp = xm.t, gate_vector = cluster.labels, level = 'Astro1')

# writing level = 1 resulted empty output
# putting 'Astro1' is running but taking a very long time
# I believe I will have to down sample the Seurat object.  I thing I might want to subset the AIW002 first



```


Try checking the results

```{r}
hgate_pheno(hg_output)

```



