---
title: "R Notebook Create Reference matrix"
output: html_notebook
---

Here I will read in each of the datasets
1. 2D Flow
2. BenBarres bulk RNA from human isolated cells
3. scRNAseq from in house MBO 47 days
4. scRNAseq public data from human brain tissues - multiple datasets to be merged
5. scRNAseq from public data from cerebral organoids

Each dataset will normalized from 0-1

```{r}
# load libraries
library(dplyr)
library(ggplot2)

```



```{r}
# I'll start with the scRNAseq from brain tissue because there are multiple datasets. From human cell landscape
# values manually copied by cluster
# all neuron averaged, all astrocytes averaged, all radial glia averaged. 

fb3 <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/fetalbrain3.csv")
fb5 <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/fetalbrain5.csv")
lake <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/Lake.csv")
lamanno <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/lamanno.csv")
zhang <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/HumanCellLandscape/zhang.csv")

# column names each of the 5 HCL datasheets
colnames(fb3)
colnames(fb5)
colnames(lake)
colnames(lamanno)
colnames(zhang)

#cell_types

```

Get the markers names - Need to be sure that are all in the same order

```{r}
head(fb3)

markers <- as.vector(fb3['Marker'])
class(markers)
markers



```



Transform values to be between 0 and 1 

```{r}

df.list <- list(fb3,fb5,lake, lamanno, zhang)
df.names <- c("fb3","fb5","lake", "lamanno", "zhang")

# normalize each df

norm.list <- lapply(df.list, function(x) {
  source.name <- df.names[df]
  # get marker names
  markers <- as.vector(x['Marker'])
  # make a df with no character to normalize
  df.value <- x[0:13,3:13]
  # create normalized matrix
  zero.one <- apply(df.value, MARGIN = 2, FUN = function(X) (X - min(X))/diff(range(X)))
  x <- cbind(markers,zero.one)
  x

  })


#I now have a list of normized df 
fb3.norm <- norm.list[[1]]
class(fb3.norm)
head(fb3.norm)

fb5.norm <- norm.list[[2]]
lake.norm <- norm.list[[3]]
lamanno.norm <- norm.list[[4]]
zhang.norm <- norm.list[[5]]

fb3.norm[1,2] # row, column
fb5.norm[1,2]
lake.norm[1,2]
lamanno.norm[1,2]
zhang.norm[1,2]

aqp4 <- c(fb3.norm[1,2],fb5.norm[1,2],lake.norm[1,2],lamanno.norm[1,2],zhang.norm[1,2])

mean.aqp4 = mean(na.omit(aqp4))
mean.aqp4
markers <- as.vector(lake.norm['Marker'])

col.names.df <- as.vector(colnames(lake.norm))


```

Next I will get the mean values for each celltype/marker from across the 5 scRNAseq dataset (public)

```{r}

# we need to make an array
#  It takes rows(x), columns(y), and a number of matrices of dimensions nrow * ncol(z).

arr <- array(as.numeric(unlist(norm.list)), dim = c(13, 12, 5))

#ans1 <- apply(lapply(arr, as.matrix), c(2, 3), mean)
#ans1 = aaply(laply(all.dat, as.matrix), c(2, 3), mean) 

library("abind")

all.matrix <- abind(arr, along=3)
df.mean <- apply(all.matrix, c(1,2), mean,na.rm=TRUE)
head(df.mean)

# I have some values now - I need to check them
# yes it has worked but column 1 is the markers - which all turned into NaN

# add the column names back in
colnames(df.mean) <- col.names.df 
head(df.mean)
# remove the NaN column
mean.exp <- df.mean[,2:12]
# replace the antibody names
mean.exp<- cbind(markers,mean.exp)
head(mean.exp)

write.csv(mean.exp, "scRNA.pub.mean.exp.csv")





```


Now I need to prepare each of the other expression matrices in the same way


```{r}
bb <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/Inputs_for_reference_combined/BenBarres_marker_exp_input.csv")

ussc <- read.csv("/home/rhalena/Documents/Documents/MyPapers/FACS/Analysis/Reference_matrix/Inputs_for_reference_combined/UCSC_organoid_average_by_cluster.csv")


head(bb)
head(ussc)



```








I want to visualize the data I'll need to make a 'long table'

```{r}
library(tidyr)
df.long <- as.data.frame(pivot_longer(mean.exp, cols=2:12, names_to = "Cell Type", values_to = "Expression" ))

write.csv(df.long, "scRNA_public_data_average_long.csv")

test.df <- read.csv("scRNA_public_data_average_long.csv")

test.df['Marker'] <- as.factor(test.df['Marker'])
test.df['Cell.Type'] <- as.factor(test.df['Cell.Type'])

ggplot(data= test.df, mapping = aes(x = 'Cell.Type', y= 'Expression', color= 'Marker')) + geom_boxplot()



ggplot(test.df, aes(x = 'Marker', y= 'Cell Type')) +
  geom_raster(aes(fill='Expression')) + 
  scale_fill_gradient(low="grey90", high="red") +
  labs(x="letters", y="LETTERS", title="Matrix") +
  theme_bw() + theme(axis.text.x=element_text(size=9, angle=0, vjust=0.3),
                     axis.text.y=element_text(size=9),
                     plot.title=element_text(size=11))

  

  
```





```{r}

print(colnames(fb3))
print(colnames(fb5))
# column names were not the same
fb5 <- fb5 %>% rename(Epithelial = Edpendymal.Epithealial)

# rbind worked to join two df but it doesn't add anything automatically to indicate where the data came from these are just turning into replicates 
# add source column
fb5$Source <- "fb5"
head(fb5)


fb3 <- fb3 %>% rename(Epithelial = Edpendymal)
# add source column for fb3
fb3$Source <- "fb3"

```
Z-score each matrix get a mean of 0 and std of 1 

```{r}

sdat <- scale(fb3)


rm.cat <- fb3 %>% select(-"Annotation",-"Marker",-"Source")
scale.data <- scale(rm.cat)

rowMeans(scale.data, na.rm = TRUE)
colMeans(scale.data, na.rm = TRUE)

tryrow <- apply(rm.cat,1,scale)
rowMeans(tryrow, na.rm = TRUE)

df <-rm.cat

# margin is the dimension - 1 is rows and 2 is columns 

# example of apply
zscore <- apply(df, 1, max, na.rm = TRUE)

zscore  
# example of a function

# z = y- mean(y) /std 

# z = x - mean/ sd

# this might work more easily if I transpose the table so I z score the columns
tdf <- t(df)


tdf$z1 <- (tdf[1] - mean(tdf[1], na.rm = TRUE))/ sd(tdf[1], na.rm = TRUE)


zscorefun <- function(df) {
  new.df <- 
  return(new.df)
}

df.scale2 <- (apply(rm.cat, 1, x - mean/sd))

```



```{r}

twodf <- rbind(fb3,fb5)
head(twodf)

# perfect this works
# now I realize I need to z score first


```

