---
title: "R Notebook"
output: html_notebook
---

PhenoID supplemental Figures

S1 - heatmap of 9 MBO samples

```{r}
# Figure S1 
# heatmap of all 9 samples


# S1
# set up the environment 

library(Seurat)
library(dplyr)
library(ggplot2)
library(reshape2)

# read in the annotated data ojbect
seu <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outs/clusters/AllcellLablesMarch25.Rds")

AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

Idents(seu) <- 'Batch'
seu.sub <- subset(x= seu, downsample = 500)
#DoHeatmap(seu.sub, features = AB.order, group.by = "Batch")

DoHeatmap(seu.sub, group.by = "Batch", features = AB, size= 6,slot = "scale.data", group.colors = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"), 
          disp.max = 1.5, disp.min = -1.1) + 
  scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 10))


# save
outpath = "/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/"
pdf(paste(outpath,"HM9hMOs.pdf",sep = ""))
DoHeatmap(seu.sub, group.by = "Batch", features = AB, size= 6,slot = "scale.data", group.colors = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"), 
          disp.max = 1.5, disp.min = -1.1) + 
  scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 10))
dev.off()




```


S8 - UMAPs and heatmaps from 3 pre-processing methods


```{r}
# 
output_path <- "/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/"

input_name <- "cat"  # processing type for file name
clust_method <- "Louvain" # cluster type for file name

input_path <- "/Users/rhalenathomas/GITHUB/PhenoID_single_cell_flow_cytometry_analysis/Old/preprocessing/outputs/prepro_outsflowset.csv"
df <- read.csv(input_path) # read in the dataframe

# create a df with just the expression
# need a way to automate this selection
# I only want the expression values
df2 <- df %>% dplyr::select(c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4"))


# the order of the DF is set by the order the columns are written above
# create a matrix for later
m <- as.matrix(df2)

# create the seurat object for visualization
tm <- t(df2)
rownames(tm) <- colnames(df2)
colnames(tm) <- rownames(df2)
seu <- CreateSeuratObject(tm)

df$Batch <- as.factor(df$Batch) # so that when added to seurat object batch has levels
# add the meta data back in for sample groups
seu <- AddMetaData(object=seu, metadata=df$Batch, col.name = 'Sample')
# create the vector for the antibodies names for feature plotting later
AB <- colnames(df2)
# add to scale data slot
seu <- ScaleData(seu)
Idents(seu) <- 'Sample'

# see the plot
DoHeatmap(seu, group.by = 'Sample', features = AB, size= 5,slot = "scale.data",
        angle = 90, group.colors = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"),label = FALSE,
        hjust = 0, group.bar.height = 0.05, disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 13)) 
  
# save the plot
png(paste(output_path,"HM_concat.png"), height = 600, width = 800)
DoHeatmap(seu, group.by = 'Sample', features = AB, size= 5,slot = "scale.data",
        angle = 90, group.colors = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"),label = FALSE,
        hjust = 0, group.bar.height = 0.05, disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 13)) 
dev.off()

# create PCA 
seu <- RunPCA(seu, features = AB, npcs = 12, approx = FALSE)
seu <- FindNeighbors(seu, dims = 1:12, k.param = 60)
seu <- FindClusters(seu, resolution = 0.2)
seu <- RunUMAP(seu, dims = 1:12, n.neighbors = 60L)

DimPlot(seu, reduction = "umap")

DimPlot(seu, reduction = "umap", group.by = 'Sample', cols = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"))


png(paste(output_path,"UMAP_concat.png"), height = 600, width = 800)
DimPlot(seu, reduction = "umap", group.by = 'Sample', cols = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"))+ 
  theme(axis.text.y = element_text(size = 12),axis.text.x = element_text(size = 12))
dev.off()

########## transform without retro transform

input_name <- "trans"  # processing type for file name
clust_method <- "Louvain" # cluster type for file name

input_path <- "/Users/rhalenathomas/GITHUB/PhenoID_single_cell_flow_cytometry_analysis/Old/preprocessing/outputs/prepro_outstransformed_flowset.csv"
df <- read.csv(input_path) # read in the dataframe

# create a df with just the expression
# need a way to automate this selection
# I only want the expression values
df2 <- df %>% dplyr::select(c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4"))


# the order of the DF is set by the order the columns are written above
# create a matrix for later
m <- as.matrix(df2)

# create the seurat object for visualization
tm <- t(df2)
rownames(tm) <- colnames(df2)
colnames(tm) <- rownames(df2)
seu <- CreateSeuratObject(tm)

df$Batch <- as.factor(df$Batch) # so that when added to seurat object batch has levels
# add the meta data back in for sample groups
seu <- AddMetaData(object=seu, metadata=df$Batch, col.name = 'Sample')
# create the vector for the antibodies names for feature plotting later
AB <- colnames(df2)
# add to scale data slot
seu <- ScaleData(seu)
Idents(seu) <- 'Sample'

# see the plot
DoHeatmap(seu, group.by = 'Sample', features = AB, size= 5,slot = "scale.data",
        angle = 90, group.colors = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"),label = FALSE,
        hjust = 0, group.bar.height = 0.05, disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 13)) 
  
# save the plot
png(paste(output_path,"HM_trans.png"), height = 600, width = 800)
DoHeatmap(seu, group.by = 'Sample', features = AB, size= 5,slot = "scale.data",
        angle = 90, group.colors = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"),label = FALSE,
        hjust = 0, group.bar.height = 0.05, disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 13)) 
dev.off()

# create PCA 
seu <- RunPCA(seu, features = AB, npcs = 12, approx = FALSE)
seu <- FindNeighbors(seu, dims = 1:12, k.param = 60)
seu <- FindClusters(seu, resolution = 0.2)
seu <- RunUMAP(seu, dims = 1:12, n.neighbors = 60L)

DimPlot(seu, reduction = "umap")

DimPlot(seu, reduction = "umap", group.by = 'Sample', cols = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"))+ 
  theme(axis.text.y = element_text(size = 12),axis.text.x = element_text(size = 12))

png(paste(output_path,"UMAP_trans.png"), height = 600, width = 800)
DimPlot(seu, reduction = "umap", group.by = 'Sample', cols = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"))+ 
  theme(axis.text.y = element_text(size = 12),axis.text.x = element_text(size = 12))
dev.off()

########## transform and retro transform #############

input_name <- "retro"  # processing type for file name
clust_method <- "Louvain" # cluster type for file name

input_path <- "/Users/rhalenathomas/GITHUB/PhenoID_single_cell_flow_cytometry_analysis/Old/preprocessing/outputs/prepro_outsretrotransformed_flowset.csv"
df <- read.csv(input_path) # read in the dataframe

# create a df with just the expression
# need a way to automate this selection
# I only want the expression values
df2 <- df %>% dplyr::select(c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4"))


# the order of the DF is set by the order the columns are written above
# create a matrix for later
m <- as.matrix(df2)

# create the seurat object for visualization
tm <- t(df2)
rownames(tm) <- colnames(df2)
colnames(tm) <- rownames(df2)
seu <- CreateSeuratObject(tm)

df$Batch <- as.factor(df$Batch) # so that when added to seurat object batch has levels
# add the meta data back in for sample groups
seu <- AddMetaData(object=seu, metadata=df$Batch, col.name = 'Sample')
# create the vector for the antibodies names for feature plotting later
AB <- colnames(df2)
# add to scale data slot
seu <- ScaleData(seu)
Idents(seu) <- 'Sample'

# see the plot
DoHeatmap(seu, group.by = 'Sample', features = AB, size= 5,slot = "scale.data",
        angle = 90, group.colors = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"),label = FALSE,
        hjust = 0, group.bar.height = 0.05, disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 13)) 
  
# save the plot
png(paste(output_path,"HM_retro.png"), height = 600, width = 800)
DoHeatmap(seu, group.by = 'Sample', features = AB, size= 5,slot = "scale.data",
        angle = 90, group.colors = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"),label = FALSE,
        hjust = 0, group.bar.height = 0.05, disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 13)) 
dev.off()

# create PCA 
seu <- RunPCA(seu, features = AB, npcs = 12, approx = FALSE)
seu <- FindNeighbors(seu, dims = 1:12, k.param = 60)
seu <- FindClusters(seu, resolution = 0.2)
seu <- RunUMAP(seu, dims = 1:12, n.neighbors = 60L)

DimPlot(seu, reduction = "umap")

DimPlot(seu, reduction = "umap", group.by = 'Sample', cols = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"))+ 
  theme(axis.text.y = element_text(size = 12),axis.text.x = element_text(size = 12))

png(paste(output_path,"UMAP_retro.png"), height = 600, width = 800)
DimPlot(seu, reduction = "umap", group.by = 'Sample', shuffle = TRUE, cols = c("cadetblue1","cadetblue2","cadetblue3", "bisque1","bisque2","bisque3","darkolivegreen1","darkolivegreen2","darkolivegreen3"))+ 
  theme(axis.text.y = element_text(size = 12),axis.text.x = element_text(size = 12))
dev.off()




```


Figure 3 related Sup
Correlation 

```{r}
output_path <- "/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/Corr/"

# read in correlation output
corr_cell.df <- read.csv("/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/Corr/corr_celltypes.csv")
freq.df <- read.csv("/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/Corr/Frequencytabletypes.csv")
# plot the main groups - but the correlation co-efficient for the assigned group
df <- corr_cell.df %>% filter(!grepl('-',cell.label))

# figure S9A 
# showing the assigned cell types main cell types only
pdf(paste(output_path,"Vln.cor.assigned.pdf"),height = 4, width = 10)
ggplot(df, aes(x=cell.label, y=cor.1, fill = best.cell.type))+ geom_violin(trim = FALSE)+ ylim(0,1)+theme_classic()+
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90, size = 12)) + ylab("correlation coefficient") + xlab("Assigned Cell Type") +
  guides(fill = guide_legend(title = "Cell Phenotype"))
dev.off()

# figure S9B
# box plot of assigned cell type correlation with 
pdf(paste(output_path,"box.cor.alltypes.assign.pdf"),height = 4, width = 10)
ggplot(corr_cell.df, aes(x=cell.label, y=cor.1))+ geom_boxplot()+ ylim(0,1)+theme_classic()+
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90, size = 12)) + ylab("correlation coefficient") + xlab("Assigned Cell Type") +
  guides(fill = guide_legend(title = "Cell Phenotype")) 
dev.off()

## Figure S10 and S11 connected columns are in the functions and saved from there


## frequency bar charts
df <- corr_cell.df

# remove different elements --- must have at least 100 cells
 df.filter <- df %>% group_by(cell.label) %>% dplyr::filter(n()> 100)
  # # plot
  pdf(paste(output_path,"FreqCellTypes2.pdf",sep=""),width =12, height = 6)
  plot1 <- ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90))+ xlab('Assigned cell type') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Cell Phenotype")) 
  print(plot1)
  dev.off()

## remove the unknown cell types
df.filter <- df %>% filter(!grepl('unknown',cell.label))

pdf(paste(output_path,"BarFreq.allknown.pdf",sep=""),width =12, height = 6)
  plot1 <- ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90, hjust = 1))+ xlab('Assigned cell type') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Cell Phenotype")) 
  print(plot1)
dev.off()


## plot the frequency 
df.filter <- df %>% filter(!grepl('unknown',cell.label))
# filter for only frequent double cell types
df.filter <- df.filter %>% group_by(cell.label) %>% dplyr::filter(n()< 500)

pdf(paste(output_path,"BarFreq.lowfreq.pdf",sep=""),width =12, height = 6)
  plot1 <- ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90, hjust = 1))+ xlab('Assigned cell type') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Cell Phenotype")) 
  print(plot1)
dev.off()




```


Figure 3 cluster annotation
Visualize and quantify correlation
Heatmap
Feature plots


```{r}
# output path for figures
output_path <- "/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/"


# visualize correlation with clusters
seu <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/cluster_parameters/retro-louv-moreparm/retrotransLouvainSeuratObject60.Rds")

seu <- RunUMAP(seu,spread = 1, min.dist = 0.05, dims = 1:12)

# now look at the correlation assignments
cor <- read.csv("/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/Corr/corr_celltypes.csv")

# add the predicted correlation cell types to the seurat object
seu <- AddMetaData(object=seu, metadata=cor$cell.label, col.name = 'cor.labels')
# check the labels
unique(seu$cor.labels)
# plot the labels on the UMAP

#Figure S13 A
png(paste(output_path,"UMAPcorlabels.png"), width = 1100, height = 500)
DimPlot(seu, group.by = 'cor.labels', label = TRUE, repel = TRUE)
dev.off()

# Figure S13B 
# Bar chart of the cell types in the different clusters
# get the number of cells with each predictions per cluster
cor.lables <- as.data.frame(table(seu$RNA_snn_res.0.7, seu$cor.labels))
pr.cor.lables <- as.data.frame(prop.table(table(seu$RNA_snn_res.0.7, seu$cor.labels)))

# rename the df
colnames(cor.lables) <- c("Cluster","Predicted.Cell.Type","Number.Cells")


pdf(paste(output_path,"BarcorlabelsbyCluster.pdf"), width = 10, height = 3)
ggplot(cor.lables, aes(y = Number.Cells, x =Cluster, fill = Predicted.Cell.Type)) + geom_bar(position = "stack", stat= "identity") +
  theme_classic() +
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90, hjust = 0))+ xlab('Cluster Number') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Correlation Predicted Cell Type")) 
dev.off()

# try to filter out the unknown
df.filter <- cor.lables %>% filter(!grepl('unknown',Predicted.Cell.Type))

# Figure S13C
pdf(paste(output_path,"BarcorlabelsbyCluster.rm.unknown.pdf"), width = 10, height = 3)
ggplot(df.filter, aes(y = Number.Cells, x =Cluster, fill = Predicted.Cell.Type)) + geom_bar(position = "stack", stat= "identity") +
  theme_classic() +
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90, hjust = 0))+ xlab('Cluster Number') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Correlation Predicted Cell Type")) 
dev.off()

# create a table with the most frequent cell type prediction ignoring unknown
top.cells <- df.filter  %>% group_by(Cluster) %>% top_n(3, Number.Cells)

top.cell <- df.filter  %>% group_by(Cluster) %>% top_n(1, Number.Cells)
# put in excel sheet to predict cell types


### visualize expression 

# feature plots
# Figure S14
png(paste(output_path,"FtUMAP9000cells.neuron.png"), width = 1100, height = 200)
FeaturePlot(seu, features = c("CD24","CD56","CD29","CD15","CD71"), slot = 'scale.data',min.cutoff = 'q3', max.cutoff ='q97', ncol = 5)
dev.off()

png(paste(output_path,"FtUMAP9000cells.astro.png"), width = 1100, height = 200)
FeaturePlot(seu, features = c("CD44","GLAST","AQP4","HepaCAM","CD133"), slot = 'scale.data',min.cutoff = 'q3', max.cutoff ='q97', ncol = 5)
dev.off()


png(paste(output_path,"FtUMAP9000cells.olig.png"), width = 900, height = 200)
FeaturePlot(seu, features = c("O4","CD140a","CD184"), slot = 'scale.data',min.cutoff = 'q3', max.cutoff ='q97', ncol = 5)
dev.off()

Idents(seu) <- 'RNA_snn_res.0.7'
png(paste(output_path,"UMAP9000cell.clusters07.png"), width = 500, height = 300)
DimPlot(seu, label = TRUE, repel = TRUE, label.size = 5) + NoLegend()
dev.off()


#### now make the heatmap for these clusters to annotate

AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

# Figure S15

pdf(paste(output_path,"HM.9000.cluster07.pdf"),width = 10, height = 6)
Idents(seu) <- 'RNA_snn_res.0.7'
DoHeatmap(seu, features = AB, size= 6,slot = "scale.data", disp.max = 2, disp.min = -1.5, angle = 0) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 15))
dev.off()


```

2D cell correlation test and assignements

```{r}


reference_path <- "/Users/rhalenathomas/GITHUB/PhenoID_single_cell_flow_cytometry_analysis/old/correlation/ReferenceMatrix9celltypesOrdered.csv"
output_path <- "/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/Corr2D/"


test_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/2Dcells_surface/preprocessing/select/2DcellsSelectflowset.csv"

find_correlation(test_path, reference_path, output_path)


# this produces a csv with the Frequencies of each cell type and a csv with the best and second best correlation coefficients
# read in those csv
corr_cell.df <- read.csv("/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/Corr2D/corr_celltypes.csv")
freq.df <- read.csv("/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/Corr2D/Frequencytabletypes.csv")


# plot the main groups - but the correlation co-efficient for the assigned group
df <- corr_cell.df %>% filter(!grepl('-',cell.label))


ggplot(df, aes(x=cell.label, y=cor.1 ))+ geom_violin()+ ylim(-0.1,1)+theme_classic()+
   theme(axis.text.x=element_text(angle=90)) + ylab("correlation coefficient") + xlab("Cell type with max correlation coefficient")


thresh1 <- 0.45
ggplot(df, aes(x=best.cell.type, y=cor.1, fill = best.cell.type))+ geom_violin(trim = FALSE)+ ylim(0,1)+theme_classic()+
   theme(text = element_text(size = 18), axis.text.x=element_text(angle=90, size = 15)) + ylab("correlation coefficient") + xlab("Cell type with max correlation coefficient") +
  geom_hline(yintercept = thresh1) +
  guides(fill = guide_legend(title = "Cell Phenotype"))

#save this plot for S13
pdf(paste(output_path,"Vln.max.cor.2Dcells.cells.pdf"),height = 4, width = 6)
ggplot(df, aes(x=best.cell.type, y=cor.1, fill = best.cell.type))+ geom_violin(trim = FALSE)+ ylim(0,1)+theme_classic()+
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90, size = 12)) + ylab("correlation coefficient") + xlab("Cell type with max correlation coefficient") +
  geom_hline(yintercept = thresh1) +
  guides(fill = guide_legend(title = "Cell Phenotype"))
dev.off()

### compare correlation with the cell types 

# add the correlation predictions to the meta data

seu <- AddMetaData(object=seu, metadata=corr_cell.df$cell.label, col.name = 'cor.labels')
seu <- AddMetaData(object=seu, metadata=corr_cell.df$best.cell.type, col.name = 'main.cor.labels')

### save with the correlation labels 
png(paste(output_path,"UMAPcor.lables.png"), width = 800, height = 300)
DimPlot(seu, group.by = 'cor.labels')
dev.off()


DimPlot(seu, group.by = 'Batch')
DimPlot(seu, group.by = 'Lables')

# show how many cells are predict in each cell type

t.lables <- as.data.frame(table(seu$Batch, seu$cor.labels))
pr.t.lables <- as.data.frame(prop.table(table(seu$Batch, seu$cor.labels)))

# try bar chart
png(paste(output_path,"Bar.cor.lables.input.png"), width = 900, height = 300)
ggplot(t.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity") + 
  theme_classic() + theme(text = element_text(size = 15), axis.text.x=element_text(hjust = 1, size = 15),axis.text.y=element_text(size = 15) ) + 
  xlab('2D cell culture') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Correltation Predicted Cell Type")) 
dev.off()

# with only the main cell types
t.lables <- as.data.frame(table(seu$Batch, seu$main.cor.labels))
pr.t.lables <- as.data.frame(prop.table(table(seu$Batch, seu$main.cor.labels)))

# try bar chart
png(paste(output_path,"Bar.cor.best.cell.input.png"), width = 600, height = 300)
ggplot(t.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity") + 
  theme_classic() + theme(text = element_text(size = 18), axis.text.x=element_text(hjust = 1, size = 18, angle = 90),axis.text.y=element_text(size = 18), legend.text = element_text(size=18)) + 
  xlab('2D cell culture') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Correltation Predicted Cell Type")) 
dev.off()




# find the top 3 in cell types predicted in each cluster 

top.labs <- t.lables  %>% group_by(Var1)  %>% top_n(3, Freq)
top.labs

# find the top cell in each cluster
top.lab <- t.lables  %>% group_by(Var1)  %>% top_n(1, Freq)
top.lab







```








Get the cell counts

```{r}

table(seu$Batch)
# total cells 
# total cells 73578

```


Cluster annotation of the 9000 cells from 9 hMOs

```{r}
# read in the annotated object
seu <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/Seu9000annot.08072021.RDS")


# more subgroups annotated
Idents(seu) <- "RNA_snn_res.0.7"
cluster.ids <- c("Unknown","Mixed","Neurons 1","Radial Glia 1","Epithelial","Astrocytes 1",
                 "Neurons 2","Astrocytes 2","Astrocytes 3","Astrocytes mature","Neurons 3","NPC",
                 "Radial Glia 2","Radial Glia 3","Endothelial",
                 "Oligodendrocytes","Stem-like 1","Stem-like 2", "Neural stem")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$subgroups <- Idents(seu)

# 19 levels - every cluster is it's own subgroup
# change the order of the cell types on the legend of the umap
cell.type.order <- c("Astrocytes 1", "Astrocytes 2","Astrocytes 3","Astrocytes mature",
                     "Radial Glia 1","Radial Glia 2","Radial Glia 3",
                     "Epithelial","Endothelial",
                     "Neural stem","NPC","Neurons 1","Neurons 2","Neurons 3",
                     "Oligodendrocytes","Stem-like 1","Stem-like 2",
                     "Mixed","Unknown")

cell.type.order <- rev(cell.type.order)

# colour order to match cell type order
clust.colours <- c("chocolate1","chocolate2","chocolate3","orangered",
                   "lightsalmon", "pink","lightpink3",
                   "steelblue3","deepskyblue",
                   "plum","plum3","purple","orchid2","mediumpurple2",
                   "seagreen3","tomato3","tomato4","burlywood3","grey90")
# Figure S16 
pdf(paste(output_path,"UMAP.9000.subgroups.pdf"),width = 15, height = 8)
DimPlot(seu, order = cell.type.order, cols = clust.colours, shuffle = TRUE, 
        raster=FALSE, pt.size = 0.1, label = FALSE) +
  theme(legend.text = element_text(size=24), axis.title.y = element_text(size=24), 
        axis.title.x = element_text(size=24), axis.text.y = element_text(size =24),
        axis.text.x = element_text(size =24))
dev.off()


# matching heatmap
levels(seu) <- c("Astrocytes 1", "Astrocytes 2","Astrocytes 3","Astrocytes mature",
                     "Radial Glia 1","Radial Glia 2","Radial Glia 3",
                     "Epithelial","Endothelial",
                     "Neural stem","NPC","Neurons 1","Neurons 2","Neurons 3",
                     "Oligodendrocytes","Stem-like 1","Stem-like 2",
                     "Mixed","Unknown")

pdf(paste(output_path,"HM9000.subtypes.08072022.pdf"),width = 8, height = 5)
DoHeatmap(seu, features = AB, size= 6,slot = "scale.data", group.colors = clust.colours, disp.max = 2, disp.min = -1.5,
          angle = 90) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 15))
dev.off()



saveRDS(seu, "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/Seu9000annot.08072021.RDS")


```

Random Forest

```{r}
# training run by the function - in the Figure 4 workbook

# read in the model

rf <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/RFM_trained.11072022.Rds")

output_path <- "/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/"

# now the predictions
p1 <- predict(rf, train)
confusionMatrix(p1, train$lables)

p2 <- predict(rf, test)
c2 <- confusionMatrix(p2, test$lables)
c2.table <- as.data.frame(c2$table)

# try to plot the results

library(ggplot2)
ggplot(data =  c2.table, mapping = aes(x = Prediction, y = Reference, fill= Freq)) +
  geom_tile(aes(fill = Freq), colour = "white") +
  geom_text(aes(label = sprintf("%1.0f",Freq)), vjust = 1) +
  scale_fill_gradient(low = "white", high = "steelblue") + theme(axis.text.x = element_text(angle = 90))

### figure S17 A

# save the plot
pdf(paste(output_path, "RFMconfustion.Test.results.pdf",sep=""), width = 8, height = 6)
ggplot(data =  c2.table, mapping = aes(x = Prediction, y = Reference, fill= Freq)) +
  geom_tile(aes(fill = Freq), colour = "white") +
  geom_text(aes(label = sprintf("%1.0f",Freq)), vjust = 0.5, colour = "white", size = 12) +
  scale_fill_gradient2(low = "purple3", mid = "grey",high = "cyan", midpoint = 2000) +
  theme(axis.text.x = element_text(angle = 90, hjust =1, size = 15), axis.text.y = element_text(size = 15), axis.title.y = element_text(size=18),axis.title.x = element_text(size=18)) 
dev.off()


## evaluate the importance of each feature

importance(rf) 
varImpPlot(rf) # higher values are more important


pdf(paste(output_path, "RFM.importance.pdf",sep=""), width = 6, height = 5)
varImpPlot(rf) # higher values are more important
dev.off()



library(randomForestExplainer)
importance_frame <- measure_importance(rf)
pdf(paste(output_path, "RFM.importanceMDS.pdf",sep=""), width = 6, height = 5)
plot_multi_way_importance(importance_frame, 
                          x_measure = "mean_min_depth", 
                          y_measure = "gini_decrease",
                          size_measure = "p_value", no_of_labels = 13)

dev.off()

#### plot the error rates over number of trees

oob.err.data1 <- data.frame(
  Trees = rep(1:nrow(rf$err.rate), 20), 
  Type = rep(c("OOB","Unknown","Mixed","Neurons 1","Neurons 2","Neurons 3",
               "NPC","Neural stem","Radial Glia 1","Radial Glia 2","Radial Glia 3", 
               "Epithelial","Endothelial","Oligodendrocytes",
               "Astrocytes 1","Astrocytes 2","Astrocytes 3","Astrocytes mature",
               "Stem-like 1","Stem-like 2"), each = nrow(rf$err.rate)),
  Error = c(rf$err.rate[,"OOB"], rf$err.rate[,"Unknown"], rf$err.rate[,"Mixed"], 
            rf$err.rate[,"Neurons 1"],rf$err.rate[,"Neurons 2"],rf$err.rate[,"Neurons 3"],
            rf$err.rate[,"NPC"],rf$err.rate[,"Neural stem"],rf$err.rate[,"Radial Glia 1"],
            rf$err.rate[,"Radial Glia 2"],rf$err.rate[,"Radial Glia 3"],
            rf$err.rate[,"Epithelial"],rf$err.rate[,"Endothelial"],rf$err.rate[,"Oligodendrocytes"],
            rf$err.rate[,"Astrocytes 1"],rf$err.rate[,"Astrocytes 2"],rf$err.rate[,"Astrocytes 3"],
            rf$err.rate[,"Astrocytes mature"],rf$err.rate[,"Stem-like 1"],rf$err.rate[,"Stem-like 2"]
            ))

# this works but we need a way to not have to type in all the cell type names
pdf(paste(output_path, "RFM.error.pdf",sep=""), width = 8, height = 6)
ggplot(data = oob.err.data1, aes(x = Trees, y= Error)) + geom_line(aes(color = Type)) +
  theme_classic() + theme(axis.text.x = element_text(hjust =1, size = 15), axis.text.y = element_text(size = 15), axis.title.y = element_text(size=18),axis.title.x = element_text(size=18),legend.text=element_text(size=15)) 
dev.off()




```


Annotate full dataset - visualize RFM, CAM and seurat label transfer on UMAP

```{r}

# read in data object with predictions RFM, CAM and seurat label transfer
outpath <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/"
seu <- readRDS(paste(outpath,"All9MOannaote.12072022.Rds"))

# improve visualization with UMAP
seu <- RunUMAP(seu,spread = 3, min.dist = 0.005, dims = 1:12)
DimPlot(seu)
#make the UMAPs of the predictions


### Figure S18 UMAPs with predictions ####

# random forest
png(paste(output_path,"UMAP.rfm.pred.png"), width = 600, height = 300)
Idents(seu) <- 'rf.pred'
DimPlot(seu, label = TRUE)
dev.off()

# correlation
png(paste(output_path,"UMAP.cor.pred.png"), width = 900, height = 350)
Idents(seu) <- 'cor.labels'
DimPlot(seu, label = FALSE)
dev.off()

# seurat transfer labels
png(paste(output_path,"UMAP.seurat.pred.png"), width = 600, height = 350)
Idents(seu) <- 'predicted.id'
DimPlot(seu, label = TRUE)
dev.off()

# resolution 1.5
png(paste(output_path,"UMAP.clusters.res1.5.png"), width = 500, height = 300)
Idents(seu) <- 'seurat_clusters'
DimPlot(seu, label = TRUE)
dev.off()

#### cluster number selected by using RAND Index

# calculate the proportion of each cell label in 

# feature plots
# feature plots
# Figure S14
png(paste(output_path,"FtUMAPall.cells.neuron.png"), width = 1100, height = 200)
FeaturePlot(seu, features = c("CD24","CD56","CD29","CD15","CD71"), slot = 'scale.data',min.cutoff = 'q3', max.cutoff ='q97', ncol = 5)
dev.off()

png(paste(output_path,"FtUMAPall.cells.astro.png"), width = 1100, height = 200)
FeaturePlot(seu, features = c("CD44","GLAST","AQP4","HepaCAM","CD133"), slot = 'scale.data',min.cutoff = 'q3', max.cutoff ='q97', ncol = 5)
dev.off()

png(paste(output_path,"FtUMAPall.cells.olig.png"), width = 900, height = 200)
FeaturePlot(seu, features = c("O4","CD140a","CD184"), slot = 'scale.data',min.cutoff = 'q3', max.cutoff ='q97', ncol = 5)
dev.off()


#### heatmap by clusters ####
AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

# Figure S15

#pdf(paste(output_path,"HM.all.cluster1.5.pdf"),width = 10, height = 6)
Idents(seu) <- 'RNA_snn_res.1.5'
png(paste(output_path,"HM.all.cells.png"), width = 900, height = 800)
DoHeatmap(seu, features = AB, size= 6,slot = "scale.data", disp.max = 2, disp.min = -1.5, angle = 0) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 15))
dev.off()

seu.sub <- subset(seu, downsample(n- 200))
DoHeatmap(seu.sub)



```

Explore numbers of cells from RFM

```{r}

# get the counts of each cell type
frq.df <- as.data.frame(table(results.df['rfm.pred']))

head(frq.df)

# plot directly from the prediction df



ggplot(results.df, aes(x = factor(rfm.pred), fill = factor(rfm.pred))) + geom_bar()+theme_classic()+
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90))+ xlab('RFM predicted') + ylab('Number of cells') 


### get the number of cells predicted by RFM in each cluster

# make a df
t.lables <- as.data.frame(table(seu$seurat_clusters, seu$rf.pred))
pr.t.lables <- as.data.frame(prop.table(table(seu$seurat_clusters, seu$rf.pred)))

# try bar chart
pdf(paste(output_path,"Bar.rf.pred.count.all.cells.pdf"), width = 10, height = 4)
ggplot(t.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity") + 
  theme_classic() + theme(text = element_text(size = 15), axis.text.x=element_text(hjust = 1)) + 
  xlab('Cluster Number') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "RFM Predicted Cell Type")) 
dev.off()
# find the top 3 in cell types predicted in each cluster 

top.labs <- t.lables  %>% group_by(Var1)  %>% top_n(3, Freq)
top.labs

# find the top cell in each cluster
top.lab <- t.lables  %>% group_by(Var1)  %>% top_n(1, Freq)
top.lab





```






Explore the correlation assignment output for the full 9 hMO dataset

```{r}

corr_cell.df <- read.csv("/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/Corr_all/corr_celltypes.csv")
df <- corr_cell.df

# remove different elements --- must have at least 100 cells
df.filter <- df %>% group_by(cell.label) %>% dplyr::filter(n()> 100)
  # # plot
 # pdf(paste(output_path,"FreqCellTypes2.pdf",sep=""),width =12, height = 6)
  plot1 <- ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90))+ xlab('Assigned cell type') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Cell Phenotype")) 
  print(plot1)#  dev.off()

## remove the unknown cell types
df.filter <- df %>% dplyr::filter(!grepl('unknown',cell.label))

#pdf(paste(output_path,"BarFreq.allknown.pdf",sep=""),width =12, height = 6)
  plot1 <- ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90, hjust = 1))+ xlab('Assigned cell type') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Cell Phenotype")) 
  print(plot1)
#dev.off()


## plot the frequency 
df.filter <- df %>% dplyr::filter(!grepl('unknown',cell.label))
# filter for only frequent double cell types
df.filter <- df.filter %>% group_by(cell.label) %>% dplyr::filter(n()< 500)

#pdf(paste(output_path,"BarFreq.lowfreq.pdf",sep=""),width =12, height = 6)
  plot1 <- ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90, hjust = 1))+ xlab('Assigned cell type') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Cell Phenotype")) 
  print(plot1)
#dev.off()

# visualize no unknown and top cells so readable
df.filter <- df %>% dplyr::filter(!grepl('unknown',cell.label))
# filter for only frequent double cell types
df.filter <- df.filter %>% group_by(cell.label) %>% dplyr::filter(n() > 500)
  plot1 <- ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
   theme(text = element_text(size = 15), axis.text.x=element_text(angle=90, hjust = 1))+ xlab('Assigned cell type') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Cell Phenotype")) 
  print(plot1)
  
  
  
  # make a df with frequency per cluster
t.lables <- as.data.frame(table(seu$seurat_clusters, seu$cor.labels))
pr.t.lables <- as.data.frame(prop.table(table(seu$seurat_clusters, seu$cor.labels)))

# plot bar chart

pdf(paste(output_path,"Bar.cor.pred.count.all.cells.pdf"), width = 10, height = 4)
ggplot(t.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity") + 
  theme_classic() + theme(text = element_text(size = 15), axis.text.x=element_text(hjust = 1)) + 
  xlab('Cluster Number') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Correlation Predicted Cell Type")) 
dev.off()



# remove unknown for plotting and top label
t.l.filt <- t.lables %>% dplyr::filter(!grepl('unknown',Var2))

# get ride of low frequency cells too
t.l.filt <- t.l.filt %>% dplyr::filter(Freq > 100)
# make bar chart
pdf(paste(output_path,"Bar.cor.pred.count.all.cells.filt.pdf"), width = 14, height = 4)
ggplot(t.l.filt, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity") + 
  theme_classic() + theme(text = element_text(size = 15), axis.text.x=element_text(hjust = 1)) + 
  xlab('Cluster Number') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Correlation Predicted Cell Type")) 
dev.off()

### figure S18

# keep uknown and remove the low freq cells for plotting and top label


# get ride of low frequency cells too
t.l.filt <- t.lables %>% dplyr::filter(Freq > 100)
# make bar chart
pdf(paste(output_path,"Bar.cor.pred.count.all.cells.filt2.pdf"), width = 11, height = 4)
ggplot(t.l.filt, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity") + 
  theme_classic() + theme(text = element_text(size = 15)) + 
  xlab('Cluster Number') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Correlation Predicted Cell Type")) 
dev.off()


# find the top 3 in cell types predicted in each cluster 

top.labs <- t.lables  %>% group_by(Var1)  %>% top_n(4, Freq)
top.labs

# find the top cell in each cluster
top.lab <- t.l.filt  %>% group_by(Var1)  %>% top_n(1, Freq)
top.lab


# try bar chart proportion

  

```


Explore the Seurat transfer labels

```{r}

# make a df
t.lables <- as.data.frame(table(seu$seurat_clusters, seu$predicted.id))
pr.t.lables <- as.data.frame(prop.table(table(seu$seurat_clusters, seu$predicted.id)))

# try bar chart
pdf(paste(output_path,"Bar.seu.pred.count.all.cells.pdf"), width = 10, height = 4)
ggplot(t.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity") + 
  theme_classic() + theme(text = element_text(size = 15), axis.text.x=element_text(hjust = 1)) + 
  xlab('Cluster Number') + ylab('Number of cells') +
    guides(fill = guide_legend(title = "Label Transfer Predicted Cell Type")) 
dev.off()
# find the top 3 in cell types predicted in each cluster 

top.labs <- t.lables  %>% group_by(Var1)  %>% top_n(3, Freq)
top.labs

# find the top cell in each cluster
top.lab <- t.lables  %>% group_by(Var1)  %>% top_n(1, Freq)
top.lab




```


Figure 4 related sub


```{r}
# heat map matching annotation same groups as in figure 4A

### Figure 20A

# reorder the bars to match the UMAP
levels(seu.q) <- c("Astrocytes 1", "Astrocytes 2","Radial Glia 1","Radial Glia 2",
                     "Epithelial","Endothelial","NPC","Neurons 1","Neurons 2",
                     "Oligodendrocytes","OPC","Stem cell like","Glia Lineage","Neural Lineage")


clust.colours <- c("chocolate1","orange","lightsalmon", "pink",
                   "steelblue3","deepskyblue","plum3","purple","orchid2",
                   "seagreen3","green","tomato3","burlywood3","grey90","lemonchiffon3")
 


#pdf(paste(output_path,"HMall.14072022.pdf"),width = 8, height = 5)
DoHeatmap(seu.q, features = AB, size= 6,slot = "scale.data", group.colors = clust.colours, disp.max = 2, disp.min = -1.5,
          angle = 90) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 15))
dev.off()



```

Figure 4 related subgroups

```{r}

# Figure S20

# UMAPS with smaller subgroup annotations


Idents(seu.q) <- "seurat_clusters"
cluster.ids <- c("glia-lin","neu-lin","RG1-a","Epi-RG","Neur1-a",
                 "RG3","Glia","Astro-1","NPC","RG1-2a",
                 "Astro-m","RG1-2b","Endo-a","Stem","Neur1-b",
                 "Astro-2a","NPC-eip-rg","Neur1-2a","Neur1-2b","Astro2-3",
                 "Neur-stem-opc","Astro2b","Epi-a","Astro2-3-m-RG","Epi-b",
                 "NPC-RG","Neur3-astro3","Neur3-1","Oligo","Epi-stem",
                 "neur1-RG3","RG1-neu-npc","astro-m","Endo-b","Oligo-RG-OPC",
                 "Astro")

names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels1 <- Idents(seu.q)

DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'labels1', repel = TRUE)

### change names to be able to group same cell types together


Idents(seu.q) <- "seurat_clusters"
cluster.ids <- c("GL","NL","RG-1a","Epi-1","Neur1-a",
                 "RG-3","Glia","Astro-1","NPC-1","RG-1b",
                 "Astro-m1","RG-1c","Endo-a","Stem","Neur1-b",
                 "Astro-2a","NPC-2","Neur1-2a","Neur1-2b","Astro-3",
                 "Neur-3a","Astro-2b","Epi-2","Astro-4","Epi-3",
                 "NPC-3","Neur-3b","Neur-3c","Oligo","Epi-4",
                 "Neur1-c","RG-1b","Astro-m1","Endo-b","OPC",
                 "Astro-5")

names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels2 <- Idents(seu.q)

png(paste(output_path,"UMAP.all.cell.labels2.png"), width = 1000, height = 400)
DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'labels2', repel = TRUE, label.size = 6)
dev.off()


#### figure 4 cell type labels

Idents(seu.q) <- "seurat_clusters"
cluster.ids <- c("Glia Lineage","Neural Lineage","Radial Glia 1","Epithelial","Neurons 1",
                 "Radial Glia 2","Astrocytes 1","Astrocytes 1","Neurons 1","Radial Glia 2",
                 "Astrocytes 2","Radial Glia 1","Endothelial","Stem cell like","Neurons 1",
                 "Astrocytes 1","NPC","Neurons 2","Neurons 2","Astrocytes 1",
                 "Neurons 2","Astrocytes 1","Epithelial","Astrocytes 2","Epithelial",
                 "Radial Glia 2","Astrocytes 1","Neurons 2","Oligodendrocytes","Stem cell like",
                 "NPC","NPC","Astrocytes 2","Endothelial","OPC",
                 "Astrocytes 2")



names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$cell.types <- Idents(seu.q)


# change the order of the cell types on the legend of the umap
cell.type.order <- c("Astrocytes 1", "Astrocytes 2","Radial Glia 1","Radial Glia 2",
                     "Epithelial","Endothelial","NPC","Neurons 1","Neurons 2",
                     "Oligodendrocytes","OPC","Stem cell like","Glia Lineage","Neural Lineage")
cell.type.order <- rev(cell.type.order)

# colour order to match cell type order
clust.colours <- c("chocolate1","orange","lightsalmon", "pink",
                   "steelblue3","deepskyblue","plum3","purple","orchid2",
                   "seagreen3","green","tomato3","burlywood3","grey90","lemonchiffon3")




###### more subtypes

Idents(seu) <- "seurat_clusters"
cluster.ids <- c("Glia Lineage","Neural Lineage","Radial Glia 1a","Epithelial1",
                 "Neurons 1a","Radial Glia 2a","Astrocytes 1a","Astrocytes 1b",
                 "Neurons 1b","Radial Glia 2b","Astrocytes 2a",
                 "Radial Glia 1b","Endothelial1","Stem cell like1","Neurons 1c",
                 "Astrocytes 1c","NPC1","Neurons 2a","Neurons 2b","Astrocytes 1d",
                 "Neurons 2c","Astrocytes 1c","Epithelial2","Astrocytes 2b","Epithelial3",
                 "Radial Glia 2c","Astrocytes 1d","Neurons 2d","Oligodendrocytes",
                 "Stem cell like2",
                 "NPC2","NPC3","Astrocytes 2c","Endothelial2","OPC",
                 "Astrocytes 2c")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$most.subgroups <- Idents(seu)

clust.colours <- c("cornsilk2", "chocolate1","orange","lightsalmon", "pink",
                   "aliceblue","darkslategray4", "gold2","goldenrod4","deeppink2",
                   "steelblue3","deepskyblue","plum3","purple","orchid2",
                   "seagreen3","green","tomato3","burlywood3","grey90",
                   "lemonchiffon3","firebrick3","navy","mistyrose","turquoise4",
                   "yellowgreen","tan1","tomato4","wheat2","yellow1",
                   "sienna3","palegreen3","darkred")

DimPlot(seu, cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = TRUE)

#### combine some groups

Idents(seu) <- "seurat_clusters"
cluster.ids <- c("Glia Lineage","Neural Lineage","Radial Glia 1b","Epithelial",
                 "Neurons 1a","Radial Glia 2a","Astrocytes 1a","Astrocytes 1a",
                 "Neurons 1b","Radial Glia 2b","Astrocytes 2a",
                 "Radial Glia 1a","Endothelial","Stem cell like","Neurons 1b",
                 "Astrocytes 1b","NPC","Neurons 2a","Neurons 2b","Astrocytes 1b",
                 "Neurons 2a","Astrocytes 1b","Epithelial","Astrocytes 2b","Epithelial",
                 "Radial Glia 2b","Astrocytes 1b","Neurons 2b","Oligodendrocytes",
                 "Stem cell like",
                 "NPC","NPC","Astrocytes 2a","Endothelial","OPC",
                 "Astrocytes 2a")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$subgroups <- Idents(seu)



DimPlot(seu, cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = TRUE)



sub.cells.order <- c("Astrocytes 1a","Astrocytes 1b","Astrocytes 2a","Astrocytes 2b",
                     "Radial Glia 1a","Radial Glia 1b","Radial Glia 2a","Radial Glia 2b",
                     "Epithelial","Endothelial",
                     "NPC","Neurons 1a","Neurons 1b","Neurons 2a","Neurons 2b",
                     "Oligodendrocytes","OPC","Stem cell like","Glia Lineage","Neural Lineage")
sub.cells.order <- rev(sub.cells.order)

clust.colours <- c("chocolate1","chocolate4","orange","darkorange2",
                   "lightsalmon", "pink","deeppink1","hotpink3",
                   "steelblue3","deepskyblue",
                   "plum3","purple","mediumpurple3","orchid2","darkorchid3",
                   "seagreen3","green","tomato3",
                   "burlywood3","grey90")

# save the subgroup plots

png(paste(output_path,"UMAPlabelled.all.subgroups.png"),width = 900, height = 400)
DimPlot(seu, order = sub.cells.order, cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = TRUE, label.size =6, repel = TRUE) +
  theme(legend.text = element_text(size=18), axis.title.y = element_text(size=14), 
        axis.title.x = element_text(size=18), axis.text.y = element_text(size =14),
        axis.text.x = element_text(size =18))
dev.off()




# save the annotated dataset 
saveRDS(seu, (paste(outpath,"All9MOannaote.12072022.Rds")))



```



Figure S20 - make dot plot and heatmap

```{r}


AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

# add heatmap and dotplot

#dot plot

express.by.cluster <- as.data.frame(AverageExpression(seu, features = AB, group.by = 'subgroups', slot = 'scale.data'))
express.by.cluster <- as.data.frame(scale(express.by.cluster))

col.names.RNA <- colnames(express.by.cluster)
# need to keep the same order

re.col.name <-  c( "Glia Lineage", "Neural Lineage", "Radial Glia 1b",
                   "Epithelial","Neurons 1a", "Radial Glia 2a","Astrocytes 1a",
                   "Neurons 1b","Radial Glia 2b","Astrocytes 2a","Radial Glia 1a",
                   "Endothelial","Stem cell like","Astrocytes 1b","NPC",
                   "Neurons 2a", "Neurons2b", "Astrocytes 2b", "Oligodendrocytes", 
                   "OPC" )

names(express.by.cluster) <- re.col.name
AB2 <- row.names(express.by.cluster)

express.by.cluster <- scale(express.by.cluster)
ex.data <- cbind(AB2,express.by.cluster)


# reformat
longData<- melt(ex.data)

# get the proportion of cells expressing a marker
a <- DotPlot(seu, features = AB, group.by = 'subgroups')
a$data
pct.exp <- as.data.frame(a$data) %>% select(features.plot, id, pct.exp)

# add the mean expression and the percent cells expressing together
# rename the AB and cell labels 
colnames(longData) <- c("features.plot","id","expression")

df.exp.pct <- merge(longData, pct.exp, by = c("features.plot", "id"))



# reorder 
unique(df.exp.pct$id)
cell.type.order <- c("Astrocytes 1a","Astrocytes 1b","Astrocytes 2a","Astrocytes 2b",
                     "Radial Glia 1a","Radial Glia 1b","Radial Glia 2a","Radial Glia 2b",
                     "Epithelial","Endothelial",
                     "NPC","Neurons 1a","Neurons 1b","Neurons 2a","Neurons 2b",
                     "Oligodendrocytes","OPC","Stem cell like","Glia Lineage","Neural Lineage")
cell.type.order <- rev(cell.type.order)

data <- df.exp.pct %>% mutate(Cell.type = factor(id, levels = cell.type.order))
data <- data %>% mutate(Marker = factor(features.plot, levels = AB))

# save the plot

#pdf(paste(output_path,"Dotplot.labelled.all.12072022.pdf"),width = 8, height = 5)
ggplot(data = data, aes(x = Marker, y = Cell.type, color = expression, size = pct.exp)) +
  geom_point() + theme_classic() + scale_colour_continuous(type = "gradient")
  scale_color_gradient2(low = "#154c79", mid = "#eeeee4", high = "#e28743") + ylab("Cell Phenotypes") + xlab("Antibodies") + RotatedAxis() +  theme(legend.text = element_text(size=15), axis.title.y = element_text(size=15), 
        axis.title.x = element_text(size=15), axis.text.y = element_text(size =15),
        axis.text.x = element_text(size =15, angle = 90)) 
dev.off()


# heatmap


clust.colours <- c("chocolate1","chocolate4","orange","darkorange2",
                   "lightsalmon", "pink","deeppink1","hotpink3",
                   "steelblue3","deepskyblue",
                   "plum3","purple","mediumpurple3","orchid2","darkorchid3",
                   "seagreen3","green","tomato3",
                   "burlywood3","grey90")


cell.type.order <- c("Astrocytes 1a","Astrocytes 1b","Astrocytes 2a","Astrocytes 2b",
                     "Radial Glia 1a","Radial Glia 1b","Radial Glia 2a","Radial Glia 2b",
                     "Epithelial","Endothelial",
                     "NPC","Neurons 1a","Neurons 1b","Neurons 2a","Neurons 2b",
                     "Oligodendrocytes","OPC","Stem cell like","Glia Lineage","Neural Lineage")

# get the bars in the correct/ matching order
Idents(seu) <- 'subgroups'
seu$subgroups <- factor(x = seu$subgroups, levels = cell.type.order)

Idents(seu) <- 'subgroups'
pdf(paste(output_path,"HMsubgroup.all.cells.pdf"), width = 9, height = 4)
DoHeatmap(seu, features = AB, size= 5,slot = "scale.data", group.colors = clust.colours, 
          disp.max = 2, disp.min = -1.5, angle = 90) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 12))
dev.off()

unique(seu$subgroups)


```






