---
title: "R Notebook"
output: html_notebook
---

Annotate the full 9MBO dataset
1. Use Seurat label transfer
2. Use Random Forest Trained model add labels 
3. Correlation assignment


```{r}
# load libraries
library(clusterSim) 
library(FlowSOM)
library(flowCore)
library(cluster)
library(fpc)
library(clv)
library(Seurat)
library(dplyr)
library(ggplot2)
library(clustree)
library(reshape2)

```


```{r}
# keep feature order the same as in other figures

AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

outpath <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outs/correlation/correlationscorr_celltypes.csv"

```


Seurat label transfer

```{r}
# read in the seurat object with clusters to annotate
# single cell reference mapping
# uses the function TransferData

# annotate res 0.5

seu.q <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outs/clusters/Louvain/Allcellsretros_LouvainSeuratObject200.Rds")

# the current labelled seurat object will be the reference  
#saveRDS(seu,"/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/Louvkn60DifferentCellLabels220220318.Rds")
  
#reference 
seu.r<- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/Louvkn60DifferentCellLabels220220318.Rds")


# find the anchors for label transfer
anchors <- FindTransferAnchors(reference = seu.r, query = seu.q,features = AB ,reference.reduction = "pca", dim= 1:10) 
# this step is computationally intensive and take a long time
# refdata = reference$cell_lables_wanted
predictions <- TransferData(anchorset = anchors, refdata = seu.r$labels6, dims = 1:10)

seu.q <- AddMetaData(seu.q, metadata = predictions)


  # label transfer in cluster labels

t.lables <- as.data.frame(table(seu.q$RNA_snn_res.1, seu.q$predicted.id))
pr.t.lables <- as.data.frame(prop.table(table(seu.q$RNA_snn_res.1, seu.q$predicted.id)))
t.lables$Freq <- as.double(t.lables$Freq)
# try to make a heatmap


# try bar chart
ggplot(t.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity")

# find the top 2 in cell in the group

top.labs <- t.lables  %>% group_by(Var1)  %>% top_n(3, Freq)
top.labs

top.lab <- t.lables  %>% group_by(Var1)  %>% top_n(1, Freq)
top.lab






```

See visualize

```{r}

DimPlot(seu.q, group.by = 'predicted.id', label = TRUE, repel = TRUE)
DimPlot(seu.q, group.by = 'RNA_snn_res.1', label = TRUE, repel = TRUE)

FeaturePlot(seu.q, features = "CD140a", slot = 'scale.data', min.cutoff = 0, max.cutoff = 8)

```

Correlation assignment

```{r}

cor <- read.csv("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outs/correlation/correlationscorr_celltypes.csv")

seu.q <- AddMetaData(object=seu.q, metadata=cor$cell.label, col.name = 'cor.labels')
unique(seu.q$cor.labels)

DimPlot(seu.q, group.by = 'cor.labels', label = TRUE, repel = TRUE) + theme(legend.position ="none")
DimPlot(seu.q, group.by = 'RNA_snn_res.0.5', label = TRUE, repel = TRUE) + theme(legend.position ="none")
DimPlot(seu.q, group.by = 'RNA_snn_res.1', label = TRUE, repel = TRUE) + theme(legend.position ="none")
DimPlot(seu.q, group.by = 'RNA_snn_res.1.5', label = TRUE, repel = TRUE) + theme(legend.position ="none")


# have a look at the correlations

df.filter <- cor %>% group_by(cell.label) %>% dplyr::filter(n()> 1000)

plot1 <- ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
   theme(axis.text.x=element_text(angle=90))+ xlab('Assigned cell type') + ylab('number of cell')
  print(plot1)

  # # plot
  pdf(paste(outpath,"FreqCellTypes_filt1000.pdf",sep=""),width =12, height = 6)
  plot1 <- ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
   theme(axis.text.x=element_text(angle=90))+ xlab('Assigned cell type') + ylab('number of cell')
  print(plot1)
  dev.off()

#plot the proportion of different cell types
  
  # cor labels per cluster label??

cor.lables <- as.data.frame(table(seu.q$RNA_snn_res.1, seu.q$cor.labels))
pr.cor.lables <- as.data.frame(prop.table(table(seu.q$RNA_snn_res.1, seu.q$cor.labels)))
cor.lables$Freq <- as.double(cor.lables$Freq)
# try to make a heatmap


# try bar chart
ggplot(cor.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity")

# find the top 2 in cell in the group

top.cells <- cor.lables  %>% group_by(Var1)  %>% top_n(3, Freq)
top.cells

top.cell <- cor.lables  %>% group_by(Var1)  %>% top_n(1, Freq)
top.cell





```


Random Forest labels
Model already trained and saved.

```{r}

rf.pred <- read.csv("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/RandomForest/ prediction_march15rf_allcells.csv")

seu.q <- AddMetaData(object=seu.q, metadata= as.factor(rf.pred$x), col.name = 'rf.pred')
unique(seu.q$rf.pred)

DimPlot(seu.q, group.by = 'rf.pred', label = TRUE)



rf.lables <- as.data.frame(table(seu.q$RNA_snn_res.1, seu.q$rf.pred))
rf.lables$Freq <- as.double(rf.lables$Freq) 
# try to make a heatmap


# try bar chart
ggplot(rf.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity") + theme(axis.text.x = element_text(angle = 90))

# find the top 2 in cell in the group

rf.cells <- rf.lables  %>% group_by(Var1)  %>% top_n(3, Freq)
rf.cells

rf.cell <- rf.lables  %>% group_by(Var1)  %>% top_n(1, Freq)
rf.cell




```
With the March 25 labels

```{r}
rf.pred2 <- read.csv("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/RandomForest/ prediction_march25rf_allcells.csv")

seu.q <- AddMetaData(object=seu.q, metadata= as.factor(rf.pred2$x), col.name = 'rf.pred2')
unique(seu.q$rf.pred2)

DimPlot(seu.q, group.by = 'rf.pred2', label = TRUE)


rf.lables <- as.data.frame(table(seu.q$RNA_snn_res.1, seu.q$rf.pred2))
rf.lables$Freq <- as.double(rf.lables$Freq) 
# try to make a heatmap


# try bar chart
ggplot(rf.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity") + theme(axis.text.x = element_text(angle = 90))

# find the top 2 in cell in the group

rf.cells <- rf.lables  %>% group_by(Var1)  %>% top_n(3, Freq)
rf.cells

rf.cell <- rf.lables  %>% group_by(Var1)  %>% top_n(1, Freq)
rf.cell



```


Add some labels and see how they look

kn200 res 1	Labels March25
0	Mix
1	NPC
2	Astroctye
3	RadialGlia-1
4	RadialGlia-2
5	Neuron/Epithelial/NPC/Astro
6	Astroctye
7	Astroctye
8	RadialGlia-3
9	Neuron/OPC/mix
10	Astrocytes
11	Endothelial
12	Astrocytes
13	Neurons1
14	Early neurons
15	Neurons2
16	Astrocytes
17	Neurons3 
18	NPC/Neurons
19	RadialGlia-2
20	oligodendrocyte
21	Epithelial
22	Neurons
	

```{r}


Idents(seu.q) <- "RNA_snn_res.1"
cluster.ids <- c("Mix","NPC","Astro1","RG1","RG2","N-E-NPC-Astro","Astro2","Astro3","RG3","N-OPC-Mix","Astro4","Endothelial","Astro4","Neurons1","EarlyNeurons","Neurons2","Astro5","Neurons3","NPC-Neuro","RG2","Oligo","Epithelial","Neurons4")

names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels <- Idents(seu.q)

DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'labels', repel = TRUE)

# group the save cells together

Idents(seu.q) <- "RNA_snn_res.1"
cluster.ids <- c("Mix","NPC","Astro1","RG2","RG1","Epithelial","Astro2","Astro1","RG2","Neuron-OPC-Mix","Astro3","Endothelial","Astro4","Neurons","EarlyNeurons","Neurons","Astro5","Neurons","NPC2","RG3","Oligo","Epithelial","Neurons")


names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels2 <- Idents(seu.q)

DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'labels2', repel = TRUE)

Idents(seu.q) <- "RNA_snn_res.1"
cluster.ids <- c("Neuro-Astro-Mix","NPC","Astro1","RG2","RG1","Epithelial","Astro1","Astro1","RG2","Neuron-OPC-Mix","Astro3","Endothelial","Astro2","Neurons","EarlyNeurons","Neurons","Astro3","Neurons","NPC","RG1","Oligo","Epithelial","Neurons")


names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels3 <- Idents(seu.q)

DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'labels3', repel = TRUE)

# look at the heat map

# heat map takes super long to run with so much data
DoHeatmap(seu.q, group.by = 'labels3', features = AB)


# grouping for gating maintaining subgroups
# for gating I will subset out the mix populations

Idents(seu.q) <- "RNA_snn_res.1"
cluster.ids <- c("Mix","Neurons-CD24+","Astro-CD44+","RG-CD184+","RG-CD133+","Mix","Astro-Glast+","Astro-CD44+","RG-CD184+","Mix","Astro-CD44+","Endothelial","Astro-CD44+","Neurons-CD56+","Neurons-CD24+CD56+","Neurons-CD56+","Astro-CD44+","Neurons-CD56+","Neurons-CD24+","RG-CD133+","Oligo","Epithelial","Neurons-CD24+CD56+")


names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels.groups <- Idents(seu.q)

DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'labels.groups', repel = TRUE)


# grouping for gating maintaining subgroups
# remain for even more basic groups 
# I'm leaving CD56 neurons as one thing and the CD15/CD24 neurons together
# joining the two astro together
# joining the two RG together


Idents(seu.q) <- "RNA_snn_res.1"
cluster.ids <- c("Mix","Neurons1","Astro","RG","RG","Mix","Astro","Astro","RG","Mix","Astro","Endothelial","Astro","Neurons2","Neurons1","Neurons2","Astro","Neurons2","Neurons1","RG","Oligo","Epithelial","Neurons1")


names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels.main.groups <- Idents(seu.q)

DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'labels.main.groups', repel = TRUE)


# now try with more groups than before
# split astrocytes into more groups 
# astro1 and 3 together 
# astro2 alone
# astro4 and 5 together

cluster.ids <- c("Mix","Neurons-CD24+","Astro-CD44+","RG-CD184+","RG-CD133+","Mix","Astro-Glast+","Astro-CD44+","RG-CD184+","Mix","Astro-CD44+","Endothelial","Astro-CD44+","Neurons-CD56+","Neurons-CD24+CD56+","Neurons-CD56+","Astro-CD44+","Neurons-CD56+","Neurons-CD24+","RG-CD133+","Oligo","Epithelial","Neurons-CD24+CD56+")

cluster.ids <- c("Mix","NPC","Astro1","RG1","RG2","N-E-NPC-Astro","Astro2","Astro3","RG3","N-OPC-Mix","Astro4","Endothelial","Astro4","Neurons1","EarlyNeurons","Neurons2","Astro5","Neurons3","NPC-Neuro","RG2","Oligo","Epithelial","Neurons4")


Idents(seu.q) <- "RNA_snn_res.1"  #set to the clusters you want to rename - it's probably 'seurat.clusters' in your oject
cluster.ids <- c("Mix","Neurons1","Astro1","RG1","RG2","Mix","Astro2","Astro1","RG1","Mix","Astro3","Endothelial","Astro3","Neurons2","Neurons3","Neurons2","Astro3","Neurons2","Neurons1","RG2","Oligo","Epithelial","Neurons3")
# vector with the new names - you need this vector from me

names(cluster.ids) <- levels(seu.q)    # where seu.q is your seurat object
seu.q <- RenameIdents(seu.q, cluster.ids) 
seu.q$cell.types <- Idents(seu.q)   # cell.types is a new slot in metadata you can call this whatever you want

DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'cell.types', repel = TRUE)



saveRDS(seu.q, "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outs/clusters/AllcellLablesMarch25.Rds")



```


Quick check proportion of cell types across samples

```{r}

table(Idents(seu.q), seu.q$Batch)

# proportion across channels
prp.t <- as.data.frame(prop.table(table(Idents(seu.q))))



```






Proportions of cell types

```{r}
sample.lables <- as.data.frame(table(seu.q$Batch, seu.q$labels))

sample.lables$Freq <- as.double(sample.lables$Freq)
# try to make a heatmap


# try bar chart
ggplot(sample.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity")+ theme_classic() + theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))  + xlab('Sample') + ylab('Cell count') + RotatedAxis()





# bar chart of with percent 
ggplot(sample.lables, aes(x = Var1,y=Freq ,fill = Var2)) + geom_bar(position= "fill", stat = "identity") + scale_y_continuous(labels = scales::percent_format()) + theme_classic() + theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))  + xlab('Sample') + ylab('Percent of Cell type') + RotatedAxis()



```

Calculate if the proportions are different

```{r}
# https://github.com/rpolicastro/scProportionTest

# now made into a package

#devtools::install_github("rpolicastro/scProportionTest")
library("scProportionTest")

prop_test <- sc_utils(seu.r)

unique(seu.r$Batch)

prop_test <- permutation_test(
	prop_test, cluster_identity = "labels6",
	sample_1 = "3450_0306", sample_2 = "3450_0317A",
	sample_identity = "Batch"
)

permutation_plot(prop_test)

```



```{r}

# I want to compare proportions across groups - like a 2 way anova
# need to setup the data

 # sample.lables this is the cell counts per cluster from above


wide.data <- reshape(data= sample.lables,idvar = Var1 )






```




See ridge plots for grouping clusters together

```{r}
RidgePlot(seu.q, group.by = 'labels', features = "CD24", log= TRUE)
RidgePlot(seu.q, group.by = 'labels', features = "CD56", log= TRUE)
RidgePlot(seu.q, group.by = 'labels', features = "CD29", log= TRUE)
RidgePlot(seu.q, group.by = 'labels', features = "CD15", log= TRUE)

```




