---
title: "Cluster Parameters Louvain Seurat"
output: html_notebook
---

Read in preprocessed csv created from fsc to flowset files 

Create Seurat object
Normalize or not and scale 
Create UMAP - UMAP conditions already compared - used final conditions for UMAP - can change later
Keep UMAP consistant across clustering method for comparative visualization

run standard processing (make PCA) and cluster
Later compare clustering with directly from AB no PCA 

```{r}
# set up environment
library(FlowSOM)
library(flowCore)
library(cluster)
library(fpc)
library(clv)
library(Seurat)
library(dplyr)
library(ggplot2)
rm(list=ls())

```
Take in the data from the preprocessing output and create the expression matrix

```{r}
# input pathway
input_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/prepro_outsflowset.csv"

# read in data
df <- read.csv(input_path)
# print info to log 
print(dim(df)) # this is specific df has 73578 cells
# the preprocessing output csv needs to be cleaned - it contains live dead, FSC, SSC and the sample column
print(colnames(df))
# create a df with just the expression 
# need a way to automate this selection 
# I only want the expression values
df2 <- df %>% select(c("AQP4", "CD24", "CD44","CD184","CD15","HepaCAM","CD29","CD56", "O4","CD140a","CD133","GLAST","CD71"))
# the order of the DF is set by the order the colunms are written above

m <- as.matrix(df2) 

range(df2)




```



Make the starting seurat object

```{r}
# add expression values to the seurat object
# transpose the df
tm <- t(df2)
rownames(tm) <- colnames(df2)
colnames(tm) <- rownames(df2)
seu <- CreateSeuratObject(tm) # create a seurat object 

# add the meta data back in for sample groups
seu <- AddMetaData(object=seu, metadata=df$Batch, col.name = 'Batch')
# this doesn't work for making levels

# create the vector for the antibodies names for feature plotting later
AB <- colnames(df2)

# add to scale data slot
seu <- ScaleData(seu)

# check the data
DoHeatmap(seu, group.by = "Batch", features = AB) 




```


```{r}
# PCA 
# find neighbours needs PCA input 
seu <- RunPCA(seu, features = AB, npcs = 25)

# check out the PCA

VizDimLoadings(seu, dims = 1:2, reduction = "pca")
DimPlot(seu, reduction = "pca", group.by = "Batch")
ElbowPlot(seu)
# there are only 13 inputs use dims = 13


```



Loop to cluster at different kn and resolultions, make plots and calculate statistics

```{r}

# change to proper pathway
output_path = "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/test/flowset/"

kn = c(25,50,100,150,200,250,300)
resolutions = c(0.01,0.1,0.25,0.5,0.75,1.0,1.5,2)

# save a data object for each kn - will only keep temporarily
# the clusters will write over with each new kn

kn = 25
resolutions = 0.5

ch_li <-list()

for (i in kn){
  seu <- FindNeighbors(seu, dims = 1:12, k = i)
  seu <- RunUMAP(seu, dims = 1:12, n.neighbors = i)
  # save feature plots of this UMAP
  # just for testing print
  print(FeaturePlot(seu, features = AB,slot = 'scale.data',min.cutoff = 'q1', max.cutoff ='99',label.size = 1))
  # file name
  UMAP_name = paste("UMAPfeatures_kn",i,".pdf",sep="")
  print(UMAP_name) #testing 
  # save feature plots UMAP
  pdf(paste(output_path,UMAP_name,sep=""),width =20, height = 10)
    print(FeaturePlot(seu, features = AB,slot = 'scale.data',min.cutoff = 'q1', max.cutoff ='99',label.size = 1))
  dev.off()
  
  for (j in resolutions) {
    seu <- FindClusters(seu, resolution = j)
    
    # calculate the statistics
    # silouette
    # CHI
    ch = calinhara(m,seu@meta.data$seurat_cluster, )
    ch = calinhara(m,seu@meta.data$seurat_clusters,length((unique(seu@meta.data$seurat_clusters))))
    ch_li[j] <- ch
    # Davies
    
    # send stats to stats_list (or df or whatever works)
    
    # make plots
    # UMAP
    UMAP_name = paste("UMAPclusters_kn",i,"_res_",j,".pdf",sep="")
  print(UMAP_name) #testing 
   pdf(paste(output_path,UMAP_name,sep=""),width =8, height = 5)
  # save UMAP grouped
    print(DimPlot(seu,reduction = "umap", repel = TRUE, label = TRUE)) # will automatically group by active ident
  dev.off()
    # heatmap
  print(DimPlot(seu,reduction = "umap", repel = TRUE, label = TRUE))

    # save stats for each resolution
    # write.csv(stats_list, paste(output_path,list_name,sep=""))
  }
  # save seurat object
  seu_name = paste("SeuratObject",i,"rds",sep="")
  saveRDS(seu, paste(output_path,seu_name,sep=""))
   # make clustree plot
  # save all stats outputs for each kn
}

```

