---
title: "Cluster Parameters Louvain Seurat"
output: html_notebook
---

Read in preprocessed csv created from fsc to flowset files 

Create Seurat object
Normalize or not and scale 
Create UMAP - UMAP conditions already compared - used final conditions for UMAP - can change later
Keep UMAP consistant across clustering method for comparative visualization

run standard processing (make PCA) and cluster
Later compare clustering with directly from AB no PCA 

```{r}
# set up environment

```



```{r}
# read in data




```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

