---
title: "Test Rphenograph clustering Jan 25 2022"
output: html_notebook
---


Load libraries - see what is needed as I go

```{r}
library(FlowSOM)
#library(flowCore)
#library(cluster)
#library(fpc)
#library(clv)
library(Seurat)
library(dplyr)
library(Rphenograph)
rm(list=ls())


```

Code from website to test installation and to figure out how the inputs work.

```{r}
iris_unique <- unique(iris) # Remove duplicates from original dataframe
data <- as.matrix(iris_unique[,1:4]) # change to matrix
Rphenograph_out <- Rphenograph(data, k = 45) # run function to get clusters
# creates an object class List with two lists
# [[1]] igraph has 10 lists
# [[2]] communities list with 3 lists: membership, memberships, modularity

modularity(Rphenograph_out[[2]]) # returns a single value must be a modularity calculation 
membership(Rphenograph_out[[2]]) # I think this is which cluster a data point belongs to but what does the function do? This lets us see 3 groups with index numbers 1-149

iris_unique$phenograph_cluster <- factor(membership(Rphenograph_out[[2]])) # add cluster IDs into original DF

ggplot(iris_unique, aes(x=Sepal.Length, y=Sepal.Width, col=Species, shape=phenograph_cluster)) + geom_point(size = 3)+theme_bw()


```

The phenograph example works - Apply to the flow data

```{r}
# read in the data and create an expression matrix 

#input file path, change if needed
fileName <-"/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/prepro_outsflowset.csv"

# note: current matrix sample ID have cell index # attached. 

df <- read.csv(fileName)
head(df)
print(dim(df)) # this is specific df has 73578 cells
# the preprocessing output csv needs to be cleaned - it contains live dead, FSC, SSC and the sample column
df2 <- df %>% select(-c("Live.Dead",FSC,SSC,X,Batch,cell))

m <- as.matrix(df2) # make a matrix as input to phenograph


```
```{r}

# Rphenograph seems to be just one function and we can adjust the K for number of neighbours
Rphenograph_out_flow <- Rphenograph(m, k = 50)


modularity(Rphenograph_out_flow[[2]])
membership(Rphenograph_out_flow[[2]])

# add cluster ID back into original df
df$phenograph_cluster <- factor(membership(Rphenograph_out_flow[[2]]))

# how many clusters are there?
unique(df$phenograph_cluster)
# 30 levels !!!! Way to many - because the k is low. 

#ggplot(iris_unique, aes(x=Sepal.Length, y=Sepal.Width, col=Species, shape=phenograph_cluster)) + geom_point(size = 3)+theme_bw()

```

Try with a higher k

```{r}
# Rphenograph seems to be just one function and we can adjust the K for number of neighbours
Rphenograph_out_flow <- Rphenograph(m, k = 271)


modularity(Rphenograph_out_flow[[2]])
membership(Rphenograph_out_flow[[2]])

# add cluster ID back into original df
df$phenograph_clusterk271 <- factor(membership(Rphenograph_out_flow[[2]]))


# 30 levels !!!! Way to many - because the k is low. 

```


```{r}

# how many clusters are there?
unique(df$phenograph_clusterk271)
# there are still 22 clusters 
# computation time is much longer with higher K

```

```{r}

# I'll have a look at the clustering quickly using two AB


ggplot(df, aes(x=CD44, y=CD71, col = phenograph_clusterk271)) + geom_point(size = 1)+theme_bw()

# can't tell much
# I'll try and add the phenograph cluster indexs into the seurat object


```


Save the df with the phenograph cluster indexes to save time.

```{r}
write.csv(df,"/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/prepro_outsflowsetdf+phenographk50k271.csv")


```

```{r}
# seurat object made from the save input as the phenograph clustering
seu <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/SeuratfromFlowsom.rds")

# read in the df with the phenograph clustering
df <- read.csv("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/prepro_outsflowsetdf+phenographk50k271.csv")

# add the phenograph cluster indexes
seu <- AddMetaData(object=seu, metadata=df$phenograph_clusterk271, col.name = 'Phenograph.k.271')

```

See some plots

```{r}
DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "Phenograph.k.271")


```

```{r}
# make a list of AB from the input df - can only use df if filter out all the exta parts
print(colnames(df2))
allAB <- colnames(df2)

DoHeatmap(seu, features = allAB, group.by = "Phenograph.k.271")

```
```{r}

DotPlot(seu, features = allAB, group.by = "Phenograph.k.271", cols = c("blue","red"))
```

```{r}
DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "flowSOM.k.8")

DoHeatmap(seu, features = allAB, group.by = "flowSOM.k.8")


```
Save seurat object with Phenograph and FlowSOM.k.8 clusters

```{r}
saveRDS(seu,"/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/SeuratfromFlowsomPheno.rds" )

```



Try to optimize UMAP to get better separation of groups


```{r}

# UMAP parameter
# n.neigbors or n_neighbors : number neighboring points for local approximation of manifult - larger give more global structur and looses detials

# min.dist: detrimines how tightly embedded points are range from 0.001 - 0.5 : large values ensure embeeded points are evenly dist, small values otimse accuracy in local structures
# metric
# a 
# b
# dims
# learning rate 

# assay.use = "RNA" 
# gene.use = allAB - this will run on the features instead of the PC

# test n.neigbours 
num.neighbors = c(30,50,100,150,200,250)

for (ng in num.neighbors){
  seu <- RunUMAP(seu, dims = 1:10, n.neighbors = ng)
  print(DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "Phenograph.k.271"))
}

#n.neighbors doesn't make a big difference, I think higher numbers are better visually

# test distances
dist.opt = c(0.001,0.01,0.1,0.5)

for (ds in dist.opt){
  seu <- RunUMAP(seu, dims = 1:10, n.neighbors = 250, min.dist = ds)
  print(DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "Phenograph.k.271"))
}

# low value for min.dist  push things apart more still a bit odd

# try using the AB to cluster
# this works better than the PCA

seu <- RunUMAP(seu, dims = NULL, n.neighbors = 250, min.dist = 0.005, features = allAB, slot = 'scale.data')
DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "Phenograph.k.271")
DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "RNA_snn_res.1")

# try changing spread


spread.opt <- c(0.1,0.5,0.75,1,5)
a.opt <- c(830,5.07,2.51,1.58,0.14)
b.opt <- c(1.93,1.0,0.93,0.90,0.81)

for (i in 1:5){
  seu <- RunUMAP(seu, dims = NULL, n.neighbors = 250, min.dist = 0.1 ,features = allAB, slot = 'scale.data', spread = spread.opt[i], a = a.opt[i], b = b.opt[i])
  print(DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "Phenograph.k.271"))
  print(DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "RNA_snn_res.1"))
}

# again the seurat clustering looks better with larger spread
# spread = 5 and min.dist = 0.1


```

```{r}

resolutions = c("RNA_snn_res.0.1","RNA_snn_res.0.25","RNA_snn_res.0.5","RNA_snn_res.0.75","RNA_snn_res.1")

for (res in resolutions){
 print(DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = res))
 
}



```

```{r}
# run at a higher spread
seu2 <- seu

seu2 <- RunUMAP(seu2, dims = NULL, n.neighbors = 250, min.dist = 0.01 ,features = allAB, slot = 'scale.data', spread =10, a = 0.05, b = 0.8)


```

```{r}

resolutions = c("RNA_snn_res.0.1","RNA_snn_res.0.25","RNA_snn_res.0.5","RNA_snn_res.0.75","RNA_snn_res.1")

for (res in resolutions){
 print(DimPlot(seu2, reduction = "umap", repel = TRUE, label = TRUE, group.by = res))
 
}


DimPlot(seu2, reduction = "umap", repel = TRUE, label = TRUE, group.by = "Phenograph.k.271")

# spread = 10 must be too high

```

```{r}


spread.opt <- c(2,3,5)
a.opt <- c(0.54,0.3,0.14)
b.opt <- c(0.84,0.82,0.81)

for (i in 1:5){
  seu <- RunUMAP(seu, dims = NULL, n.neighbors = 250, min.dist = 0.001 ,features = allAB, slot = 'scale.data', spread = spread.opt[i], a = a.opt[i], b = b.opt[i])
  print(DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "Phenograph.k.271"))
  print(DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "RNA_snn_res.1"))
}

# spread of 3 is best




```

```{r}

# spread 3 

dist.opt = c(0.001,0.005,0.01,0.05)

for (ds in dist.opt){
  seu <- RunUMAP(seu, dims = NULL, n.neighbors = 250, min.dist = ds ,features = allAB, slot = 'scale.data', spread = 3, a = 0.3, b = 0.82)
  print(DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "Phenograph.k.271"))
  print(DimPlot(seu, reduction = "umap", repel = TRUE, label = TRUE, group.by = "RNA_snn_res.1"))
}


```




