---
title: "R Notebook"
output: html_notebook
---



```{r}
#install 

# dependencies
install.packages(c("sp", "polyclip", "rgeos"))



# package from github

library(devtools)
install_github(repo = "ebecht/hypergate")


library("Seurat")
library("ggplot2")
library("dplyr")


```

Hypergate takes in 
1. an expression matrix
2. a vector of events to attempt to gate on - there are different ways to get these

interactive gating - try later
Clustering - this is what I want to use

hypergate is the function to optomize gating strategies

xp = a numberic matrix encoding expression 
gate_vector a vector with a few unique values --- this should be the cluster labels
level specifies what value of gate vector togate upon 

```{r}
library("hypergate")

input.xm = as.matrix(GetAssayData(seu.q, slot = 'scale.data'))
xm.t <- t(input.xm)
cluster.labels <- as.vector(seu.q$labels)


hg_output <- hypergate(xp = xm.t, gate_vector = cluster.labels, level = 'Astro1')

# writing level = 1 resulted empty output
# putting 'Astro1' is running but taking a very long time
# I believe I will have to down sample the Seurat object.  I thing I might want to subset the AIW002 first




```


I'll need to downsample the seurat object

```{r}
Idents(seu.q) <- 'Batch'

unique(seu$Batch)
# subset the AIW because we will use AIW in the FACS experiment

seu <- subset(seu.q, idents = c("AIW002_0306","AIW002_0317A", "AIW002_0317B"))

# downsampe to about 2000 cells so that hypergate can run
# there is an option to down sample per identity class

Idents(seu) <- 'labels'
seu.down <- subset(x= seu, downsample = 200)

DimPlot(seu.down)




```

```{r}
# try hypergate again on one group 
input.xm = as.matrix(GetAssayData(seu.down, slot = 'scale.data'))
xm.t <- t(input.xm)
cluster.labels <- as.vector(seu.down$labels)


hg_output <- hypergate(xp = xm.t, gate_vector = cluster.labels, level = 'Astro1')



```






Try checking the results

```{r}

gating_predicted = subset_matrix_hg(hg_output,xm.t)


table(ifelse(gating_predicted, "Gated-in", "Gated-out"), ifelse(cluster.labels == 
    'Astro1', "Events of interest", "Others"))

# as we started with 200 astro1 we captured a good amount


```

Now we can see if each cell past parameters for the threshold for each AB

```{r}

bm = boolmat(gate = hg_output, xp = xm.t)
head(bm)



```

Plots some gates


```{r}

plot_gating_strategy(gate = hg_output, xp = xm.t, gate_vector = cluster.labels, 
    level = 'Astro1', highlight = "firebrick3")

```

Channel contributions

```{r}

contributions = channels_contributions(gate = hg_output, xp = xm.t, gate_vector = cluster.labels, 
    level = 'Astro1', beta = 1)
barplot(contributions, horiz = TRUE, 
    xlab = "F1-score deterioration when the parameter is ignored")
print(contributions)

```

We could reoptimize using only the top 4 
But right now I'll leave this


```{r}
hgate_pheno(hg_output)

```

```{r}
hgate_rule(hg_output)
```

```{r}
hgate_info(hg_output)  # make a helpful table 
```

I need to define the groups I want to gate first and then use this to gate them. 

```{r}
AB.order <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")


RidgePlot(seu.down, features = "CD44", log = TRUE )
RidgePlot(seu.down, features = "CD24", log = TRUE )


```

Regroup the similar for gating and remove mix cells that might confuse

```{r}


Idents(seu.q) <- 'labels.groups'
seu <- subset(x= seu.q, idents = c("Neurons-CD24+","Astro-CD44+","RG-CD184+","RG-CD133+","Astro-Glast+","Astro-CD44+","RG-CD184+","Astro-CD44+","Endothelial","Astro-CD44+","Neurons-CD56+","Neurons-CD24+CD56+","Neurons-CD56+","Astro-CD44+","Neurons-CD56+","Neurons-CD24+","RG-CD133+","Oligo","Epithelial","Neurons-CD24+CD56+")) 

DimPlot(seu, label = TRUE, repel = TRUE)

Idents(seu) <- 'Batch'
seu <- subset(seu.q, idents = c("AIW002_0306","AIW002_0317A", "AIW002_0317B"))


# downsampe to about 2000 cells so that hypergate can run
# there is an option to down sample per identity class

Idents(seu) <- 'labels.groups'
seu.down <- subset(x= seu, downsample = 400)

DimPlot(seu.down)

table(seu.down$labels.groups)




```
Hypergate on one group

```{r}
# try hypergate again on one group 
input.xm = as.matrix(GetAssayData(seu.down, slot = 'scale.data'))
xm.t <- t(input.xm)
cluster.labels <- as.vector(seu.down$labels.groups)

#try all at once

cell.types <- c("Neurons-CD24+", "Astro-CD44+", "RG-CD184+", "RG-CD133+", "Astro-Glast+","Endothelial","Neurons-CD56+","Neurons-CD24+CD56+","Oligo","Epithelial")

for (cell in cell.types) {
  hg_output <- hypergate(xp = xm.t, gate_vector = cluster.labels, level = cell)
  gating_predicted = subset_matrix_hg(hg_output,xm.t)
  conf.table <- table(ifelse(gating_predicted, "Gated-in", "Gated-out"), ifelse(cluster.labels ==  cell, cell, "Others"))
  print(cell)
  print(conf.table)
  plot_gating_strategy(gate = hg_output, xp = xm.t, gate_vector = cluster.labels, 
    level = cell, highlight = "firebrick3")
  gate.table <- hgate_info(hg_output) 
  print(gate.table)
}


# this prints the threshold for each gate
hgate_rule(hg_output)

# clearer output - should try make a summary of the accuray 

for (cell in cell.types) {
  hg_output <- hypergate(xp = xm.t, gate_vector = cluster.labels, level = cell)
  gating_predicted = subset_matrix_hg(hg_output,xm.t)
  conf.table <- table(ifelse(gating_predicted, "Gated-in", "Gated-out"), ifelse(cluster.labels ==  cell, cell, "Others"))
  print(cell)
  print(conf.table)
  print(hgate_rule(hg_output))
}






```

Check results

```{r}
gating_predicted = subset_matrix_hg(hg_output,xm.t)


table(ifelse(gating_predicted, "Gated-in", "Gated-out"), ifelse(cluster.labels == 
    'Astro-CD44+', "Events of interest", "Others"))

#not great accuracy

bm = boolmat(gate = hg_output, xp = xm.t)
head(bm)

plot_gating_strategy(gate = hg_output, xp = xm.t, gate_vector = cluster.labels, 
    level = 'Astro-CD44+', highlight = "firebrick3")

contributions = channels_contributions(gate = hg_output, xp = xm.t, gate_vector = cluster.labels, 
    level = 'Astro-CD44+', beta = 1)
barplot(contributions, horiz = TRUE, 
    xlab = "F1-score deterioration when the parameter is ignored")
print(contributions)

hgate_pheno(hg_output)

hgate_rule(hg_output)

gate.table <- hgate_info(hg_output)  # make a helpful table 


```



Try with the input data
```{r}
# try hypergate again on one group 
input.xm = as.matrix(GetAssayData(seu.down, slot = 'data'))
xm.t <- t(input.xm)
cluster.labels <- as.vector(seu.down$labels.groups)

#try all at once

cell.types <- c("Neurons-CD24+", "Astro-CD44+", "RG-CD184+", "RG-CD133+", "Astro-Glast+","Endothelial","Neurons-CD56+","Neurons-CD24+CD56+","Oligo","Epithelial")


# clearer output - should try make a summary of the accuray 

for (cell in cell.types) {
  hg_output <- hypergate(xp = xm.t, gate_vector = cluster.labels, level = cell)
  gating_predicted = subset_matrix_hg(hg_output,xm.t)
  conf.table <- table(ifelse(gating_predicted, "Gated-in", "Gated-out"), ifelse(cluster.labels ==  cell, cell, "Others"))
  print(cell)
  print(conf.table)
  print(hgate_rule(hg_output))
  # save files with the plots 
}



```

```{r}

# visualize
DimPlot(seu.q)

```
Try keeping the mix population

```{r}


Idents(seu.q) <- 'Batch'
seu <- subset(seu.q, idents = c("AIW002_0306","AIW002_0317A", "AIW002_0317B"))


# downsampe to about 2000 cells so that hypergate can run
# there is an option to down sample per identity class

Idents(seu) <- 'labels.groups'
seu.down <- subset(x= seu, downsample = 400)

DimPlot(seu.down)

table(seu.down$labels.groups)


```


```{r}
# keeping mix in 

# try hypergate again on one group 
input.xm = as.matrix(GetAssayData(seu.down, slot = 'data'))
xm.t <- t(input.xm)
cluster.labels <- as.vector(seu.down$labels.groups)

#try all at once

cell.types <- c("Neurons-CD24+", "Astro-CD44+", "RG-CD184+", "RG-CD133+", "Astro-Glast+","Endothelial","Neurons-CD56+","Neurons-CD24+CD56+","Oligo","Epithelial","Mix")


# clearer output - should try make a summary of the accuray 

for (cell in cell.types) {
  hg_output <- hypergate(xp = xm.t, gate_vector = cluster.labels, level = cell)
  gating_predicted = subset_matrix_hg(hg_output,xm.t)
  conf.table <- table(ifelse(gating_predicted, "Gated-in", "Gated-out"), ifelse(cluster.labels ==  cell, cell, "Others"))
  print(cell)
  print(conf.table)
  print(hgate_rule(hg_output))
}


# having the mix cells in make no difference


```


Try with main groups Neurons1, Neurons2, Astro, RG

```{r}

Idents(seu.q) <- 'Batch'
seu <- subset(seu.q, idents = c("AIW002_0306","AIW002_0317A", "AIW002_0317B"))


# downsampe to about 2000 cells so that hypergate can run
# there is an option to down sample per identity class

# not enough oligo in only AIW

Idents(seu.q) <- 'labels.main.groups'
seu.down <- subset(x= seu.q, downsample = 500)

DimPlot(seu.down)

table(seu.down$labels.main.groups)




# try hypergate again on one group 
input.xm = as.matrix(GetAssayData(seu.down, slot = 'data'))
xm.t <- t(input.xm)
cluster.labels <- as.vector(seu.down$labels.main.groups)

#try all at once

cell.types <- c("Neurons1","Neurons2","Astro","RG","Endothelial","Oligo","Epithelial")



# clearer output - should try make a summary of the accuray 

for (cell in cell.types) {
  hg_output <- hypergate(xp = xm.t, gate_vector = cluster.labels, level = cell)
  gating_predicted = subset_matrix_hg(hg_output,xm.t)
  conf.table <- table(ifelse(gating_predicted, "Gated-in", "Gated-out"), ifelse(cluster.labels ==  cell, cell, "Others"))
  print(cell)
  print(conf.table)
  print(hgate_rule(hg_output))
}


# having the mix cells in makes no difference




```


Repeat with more populations

```{r}

# define more groups
#cluster.ids <- c("Mix","Neurons1","Astro1","RG1","RG2","Mix","Astro3","Astro1","RG1","Mix","Astro3","Endothelial","Astro3","Neurons2","Neurons3","Neurons2","Astro3","Neurons2","Neurons1","RG2","Oligo","Epithelial","Neurons3")

DimPlot(seu.q, label= TRUE)

Idents(seu.q) <- 'cell.types'
seu.down <- subset(x= seu.q, downsample = 800)

DimPlot(seu.down)

table(seu.down$cell.types)




# try hypergate again on one group 
input.xm = as.matrix(GetAssayData(seu.down, slot = 'data'))
xm.t <- t(input.xm)
cluster.labels <- as.vector(seu.down$cell.types)

#try all at once

cell.types <- c("Neurons1","Neurons2","Neurons3","Astro1","Astro2","Astro3","RG1","RG2","Endothelial","Oligo","Epithelial","Mix")
# 11 possible populations



# clearer output - should try make a summary of the accuray 

for (cell in cell.types) {
  hg_output <- hypergate(xp = xm.t, gate_vector = cluster.labels, level = cell)
  gating_predicted = subset_matrix_hg(hg_output,xm.t)
  conf.table <- table(ifelse(gating_predicted, "Gated-in", "Gated-out"), ifelse(cluster.labels ==  cell, cell, "Others"))
  print(cell)
  print(conf.table)
  print(hgate_rule(hg_output))
}





```

Load in libraries
```{r}
require("reshape2") #visualization
require("flowStats") #Alignment functions
require("scales") #scale colour intensity for visualization
require("dplyr")
library("hypergate")
library("Seurat")

```

Read in the data for gating

```{r}
seu.r <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outs/clusters/AllcellLablesMarch25.Rds")


```




I"m going to run the hypergates again - maybe multiple times to see how consitatant to output will be

```{r}


Idents(seu.r) <- 'cell.types'
i = 300
set.seed(i)
seu.down <- subset(x= seu.r, downsample = 200)


# try hypergate again on one group 
input.xm = as.matrix(GetAssayData(seu.down, slot = 'data'))
xm.t <- t(input.xm)
cluster.labels <- as.vector(seu.down$cell.types)

#try all at once

cell.types <- c("Neurons1","Neurons2","Neurons3","Astro1","Astro2","Astro3","RG1","RG2","Endothelial","Oligo","Epithelial","Mix")
# 11 possible populations



# clearer output - should try make a summary of the accuray 

for (cell in cell.types) {
  hg_output <- hypergate(xp = xm.t, gate_vector = cluster.labels, level = cell)
  gating_predicted = subset_matrix_hg(hg_output,xm.t)
  conf.table <- table(ifelse(gating_predicted, "Gated-in", "Gated-out"), ifelse(cluster.labels ==  cell, cell, "Others"))
  print(cell)
  print(conf.table)
  print(hgate_rule(hg_output))
}




```



Try using the smaller labeled object - will still need to subset
Also make groupings of cells

```{r}

seu <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/Louvkn60DifferentCellLabels220220318.Rds")

saveRDS(seu,"/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/Louvkn60DifferentCellLabels220220318.Rds")

AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

#DimPlot(seu)
#DoHeatmap(seu, features = AB)




```


Run with different cutoff of cells in down sampling


```{r}

unique(seu$labels6)

# get the cell type labels


```

Relabel cell type groups

```{r}
Idents(seu) <- "labels6"
cluster.ids <- c("Mix","RG1","Mix","Neurons1","Neurons2","Astro1","Astro2","RG2","Neuron3","RG3","Endothelial","Oligo","POC","Epithelial")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$gating.all <- Idents(seu)

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'gating.all', repel = TRUE)


# now fewer groups
Idents(seu) <- "labels6"
cluster.ids <- c("Mix","RG1","Mix","Neurons1","Neurons2","Astro1","Astro2","RG2","Neurons3","RG1","Endothelial","Oligo","OPC","Epithelial")
# merge cells that won't be used for gating
cluster.ids <- c("other","RG1","other","Neurons1","Neurons2","Astro1","Astro2","RG2","Neurons3","RG1","other","other","other","other")
# adjust based on hypergate results
cluster.ids <- c("other","glia","other","Neurons1","Neurons2","glia","Astro","RG2","Neurons3","glia","other","other","other","other")


names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$gating.main <- Idents(seu)

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'gating.main', repel = TRUE)

# now fewer groups
Idents(seu) <- "labels6"
cluster.ids <- c("other","RG","other","Neurons1","Neurons2","Astro","Astro","RG","Neurons1","RG","other","other","other","other")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$gating.4 <- Idents(seu)

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'gating.4', repel = TRUE)





```





```{r}
Idents(seu) <- 'gating.main'
i = 300
set.seed(i)
seu.down <- subset(x= seu, downsample = 500)


# try hypergate again on one group 
input.xm = as.matrix(GetAssayData(seu.down, slot = 'data'))
xm.t <- t(input.xm)
cluster.labels <- as.vector(seu.down$gating.main)

#try all at once
gating.main.celltypes <- c("RG1","RG2","Neurons1","Neurons2","Neurons3","Astro1","Astro2","Endothelial","Oligo","OPC","Epithelial")

# for gating.4
cell.types <- c("RG","Neurons1","Neurons2","Astro")
# gating more cell populations to try and get better accuracy.  

cell.types <- c("glia","RG2","Neurons1","Neurons2","Neurons3","Astro")
# note that RG1 and astro1 overlap and RG2 and astro 2 overlap a little bit
# RG1 and astro 1 are highy overlapping - merge together

# clearer output - should try make a summary of the accuray 

for (cell in cell.types) {
  hg_output <- hypergate(xp = xm.t, gate_vector = cluster.labels, level = cell)
  gating_predicted = subset_matrix_hg(hg_output,xm.t)
  conf.table <- table(ifelse(gating_predicted, "Gated-in", "Gated-out"), ifelse(cluster.labels ==  cell, cell, "Others"))
  print(cell)
  print(conf.table)
  print(hgate_rule(hg_output))
  # visualize the predicted cells 
  predicted.cell.type = paste(cell,"gated",sep=".")
  seu.down <- AddMetaData(object=seu.down, metadata=gating_predicted, col.name = predicted.cell.type )
  print(DimPlot(seu.down, group.by = predicted.cell.type, split.by = 'gating.main'))
  
  
  
}


# hypergate( xp = matrix, gate_vector = labels, level = thing to gate)




```

The glia is only 50/50 and no good
The other cell type accuracy is high





Try plotting the results to see where the labelled cells end up

```{r}
# how to see that gated populations???
# gating predicted is a logical if it is in or not
length(gating_predicted)
dim(seu.down)

# add the logical as a meta data 
seu.down <- AddMetaData(object=seu.down, metadata=gating_predicted, col.name = 'gate_predicted')



DimPlot(seu.down, group.by = 'gate_predicted', split.by = 'gating.main')


```











Instead of subsampling in advance I can sample using hgate_sample to also sample negative events. This just gives you an idea of how many cells you need to sample

```{r}
# this is just to check the frequency of events

input.xm = as.matrix(GetAssayData(seu, slot = 'data'))
xm.t <- t(input.xm)
cluster.labels <- as.vector(seu$gating.main)

# hypergate( xp = matrix, gate_vector = labels, level = thing to gate)
sample = hgate_sample(gate_vector = cluster.labels, level = "Astro1", size=1000)


tab = table(ifelse(sample, "In", "Out"), ifelse(cluster.labels == "Astro1", "Positive pop.", "Negative pop."))
tab[1, ]/colSums(tab)  ## Fraction of subsampled events for positive and negative populations

```




