---
title: "Figure 2 and sup"
output: html_notebook
---

From Seurat parameter exploration select kn =25, res =0.25

```{r}
# set up enviroment / load libraries
library(Seurat)
library(dplyr)
library(ggplot2)
rm(list=ls())

```

Read in the Seurat object saved from the parameter optimization loop


```{r}
# make the seurat object
input_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/2Dcells_surface/preprocessing/select/2DcellsSelectflowset.csv"
# note this is not aligned or transformed

# output pathway
output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/2Dcells_surface/Figure2/FigureStuff/"
# add input description to ouptput files
input_name <- "Flowset_Select"  # this will be the different processing types

# cluster type for file name
clust_method <- "Louvain"

# read in the dataframe
df <- read.csv(input_path)
# print info to log 
print(dim(df)) # this is specific df has 73578 cells
# the preprocessing output csv needs to be cleaned - it contains live dead, FSC, SSC and the sample column
print(colnames(df))
# create a df with just the expression 
# need a way to automate this selection 
# I only want the expression values
df2 <- df %>% select(c("AQP4", "CD24", "CD44","CD184","CD15","HepaCAM","CD29","CD56", "O4","CD140a","CD133","GLAST","CD71"))
# the order of the DF is set by the order the columns are written above
# create a matrix for later
m <- as.matrix(df2) 

# create the seurat object for visualization

tm <- t(df2)
rownames(tm) <- colnames(df2)
colnames(tm) <- rownames(df2)
seu <- CreateSeuratObject(tm) # create a seurat object 

# add the meta data back in for sample groups
seu <- AddMetaData(object=seu, metadata=df$Batch, col.name = 'Batch')
# this doesn't work for making levels
# create the vector for the antibodies names for feature plotting later
AB <- colnames(df2)
# add to scale data slot
seu <- ScaleData(seu, features = AB)

```

Make heatmap of expression by cell type sample

```{r}
# plot by batch (cell culture sample)

DoHeatmap(seu, group.by = "Batch", features = AB, size= 6,slot = "scale.data", group.colors = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","cyan4"), disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 20))

# colours : dark blue, grey, orange

#c("red","yellow","blue","green","pink")
# c("#FOF7D4","#OF7DA9","#9EBB35","#B92F22","#72B9D5") error in these colours
# check the data
pdf(paste(output_path,input_name,clust_method,"Heatmap_2DcellSamples.pdf",sep=""),width =6, height = 4)
print(DoHeatmap(seu, group.by = "Batch", features = AB, size= 6,slot = "scale.data", group.colors = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","cyan4"), disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 20)))
dev.off()

#DoHeatmap(object = pbmc_small) + scale_fill_gradientn(colors = c("blue", "white", "red"))

```


UMAPS of cell types

```{r}
# make the UMAP
seu <- RunPCA(seu, features = AB, npcs = 13)
seu <- FindNeighbors(seu, dims = 1:12, k = 25)
seu <- RunUMAP(seu, dims = 1:12, n.neighbors = 25)

DimPlot(seu, group.by = 'Batch', repel = TRUE, cols = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","cyan4")) 

pdf(paste(output_path,input_name,clust_method,"kn25UMAPcelltypes.pdf",sep=""),width =10, height = 5)
print(DimPlot(seu, group.by = 'Batch', repel = TRUE, cols = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","cyan4")) )
dev.off()

```

Make the clusters  and plot the clusters

```{r}
seu <- FindClusters(seu, resolution = 0.25)

```

Visualize the clusters

```{r}

DimPlot(seu, group.by = 'RNA_snn_res.0.25', repel = TRUE,shape.by = 'Batch', raster = FALSE, pt.size = 0.25) 

#cols.highlight = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","cyan4")

pdf(paste(output_path,input_name,clust_method,"UMAPcelltypeandClustersres0.25.pdf",sep=""),width =8, height = 5)
 print(DimPlot(seu, group.by = 'RNA_snn_res.0.25', repel = TRUE,shape.by = 'Batch', raster = FALSE, pt.size = 0.25) ) 
dev.off()


DimPlot(seu, group.by = 'Batch' , repel = TRUE,shape.by = 'RNA_snn_res.0.25', raster = FALSE, pt.size = 0.5, cols = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","cyan4")) 

#cols.highlight = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","cyan4")

pdf(paste(output_path,input_name,clust_method,"UMAPClustersres0.25andCelltype.pdf",sep=""),width =8, height = 5)
 print(DimPlot(seu, group.by = 'Batch' , repel = TRUE,shape.by = 'RNA_snn_res.0.25', raster = FALSE, pt.size = 1, cols = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","cyan4")) ) 
dev.off()



```

Heatmap of the clusters

```{r}


# colours : dark blue, grey, orange

#c("red","yellow","blue","green","pink")
# c("#FOF7D4","#OF7DA9","#9EBB35","#B92F22","#72B9D5") error in these colours
# check the data
pdf(paste(output_path,input_name,clust_method,"Heatmap_2DcellClusterkn25Res0.25.pdf",sep=""),width =12, height = 4)
print(DoHeatmap(seu, group.by = "RNA_snn_res.0.25", features = AB, size= 4,slot = "scale.data", disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 20)))
dev.off()

#c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","cyan4")

```

Label clusters and plot again

```{r}
Idents(seu) <- "RNA_snn_res.0.25"
cluster.ids <- c("Astrocytes", "Early Neurons", "iPSCs", "NPC", "Oligo","Radial Glia","Neurons","Neural Stem","Mix-AQP4+","NPC-like","iPSCs")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)

# save the labels to meta data
seu[["Lables"]] <- Idents(seu)

# plot the UMAP and heatmap

#UMAP
DimPlot(seu, group.by = 'Lables', repel = TRUE) 

pdf(paste(output_path,input_name,clust_method,"kn25UMAPcelltypesLabelled2.pdf",sep=""),width =10, height = 5)
print(DimPlot(seu, group.by = 'Lables', repel = TRUE, label = TRUE) )
dev.off()


DoHeatmap(seu, group.by = 'Lables', features = AB, size= 4,slot = "scale.data", disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 20))
# heatmap
pdf(paste(output_path,input_name,clust_method,"Heatmap_2DcellClusterkn25Res0.25Labelled.pdf",sep=""),width =12, height = 4)
print(DoHeatmap(seu, group.by = 'Lables', features = AB, size= 4,slot = "scale.data", disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 20)))
dev.off()

#c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","cyan4")


```



Save the object with labels to try and apply these labels to the organoid using the new Seurat method and see if it can work.

```{r}

saveRDS(seu, paste(output_path,input_name,clust_method,"labels.Rds",sep=""))

```


Test if the file saved correctly 

```{r}
seu <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/2Dcells_surface/Figure2/FigureStuff/Flowset_SelectSeuratlabels.Rds")
```


make the umpa with labels

```{r}

DimPlot(seu, group.by = 'Lables', repel = TRUE, label = TRUE)
DimPlot(seu, group.by = 'Batch', repel = TRUE, label = TRUE)

```



Try to hightlight the original cells

```{r}

Idents(seu) <- 'Batch'
A <- WhichCells(seu, idents = c("Astrocytes"))
I <- WhichCells(seu, idents = c("IPSC"))
N <- WhichCells(seu, idents = c("Neurons"))
NPC <- WhichCells(seu, idents = c("NPC"))
O <- WhichCells(seu, idents = c("Oligo"))


DimPlot(seu, group.by = 'Lables', cells.highlight = list(NPC,N,I,A,O),cols.highlight = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","darkblue"), label = TRUE, pt.size = 0.005, cols = "grey")

unique(seu2@meta.data$Batch)

# save the plot

pdf(paste(output_path,input_name,clust_method,"kn25UMAPcolourcelltypeLabelclusters.pdf",sep=""),width =10, height = 5)
print(DimPlot(seu, group.by = 'Lables', cells.highlight = list(NPC,N,I,A,O),cols.highlight = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","darkblue"), label = TRUE, pt.size = 0.005, cols = "grey"))
dev.off()





```




Add in the correlaiton labels

```{r}
# correlation labels from Feb 21, 2022

cor.df <- read.csv("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/2Dcells_surface/correlations/corr_celltypes_2Dcells2.csv")

head(cor.df)

# add the labels as meta data

seu <- AddMetaData(object=seu, metadata=cor.df$best.cell.type, col.name = 'celltype')
seu <- AddMetaData(object=seu, metadata=cor.df$cell.label, col.name = 'cell.label')

```


Plot the correlations

```{r}

DimPlot(seu, group.by = 'celltype', repel = TRUE, label = TRUE)
DimPlot(seu, group.by = 'cell.label', repel = TRUE, label = TRUE)



DimPlot(seu, group.by = 'celltype', cells.highlight = list(NPC,N,I,A,O),cols.highlight = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","darkblue"), label = TRUE, pt.size = 0.005, cols = "grey")


DimPlot(seu, group.by = 'cell.label', cells.highlight = list(NPC,N,I,A,O),cols.highlight = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","darkblue"), label = TRUE, pt.size = 0.005, cols = "grey")

```


save all the plots

```{r}

output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/2Dcells_surface/Figure2/FigureStuff/"

pdf(paste(output_path,input_name,clust_method,"kn25res025UMAPcorcelltype.pdf",sep=""),width =8, height = 5)
DimPlot(seu, group.by = 'celltype', repel = TRUE, label = TRUE)
dev.off()


pdf(paste(output_path,input_name,clust_method,"kn25res025UMAPcorcelllabel.pdf",sep=""),width =20, height = 5)
DimPlot(seu, group.by = 'cell.label', repel = TRUE, label = TRUE)
dev.off()


pdf(paste(output_path,input_name,clust_method,"kn25res025UMAPcorcelltypeand2D.pdf",sep=""),width =8, height = 5)
DimPlot(seu, group.by = 'celltype', cells.highlight = list(NPC,N,I,A,O),cols.highlight = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","darkblue"), label = TRUE, pt.size = 0.005, cols = "grey")
dev.off()



pdf(paste(output_path,input_name,clust_method,"kn25res025UMAPcorcelllabeland2D.pdf",sep=""),width =14, height = 6)
DimPlot(seu, group.by = 'cell.label', cells.highlight = list(NPC,N,I,A,O),cols.highlight = c("cadetblue","bisque3","darkolivegreen3","darkgoldenrod3","darkblue"), label = TRUE, pt.size = 0.005, cols = "grey")
dev.off()

```



