---
title: "R Notebook"
output: html_notebook
---


```{r}
# load libraries
# load libraries
library(clusterSim) 
library(FlowSOM)
library(flowCore)
library(cluster)
library(fpc)
library(clv)
library(Seurat)
library(dplyr)
library(ggplot2)
library(clustree)
library(reshape2)

```

Visialize correlations on UMAP

```{r}
# read in correlation

# read in Seurat object to make UMAP
seu <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/cluster_parameters/LouvainFinal/FlowAlignRetroTransLouvainSeuratObject60.Rds")


# keep feature order the same as in other figures

AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")



```


Add the cell labels into seurat object

```{r}

cor <- read.csv("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/correlation/retro/9celltypes/corr_celltypes.csv")

seu <- AddMetaData(object=seu, metadata=cor$cell.label, col.name = 'cor.labels')
unique(seu$cor.labels)

```


Make UMAP


```{r}
seu <- RunUMAP(seu,spread = 1.2, min.dist = 0.25, dims = 1:12)
DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'RNA_snn_res.0.4', repel = TRUE)

DimPlot(seu, reduction = "umap", label = FALSE, group.by = 'cor.labels', repel = TRUE) #+ theme(legend.position ="none")




```


Save UMAPS

```{r}

outpath <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/"

pdf(paste(outpath,"UMAPCorLables9celltypes202203018.pdf"), width = 20, height = 8)
DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'cor.labels', repel = TRUE, shuffle = TRUE)
dev.off()


pdf(paste(outpath,"Fig3UMAP202203018.pdf"), width = 12, height = 6)
DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'RNA_snn_res.0.4', repel = TRUE)
dev.off()



# save the clusters

#pdf(paste(outpath,"UMAPlouvk60res02Clusters.pdf"), width = 20, height = 8)
#DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'RNA_snn_res.0.5', repel = TRUE)
#dev.off()


```


```{r}
# save feature plot with new UMAP structure and better dimensions
pdf(paste(outpath,"UMAPlouvk60res02FeaturePlots.pdf"), width = 20, height = 8)
FeaturePlot(seu, reduction = "umap", features = AB,slot = 'scale.data',min.cutoff = 'q1', max.cutoff ='99',label.size = 1)+ theme(plot.title = element_text(size = 0.1))
dev.off()

FeaturePlot(seu, reduction = "umap", features = AB,slot = 'scale.data',label.size = 1)+ theme(plot.title = element_text(size = 0.1))




```


Add labels

0-Epithelial
1-Astrocytes-RG
2-Neurons
3-NPC (neural stem)
4-RG-CD133
5-RG-CD184
6-Astrocytes
7-Neurons-early
8-Astrocytes
9-Glia (RG/neuron)
10-Endothelial
11-Oligodendrocytes
12-mix
13-OPC
14-Neurons-early


```{r}
Idents(seu) <- "RNA_snn_res.0.4"
cluster.ids <- c("Epithelial","Astro-RG","Neurons","NPC","RG-1","RG-2","Astro-1","Neurons-early","Astro-2","Glia","Endothelial","Oligo","mix","OPC","Neurons-early")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$labels <- Idents(seu)

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'labels', repel = TRUE)

# few labels

Idents(seu) <- "RNA_snn_res.0.4"
cluster.ids <- c("Epithelial","Astrocytes","Neurons","Neural-Stem","RG","RG","Astrocytes","Neurons-early","Astrocytes","RG","Endothelial","Oligo","mix","OPC","Neurons-early")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$labels2 <- Idents(seu)

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'labels2', repel = TRUE)


Idents(seu) <- "RNA_snn_res.0.4"
cluster.ids <- c("Epithelial","Astro","Neurons","NPC","RG-1","RG-2","Astro","Neurons-early","Astro","Glia","Endothelial","Oligo","mix","OPC","Neurons-early")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$labels3 <- Idents(seu)

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'labels3', repel = TRUE)

Idents(seu) <- "RNA_snn_res.0.4"
cluster.ids <- c("Epithelial","Astrocytes-1","Neurons","NPC","RG-1","RG-2","Astrocytes-2","Neurons-early","Astrocytes-1","Astrocytes-2","Endothelial","Oligo","mix","OPC","Neurons-early")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$labels4 <- Idents(seu)

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'labels4', repel = TRUE)


```


```{r}
# save the labelled seurat object to apply labels to other clustering
# use an inputs to Random Forest
saveRDS(seu,"/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/Louvkn60DifferentCellLabels220220318.Rds")


# save some plots
pdf(paste(outpath,"UMAPcelltype20220318.pdf"), width = 8, height = 5)
DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'labels4', repel = TRUE, shuffle = TRUE, label.size = 5)
dev.off()



pdf(paste(outpath,"HeatMapcell20220318.pdf"), width = 8, height = 5)
DoHeatmap(seu, group.by = 'labels4', features = AB, size= 5,slot = "scale.data", disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 15))
dev.off()


```

Visualize split UMAPs by batch 

```{r}
outpath <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/"

pdf(paste(outpath,"UMAPsplitBatch_clusterlabels2.pdf"), width = 30, height = 5)
  DimPlot(seu, split.by = 'Batch', label = TRUE)
dev.off()


FeaturePlot(seu, split.by = 'Batch', features = "CD24")
DimPlot(seu, group.by = 'Batch', label = FALSE, shuffle = TRUE)


```

Plot the correlations with less mix-cell types

```{r}
 df.filter <- cor %>% group_by(cell.label) %>% dplyr::filter(n()> 500)

plot1 <- ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
   theme(axis.text.x=element_text(angle=90))+ xlab('Assigned cell type') + ylab('number of cell')
  print(plot1)

  # # plot
  pdf(paste(outpath,"FreqCellTypes_filt500.pdf",sep=""),width =12, height = 6)
  plot1 <- ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
   theme(axis.text.x=element_text(angle=90))+ xlab('Assigned cell type') + ylab('number of cell')
  print(plot1)
  dev.off()


```


Try to calculate how many of each cell type are in each cluster

```{r}

# total proportion in each cell type
table(seu$RNA_snn_res.0.4)


```


```{r}
table(seu$labels4)
#proportions
prop.table(table(seu$labels4))

```


```{r}

# cor labels per cluster label??

cor.lables <- as.data.frame(table(seu$labels4, seu$cor.labels))
pr.cor.lables <- as.data.frame(prop.table(table(seu$labels4, seu$cor.labels)))
cor.lables$Freq <- as.double(cor.lables$Freq)
# try to make a heatmap

# not working
hm <- ggplot(data = pr.cor.lables, mapping = aes(x= Var1, y = Var2, fill = Freq))
hm

# try bar chart
ggplot(cor.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity")


#Filter by most frequent per group




```


Proportions of each sample:

```{r}
sample.lables <- as.data.frame(table(seu$Batch, seu$labels4))

sample.lables$Freq <- as.double(sample.lables$Freq)
# try to make a heatmap


# try bar chart
ggplot(sample.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity")+ theme_classic() + theme(axis.text.x=element_text(angle=90))+ xlab('Batch') + ylab('Frequency of Cell type')



```


Use correlation labels to help label clusters

```{r}

cor.lables <- as.data.frame(table(seu$RNA_snn_res.1.2, seu$cor.labels))
cor.lables$Freq <- as.double(cor.lables$Freq)
# try to make a heatmap


# try bar chart
ggplot(cor.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity")
```


See the same format of UMAP
```{r}
DimPlot(seu, group.by = "RNA_snn_res.1.2", label = TRUE)

```



Make a lot of rigde plots

```{r}


RidgePlot(seu, group.by = "RNA_snn_res.1.2", features = AB, log = TRUE)
RidgePlot(seu, group.by = "RNA_snn_res.1.2", features = "CD15", log = TRUE)
RidgePlot(seu, group.by = "RNA_snn_res.1.2", features = "CD29", log = TRUE)
RidgePlot(seu, group.by = "RNA_snn_res.1.2", features = "CD184", log = TRUE)
RidgePlot(seu, group.by = "RNA_snn_res.1.2", features = "HepaCAM", log = TRUE)
RidgePlot(seu, group.by = "RNA_snn_res.1.2", features = "CD24", log = TRUE)
RidgePlot(seu, group.by = "RNA_snn_res.1.2", features = "CD56", log = TRUE)
RidgePlot(seu, group.by = "RNA_snn_res.1.2", features = "CD44", log = TRUE)
RidgePlot(seu, group.by = "RNA_snn_res.1.2", features = "AQP4", log = TRUE)
RidgePlot(seu, group.by = "RNA_snn_res.1.2", features = "CD140a", log = TRUE)

```

Add some new labels from resolution 1.2



```{r}

Idents(seu) <- "RNA_snn_res.1.2"
cluster.ids <- c("Mix","RG-1","Neurons-Mix","Neurons-1","NPC","Astro-1","Glia-1","RG-2","EarlyNeuron-endo","RG-3","Endothelial","Astro-2","Neurons","Glia-2","Astro-3","Astro-2","RG-4","Glia-3","NPC","Astro-1","Oligo","mix","OPC","Endothelial","EarlyNeuron-2","Epithelial")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$labels5 <- Idents(seu)

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'labels5', repel = TRUE)


Idents(seu) <- "RNA_snn_res.1.2"
cluster.ids <- c("Mix","RG-1","Neurons-Mix","Neurons","NPC","Astro-1","Astro-2","RG-2","EarlyNeurons","RG-3","Endothelial","Astro-2","Neurons","Astro-1","RG-1","Astro-2","RG-2","Astro-2","NPC","Astro-1","Oligo","Mix","OPC","Endothelial","EarlyNeurons","Epithelial")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$labels6 <- Idents(seu)

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'labels6', repel = TRUE)



```






See the expression in each group in ridge plot

```{r}

RidgePlot(seu, group.by = "labels4", features = AB, log = TRUE)

RidgePlot(seu, group.by = "labels6", features = AB, log = TRUE)

```





Get mean expression per group/cell type  of each antibody 

```{r}


```



Label transfer using Seurat

```{r}
# single cell reference mapping
# uses the function TransferData

# I'll try a different kn in the same data as the query

seu.q <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/cluster_parameters/LouvainFinal/FlowAlignRetroTransLouvainSeuratObject20.Rds")

# the current labelled seurat object will be the reference  
#saveRDS(seu,"/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/Louvkn60DifferentCellLabels220220318.Rds")
  
#reference 
seu.r<- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/Figure3/Louvkn60DifferentCellLabels220220318.Rds")


# find the anchors for label transfer
anchors <- FindTransferAnchors(reference = seu, query = seu.q, features = AB, reference.reduction = "pca", dim= 1:10) 
# this step is computationally intensive and take a long time

seu.q <- IntegrateData(anchorset = anchors, new.assay.name = 'integrated', dims = 1:10, features = AB)


```






Maybe the dotplot can work now???  Nope - but it worked once before so why not now????


```{r}
seu <- ScaleData(seu)
DotPlot(seu, features = AB, cols = c("red","blue"), group.by = 'RNA_snn_res.0.4', dot.scale = 8) + RotatedAxis()

```





