---
title: "R Notebook"
output: html_notebook
---

```{r}
library(ggplot2)
library(Seurat)
library(dplyr)
library(reshape2)
```



Explore correlation

Looking at the list of cell types assigned 

```{r}

# new dataframe for exploring correlations
df <- corr_df
# read in saved df 

# see the size of difference between top correlation and second highest
df$oneVstwo <- (df$`best correlation` - df$`second correlation`) < 0.05

total <- dim(df)[1] # 73578     6

# check the number of positive cells
sum(df$oneVstwo, na.rm = TRUE)

# proportion of cells with close labels
print(sum(df$oneVstwo, na.rm = TRUE)/dim(df)[1])


```


Now see what the double cell types are 

```{r}
df$'double cell type' <- paste(df$`best cell type`,df$`second cell type`,sep = "-")

# this takes all the double cell types
unique(df$`double cell type`)



```

Try to make a summary table

```{r}
df.double <- df %>% select(`double cell type`)

df.table <- as.data.frame(table(df.double))
df.table

```

Apply a cell type label only using double labels when the correlation is close
```{r}
# if difference between top correlation and second correlation is < 0.05 than add double label
# else keep top cell type label

# assign cell type as unknown if max correlation is < 0.1


df$cell.lable <- ifelse(df$oneVstwo == TRUE, paste(df$`best cell type`,df$`second cell type`,sep = "-"), df$`best cell type`)


# condition on difference between first and second top correlations
df$cell.lable <- ifelse((df$`best correlation` - df$`second correlation`) < 0.05, paste(df$`best cell type`,df$`second cell type`,sep = "-"), df$`best cell type`)

# add another condition if if max correlation is < 0.1 cell.label = "unknown"



unique(df$cell.lable)


```


See the new counts

```{r}
df.cells <- df %>% select(`cell.lable`)

df.table2 <- as.data.frame(table(df.cells))
df.table2

```


Plot the correlations and the cell types

```{r}



```



Try visualizing the cell assignments on the UMAP

```{r}
# Seurat object UMAP already created 
# read one in here
seu <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/test/Louvain2/FlowsetSeuratSeuratObject100.Rds")
# add the cell type labels as meta-data

# read in the df from correlation
df.cells <- read.csv("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/correlationscorr_df_9000MBOcelltypelabels.csv")

seu <- AddMetaData(object=seu, metadata=df.cells$cell.label, col.name = 'cor.labels')

DimPlot(seu, group.by = 'cor.labels')
```


```{r}
```



```{r}
output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/correlations"

#write.csv(corr_df,"/Users/shuming/Desktop/PhenoID_single_cell_flow_cytometry_analysis/correlation/correlation_feb14.csv", row.names = FALSE)
write.csv(df, paste(output_path, "corr_df_9000MBOcelltypelabels.csv",sep=""))


# save the UMAP made above

pdf(paste(output_path,'UMAPcorrcelllabelsFeb15.pdf',sep=""),width =20, height = 10)
  print(DimPlot(seu, group.by = 'cor.labels'))
  dev.off()


```


Look at the 2D cell type data:

```{r}
# read in the save correlations

input_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/2Dcells_surface/correlationscorr_df_2Dcells.csv"
df.2d <- read.csv(input_path)

df.cells.2 <- df.2d %>% select(best.cell.type)

df.table3 <- as.data.frame(table(df.cells.2))
df.table3



```

Plot correlations and cell type frequency or proportions

```{r}
ggplot(df, aes(x=cell.lable))+ geom_bar()+theme_classic()+
  theme(axis.text.x=element_text(angle=90))

# consider plotting after filtering for cell types with more than 10 or 100 cells
# filter 
df.filter <- df %>% group_by(cell.lable) %>% filter(n()> 100)

ggplot(df.filter, aes(x=cell.lable))+ geom_bar()+theme_classic()+
  theme(axis.text.x=element_text(angle=90))

#try a proportion
ggplot(df.filter, aes(x=cell.lable))+ geom_bar(aes(y = (..count..)/sum(..count..)))+theme_classic()+
  theme(axis.text.x=element_text(angle=90))






```

Read in the data

```{r}

df <- read.csv("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/2Dcells_surface/correlations/corr_celltypes_2Dcells.csv")

df <- df %>% select(-X.1)

```


Create long df to plot the correlations

```{r}
#df <- df %>% select(-X.1)
df.melt <- melt(df)

# second corr values are sometime so low it messes up the plot

# vln plot of the 
# need to reformat the df
ggplot(df, aes(x=best.cell.type, y=best.correlation ))+ geom_violin()+ ylim(-0.1,1)+theme_classic()+
  theme(axis.text.x=element_text(angle=90))

# this plot the best and second best corr are together 
ggplot(df.melt, aes(x=cell.lable, y=value ))+ geom_boxplot()+ ylim(-0.1,1)+theme_classic()+
  theme(axis.text.x=element_text(angle=90))

# splitting the max and 2nmax corr
ggplot(df.melt, aes(x=best.cell.type, y=value, fill= variable, colour = variable ))+ geom_boxplot()+ ylim(-0.25,1)+theme_classic()+
  theme(axis.text.x=element_text(angle=90))
# the second best correlation is so low it was removed from STEM with axis limit -0.1 and even -1



```


```{r}

library(plyr)

# this take a huge amount of time to run - next time downsample by randomly selecting rows maybe 5000
# it cannot run

# need to sample before melt to keep matching data.
df <- rename(df, c("best.correlation"= "cor.1", "second.correlation" = "cor.2"))


# down sample
set.seed(64)
df.downsample <- sample_n(df, 1000)
df.melt.down <- melt(df.downsample)

# reformat the table to work with the before after plot
# y is the measurement in df.melt = value
# x is before after in df.melt = variable
# class another variable - in the example this is different shapes - for us this is best cell type
# might use facet to split the cell type - needs to be a factor
# id is the individual id this is the X column
ggplot(df.melt.down, aes(x = variable, y = value,colour= variable, group= X)) +
  geom_line(show.legend = F, size = 0.1, color = "black") + geom_point()+ scale_color_manual(values = c("#4E84C4", "#52854C")) + ylim(-0.25,0.95) +
  facet_wrap(~(as.factor(best.cell.type))) +
#theme_few() +
#theme(legend.position = "none") +
ylab("Correlation Coefficient") +
  xlab("")



# see with the double labels

ggplot(df.melt.down, aes(x = variable, y = value,colour= variable, group= X)) +
  geom_line(show.legend = F, size = 0.1, color = "black") + geom_point()+ scale_color_manual(values = c("#4E84C4", "#52854C")) + ylim(-0.25,0.95) +
  facet_wrap(~(as.factor(cell.lable))) +
#theme_few() +
#theme(legend.position = "none") +
ylab("Correlation Coefficient") +
  xlab("")


# this will be an excellent visualization but I need to subset only the double labels, then I can plot more cells and see more clearly. 

double.cells <- df[grep("-", df$cell.lable),]

df.melt.double <- melt(double.cells)

ggplot(df.melt.double, aes(x = variable, y = value,colour= variable, group= X)) +
  geom_line(show.legend = F, size = 0.1, color = "black") + geom_point()+ scale_color_manual(values = c("#4E84C4", "#52854C")) + ylim(-0.25,0.95) +
  facet_wrap(~(as.factor(cell.lable))) +
#theme_few() +
#theme(legend.position = "none") +
ylab("Correlation Coefficient") +
  xlab("")





```
Remake Freq plot - more attractive and in order of most to least frequent

```{r}
df.filter <- df %>% group_by(cell.label) %>% dplyr::filter(n()> 100)

# bar plot of how many cells get each label
# reorder x axis to be most to least frequent @Shuming


pdf(paste(output_path,"FreqCellTypesRankColours.pdf",sep=""),width =12, height = 6)
ggplot(df.filter, aes(x=reorder(cell.label,cell.label,function(x)-length(x)), fill = cell.label))+ geom_bar()+theme_classic() +
  theme(axis.text.x=element_text(angle=90))+ xlab('Assigned cell type') + ylab('number of cell')
dev.off()




```



