---
title: "R Notebook"
output: html_notebook
---

Annotate the full 9MBO dataset
1. Use Random Forest Model (RFM) - train model from Annotated hMO subset
2. Correlation assignment model (CAM)  - run correlation 
3. Use Seurat label transfer        - transfer labels from annotated hMO subset

For each prediction method visualize on UMAP and get frequency of cell type predictions per cluster (bar chart and df)


```{r}
# load libraries

library(Seurat)
library(dplyr)
library(ggplot2)
library(reshape2)
library(randomForest)
library(scProportionTest)

```


```{r}

# output path for data objects
outpath <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/"
# output path for figures
output_path <- "/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/"
# read in files needed:
# data object with clustering done ready to label
seu.q <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outs/clusters/Louvain/Allcellsretros_LouvainSeuratObject60.Rds")
# reference hMO subset annotated
seu.r<- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/Seu9000annot.08072021.RDS")

# antibody features in order
AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")


seu <- readRDS(paste(outpath,"seuAllcells.15072022.Rds"))

```



Random Forest Classifier


```{r}

# A function will be made for training the model and running it.
# For now I'm going to run the training and run the predictions with the new model
# see the file: 'RandomForest.Fig4.July10.R'

# read in the saved model
rf <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/RFM_trained.11072022.Rds")

# run the model to predict cell types in the unlabelled full hMO dataset
# need to prepare the full dataset in a way to be read by RFM
df <- transpose(as.data.frame(GetAssayData(seu.q,slot = 'scale.data')))
dim(df)
colnames(df) <- AB
rfm.pred <- predict(rf,df)
head(rfm.pred)
results.df <- as.data.frame(rfm.pred)
head(results.df)

# save the results
write.csv(results.df, paste(outpath, "RFMpredictionsAllcells.12072022.csv"))

# add the predictions to the seurat object
seu.q <- AddMetaData(object=seu.q, metadata= as.factor(results.df$rfm.pred), col.name = 'rf.pred')
unique(seu.q$rf.pred)
# save the RFM predictions
saveRDS(seu.q, paste(outpath,"All9MOannaote.12072022.Rds"))


```


Correlation test

```{r}
# run the correlation prediction function on the full dataset
reference_path <- "/Users/rhalenathomas/GITHUB/PhenoID_single_cell_flow_cytometry_analysis/old/correlation/ReferenceMatrix9celltypesOrdered.csv"
output_path <- "/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/Corr_all/"

# full dataset
test_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outs/retrotransformed_flowset.csv"


#run the correlation function
find_correlation(test_path, reference_path, output_path)

# this produces a csv with the Frequencies of each cell type and a csv with the best and second best correlation coefficients
# read in those csv
corr_cell.df <- read.csv("/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/Corr_all/corr_celltypes.csv")
freq.df <- read.csv("/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/Corr_all/Frequencytabletypes.csv")


# add the correlation predictions to the meta data

seu.q <- AddMetaData(object=seu.q, metadata=corr_cell.df$cell.label, col.name = 'cor.labels')
unique(seu.q$cor.labels)

# save the RDS seurat object with correlation labels added

saveRDS(seu.q, paste(outpath,"All9MOannaote.12072022.Rds"))



```


Seurat label transfer

```{r}

Idents(seu.r) <- 'subgroups'
seu.r <- subset(seu.r, downsample = 500)
table(seu.r$subgroups)

anchors <- FindTransferAnchors(reference = seu.r, query = seu.q,features = AB ,reference.reduction = "pca", dim= 1:10) 
predictions <- TransferData(anchorset = anchors, refdata = seu.r$subgroups, dims = 1:10)
write.csv(predictions, paste(outpath,"all.cells.predictions.csv"))
seu.q <- readRDS( paste(outpath,"All9MOannaote.12072022.Rds"))
seu.q <- AddMetaData(seu.q, metadata = predictions)
# save the Seurat predictions

saveRDS(seu.q, paste(outpath,"All9MOannaote.12072022.Rds"))

  # label transfer in cluster labels

t.lables <- as.data.frame(table(seu.q$RNA_snn_res.1, seu.q$predicted.id))
pr.t.lables <- as.data.frame(prop.table(table(seu.q$RNA_snn_res.1, seu.q$predicted.id)))
t.lables$Freq <- as.double(t.lables$Freq)


# try bar chart
ggplot(t.lables, aes(y = Freq, x = Var1, fill = Var2)) + geom_bar(position = "stack", stat= "identity")

# find the top 2 in cell in the group

top.labs <- t.lables  %>% group_by(Var1)  %>% top_n(3, Freq)
top.labs

top.lab <- t.lables  %>% group_by(Var1)  %>% top_n(1, Freq)
top.lab






```


Add some labels 1 to the Seurat object

Label 1
glia-lineage
neural-lineage
RG1-a
Epi-RG
Neurons-1-a
RG3
Glia
Astro-1
NPC
RG1-2
Astro-m
RG1-b
Endo
Stem
Neurons-1-b
Astro-2-a
NPC-epi-RG1
Neurons1-2-a
Neurons1-2-b
Astro-2-3
Neur-Stem-OPC
Astro-2-b
Epi-a
Astro-3-2m-RG
Epi-b
NPC-RG2
Neurons-3-Astro-3
Neurons-3-1
Oligo
Epi-stem
Neuron-1-RG3
RG1-Neur1-NPC
Astro-m
Endo
Oligo-RG-OPC
astrocyte
	

```{r}

seu.q <- readRDS(paste(outpath,"All9MOannaote.12072022.Rds"))

Idents(seu.q) <- "seurat_clusters"
cluster.ids <- c("glia-lin","neu-lin","RG1-a","Epi-RG","Neur1-a",
                 "RG3","Glia","Astro-1","NPC","RG1-2a",
                 "Astro-m","RG1-2b","Endo-a","Stem","Neur1-b",
                 "Astro-2a","NPC-eip-rg","Neur1-2a","Neur1-2b","Astro2-3",
                 "Neur-stem-opc","Astro2b","Epi-a","Astro2-3-m-RG","Epi-b",
                 "NPC-RG","Neur3-astro3","Neur3-1","Oligo","Epi-stem",
                 "neur1-RG3","RG1-neu-npc","astro-m","Endo-b","Oligo-RG-OPC",
                 "Astro")

names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels1 <- Idents(seu.q)

DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'labels1', repel = TRUE)

### change names to be able to group same cell types together


Idents(seu.q) <- "seurat_clusters"
cluster.ids <- c("GL","NL","RG-1a","Epi-1","Neur1-a",
                 "RG-3","Glia","Astro-1","NPC-1","RG-1b",
                 "Astro-m1","RG-1c","Endo-a","Stem","Neur1-b",
                 "Astro-2a","NPC-2","Neur1-2a","Neur1-2b","Astro-3",
                 "Neur-3a","Astro-2b","Epi-2","Astro-4","Epi-3",
                 "NPC-3","Neur-3b","Neur-3c","Oligo","Epi-4",
                 "Neur1-c","RG-1b","Astro-m1","Endo-b","OPC",
                 "Astro-5")

names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels2 <- Idents(seu.q)

png(paste(output_path,"UMAP.all.cell.labels2.png"), width = 1000, height = 400)
DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'labels2', repel = TRUE, label.size = 6)
dev.off()


# merge together clusters that are overlapping or adjacent

Idents(seu.q) <- "seurat_clusters"
cluster.ids <- c("GL","NL","RG1a","Epi","Neur1a",
                 "RG3a","RG3b","Astro1a","Neur1b","RG2a",
                 "Astro-m1","RG1b","Endo","Stem-1","Neur1b",
                 "Astro2","Neur3","Neur2a","Neur2b","Astro3a",
                 "Neur2c","Astro3","Epi","Astro3b","Epi",
                 "RG2b","Neur3a","Neur2d","Oligo","Stem-2",
                 "Neur3b","RG3-3","Astro-m2","Endo","OPC",
                 "Astro1b")

names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels3 <- Idents(seu.q)

# larger groups

png(paste(output_path,"UMAP.all.cell.labels3.png"), width = 1000, height = 400)
DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'labels3', repel = TRUE, label.size = 6)
dev.off()

Idents(seu.q) <- "seurat_clusters"
cluster.ids <- c("GL","NL","RG1a","Epi","Neurons1",
                 "RG3a","RG3b","Astro1","Neurons1","RG2",
                 "Astro-m","RG1b","Endo","Stem","Neurons1",
                 "Astro2","Neurons3","Neurons2","Neurons2","Astro2",
                 "Neurons2","Astro2","Epi","Astro2","Epi",
                 "RG2","Neurons3","Neurons2","Oligo","Stem",
                 "Neurons3","Astro2","Astro-m","Endo","OPC",
                 "Astro1")

names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels.main <- Idents(seu.q)



#png(paste(output_path,"UMAP.all.cell.labels.main.png"), width = 1000, height = 400)
DimPlot(seu.q, reduction = "umap", label = TRUE, group.by = 'labels.main', repel = TRUE, label.size = 6)
dev.off()




### plot with different colours

clust.colours <- c("chocolate1","chocolate2","chocolate3","orangered",
                   "lightsalmon", "pink","lightpink3",
                   "steelblue3","deepskyblue",
                   "plum","plum3","purple","orchid2","mediumpurple2",
                   "seagreen3","tomato3","tomato4","burlywood3","grey90","brown")
# Figure S16 
#order = cell.type.order

#pdf(paste(output_path,"UMAP.all.groups.pdf"),width = 15, height = 8)
DimPlot(seu.q, cols = clust.colours, shuffle = TRUE, 
        raster=FALSE, pt.size = 0.1, label = TRUE, group.by = 'labels.main') +
  theme(legend.text = element_text(size=24), axis.title.y = element_text(size=24), 
        axis.title.x = element_text(size=24), axis.text.y = element_text(size =24),
        axis.text.x = element_text(size =24))
#dev.off()



Idents(seu.q) <- "seurat_clusters"
# some subgroups

cluster.ids <- c("GL","NL","RG1","Epi","Neurons1",
                 "RG3","Astro1","Astro1","Neurons1","RG2",
                 "Astro-m","RG1","Endo","Stem-like","Neurons1",
                 "Astro2","Neurons3","Neurons2","Neurons2","Astro2",
                 "Neurons2","Astro3","Epi","Astro3","Epi",
                 "RG2","Astro2","Neurons2","Oligo","Stem-like",
                 "Neurons3","Neurons3","Astro-m","Endo","OPC",
                 "Astro1")

names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels4 <- Idents(seu.q)

cell.type.order <- c("Neurons1","Neurons2","Neurons3",
                     "RG1","RG2","RG3","Astro1","Astro2","Astro3","Astro-m",
                     "Epi","Endo","OPC","Oligo","Stem-like","GL","NL")
cell.type.order <- rev(cell.type.order)

clust.colours <- c("plum4","purple3","orchid3",
                  "indianred1","pink","lightpink3",
                   "darkorange","chocolate3","sienna1","orangered",
                   "steelblue3","deepskyblue","blue","cyan",
                   "burlywood3","grey90","seashell2")

pdf(paste(output_path,"UMAP.subgroups.few.pdf"),width = 15, height = 8)
DimPlot(seu.q, cols = clust.colours, shuffle = TRUE, order = cell.type.order,
        raster=FALSE, pt.size = 0.1, label = TRUE, group.by = 'labels4')
dev.off()


### main groups
Idents(seu.q) <- "seurat_clusters"
cluster.ids <- c("Glia Lineage","Neural Lineage","Radial Glia","Epithelial","Neurons 1",
                 "Radial Glia","Astrocytes","Astrocytes","Neurons 1","Radial Glia",
                 "Astrocytes","Radial Glia","Endothelial","Stem cell like","Neurons 1",
                 "Astrocytes","NPC","Neurons 2","Neurons 2","Astrocytes",
                 "Neurons 2","Astrocytes","Epithelial","Astrocytes","Epithelial",
                 "Radial Glia","Astrocytes","Neurons 2","Oligodendrocytes","Stem cell like",
                 "NPC","NPC","Astrocytes","Endothelial","OPC",
                 "Astrocytes")

names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$labels.main <- Idents(seu.q)

cell.type.order <- c("Neurons 1","Neurons 2","NPC",
                     "Radial Glia","Astrocytes 1", "Astrocytes 2",
                     "Epithelial","Endothelial","OPC","Oligodendrocytes",
                     "Stem cell like","Glia Lineage","Neural Lineage")
cell.type.order <- rev(cell.type.order)

clust.colours <- c("plum3","purple3","orchid3",
                  "pink","darkorange",
                   "steelblue3","deepskyblue","blue","cyan",
                   "burlywood3","grey90","seashell2")



pdf(paste(output_path,"UMAP.main.groups.pdf"),width = 15, height = 8)
DimPlot(seu.q, cols = clust.colours, shuffle = TRUE, order = cell.type.order,
        raster=FALSE, pt.size = 0.1, label = FALSE, group.by = 'labels.main')
dev.off()


#### label with the same amount of levels as in the subset
#### use the same colour scheme and order as in figure 3
##### Figure 4 A plot

Idents(seu.q) <- "seurat_clusters"
cluster.ids <- c("Glia Lineage","Neural Lineage","Radial Glia 1","Epithelial","Neurons 1",
                 "Radial Glia 2","Astrocytes 1","Astrocytes 1","Neurons 1","Radial Glia 2",
                 "Astrocytes 2","Radial Glia 1","Endothelial","Stem cell like","Neurons 1",
                 "Astrocytes 1","NPC","Neurons 2","Neurons 2","Astrocytes 1",
                 "Neurons 2","Astrocytes 1","Epithelial","Astrocytes 2","Epithelial",
                 "Radial Glia 2","Astrocytes 1","Neurons 2","Oligodendrocytes","Stem cell like",
                 "NPC","NPC","Astrocytes 2","Endothelial","OPC",
                 "Astrocytes 2")



names(cluster.ids) <- levels(seu.q)
seu.q <- RenameIdents(seu.q, cluster.ids)
seu.q$cell.types <- Idents(seu.q)


# change the order of the cell types on the legend of the umap
cell.type.order <- c("Astrocytes 1", "Astrocytes 2","Radial Glia 1","Radial Glia 2",
                     "Epithelial","Endothelial","NPC","Neurons 1","Neurons 2",
                     "Oligodendrocytes","OPC","Stem cell like","Glia Lineage","Neural Lineage")
cell.type.order <- rev(cell.type.order)

# colour order to match cell type order
clust.colours <- c("chocolate1","orange","lightsalmon", "pink",
                   "steelblue3","deepskyblue","plum3","purple","orchid2",
                   "seagreen3","green","tomato3","burlywood3","grey90","lemonchiffon3")
 
Idents(seu.q) <- 'cell.types'
                  
pdf(paste(output_path,"UMAPlabelled.all.12072022.pdf"),width = 11, height = 5)
DimPlot(seu.q, order = cell.type.order, cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = FALSE) +
  theme(legend.text = element_text(size=24), axis.title.y = element_text(size=24), 
        axis.title.x = element_text(size=24), axis.text.y = element_text(size =24),
        axis.text.x = element_text(size =24))
dev.off()



# save the annotated dataset 
saveRDS(seu.q, (paste(outpath,"All9MOannaote.12072022.Rds")))


```


Make a dot plot of expression in full dataset


```{r}
# Figure 4B

# get the data

express.by.cluster <- as.data.frame(AverageExpression(seu.q, features = AB, group.by = 'cell.types', slot = 'scale.data'))
express.by.cluster <- as.data.frame(scale(express.by.cluster))

col.names.RNA <- colnames(express.by.cluster)

re.col.name <-  c("Glia Lineage","Neural Lineage","Radial Glia 1", "Epithelial",
                  "Neurons 1","Radial Glia 2","Astrocytes 1","Astrocytes 2",
                  "Endothelial","Stem cell like","NPC",
                  "Neurons 2","Oligodendrocytes","OPC")

names(express.by.cluster) <- re.col.name
AB2 <- row.names(express.by.cluster)

express.by.cluster <- scale(express.by.cluster)
ex.data <- cbind(AB2,express.by.cluster)


# reformat
longData<- melt(ex.data)

# get the proportion of cells expressing a marker
a <- DotPlot(seu.q, features = AB, group.by = 'cell.types')
a$data
pct.exp <- as.data.frame(a$data) %>% select(features.plot, id, pct.exp)

# add the mean expression and the percent cells expressing together
# rename the AB and cell labels 
colnames(longData) <- c("features.plot","id","expression")

df.exp.pct <- merge(longData, pct.exp, by = c("features.plot", "id"))



# reorder 
unique(df.exp.pct$id)
cell.type.order <- c("Astrocytes 1", "Astrocytes 2","Radial Glia 1","Radial Glia 2",
                     "Epithelial","Endothelial","NPC","Neurons 1","Neurons 2",
                     "Oligodendrocytes","OPC","Stem cell like","Glia Lineage","Neural Lineage")
cell.type.order <- rev(cell.type.order)

data <- df.exp.pct %>% mutate(Cell.type = factor(id, levels = cell.type.order))
data <- data %>% mutate(Marker = factor(features.plot, levels = AB))

# save the plot

pdf(paste(output_path,"Dotplot.labelled.all.12072022.pdf"),width = 8, height = 5)
ggplot(data = data, aes(x = Marker, y = Cell.type, color = expression, size = pct.exp)) +
  geom_point() + theme_classic() +
  scale_color_gradient2(low = "#154c79", mid = "#eeeee4", high = "#e28743") + ylab("Cell Phenotypes") + xlab("Antibodies") + RotatedAxis() +  theme(legend.text = element_text(size=15), axis.title.y = element_text(size=15), 
        axis.title.x = element_text(size=15), axis.text.y = element_text(size =15),
        axis.text.x = element_text(size =15, angle = 90)) 
dev.off()

# colors = c("#154c79", "#eeeee4", "#e28743")

```


Add groups to the seurat object

```{r}

seu <- readRDS(paste(outpath,"All9MOannaote.12072022.Rds"))

unique(seu$Batch)

# add genotype 
Idents(seu) <- "Batch"
cluster.ids <- c("3450","3450","3450","AIW002","AIW002","AIW002","AJG001","AJG001","AJG001")
names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$Genotype <- Idents(seu)

# add experiment date
Idents(seu) <- "Batch"
cluster.ids <- c("0306","0317","0317","0306","0317","0317","0306","0317","0317")
names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$Ex.date <- Idents(seu)

# add batch information
Idents(seu) <- "Batch"
cluster.ids <- c("B","A","B","B","A","B","B","A","B")
names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$mo.batch <- Idents(seu)

# add organoid age
Idents(seu) <- "Batch"
cluster.ids <- c("273","263","284","273","263","284","273","263","284")
names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$age <- Idents(seu)


DimPlot(seu)


```






Run split UMAP

```{r}


# change the order of the cell types on the legend of the umap
cell.type.order <- c("Astrocytes 1", "Astrocytes 2","Radial Glia 1","Radial Glia 2",
                     "Epithelial","Endothelial","NPC","Neurons 1","Neurons 2",
                     "Oligodendrocytes","OPC","Stem cell like","Glia Lineage","Neural Lineage")
cell.type.order <- rev(cell.type.order)

# colour order to match cell type order
clust.colours <- c("chocolate1","orange","lightsalmon", "pink",
                   "steelblue3","deepskyblue","plum3","purple","orchid2",
                   "seagreen3","green","tomato3","burlywood3","grey90","lemonchiffon3")
 
Idents(seu) <- 'cell.types'
      
DimPlot(seu, order = cell.type.order, cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = FALSE, split.by = 'Genotype')

seu$Genotype <- factor(x = seu$Genotype, levels = c("AIW002","AJG001","3450"))
           
# genotype 
pdf(paste(output_path,"UMAP.split.genotype.12072022.pdf"),width = 15, height = 3.5)
DimPlot(seu, order = cell.type.order, cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = FALSE,split.by = 'Genotype', label.size = 6) +
  theme(legend.text = element_text(size=16), axis.title.y = element_text(size=16), 
        axis.title.x = element_text(size=16), axis.text.y = element_text(size =16),
        axis.text.x = element_text(size =16))
dev.off()

# age
            
pdf(paste(output_path,"UMAP.split.age.12072022.pdf"),width = 15, height = 4)
DimPlot(seu, order = cell.type.order, cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = FALSE,split.by = 'age', label.size = 6) +
  theme(legend.text = element_text(size=18), axis.title.y = element_text(size=18), 
        axis.title.x = element_text(size=18), axis.text.y = element_text(size =18),
        axis.text.x = element_text(size =18))
dev.off()
# no visable differences

# batch 
pdf(paste(output_path,"UMAP.split.batch.12072022.pdf"),width = 15, height = 4)
DimPlot(seu, order = cell.type.order, cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = FALSE,split.by = 'mo.batch', label.size = 6) +
  theme(legend.text = element_text(size=18), axis.title.y = element_text(size=18), 
        axis.title.x = element_text(size=18), axis.text.y = element_text(size =18),
        axis.text.x = element_text(size =18))
dev.off()

# no visible differences

# ex date

pdf(paste(output_path,"UMAP.split.exdate.12072022.pdf"),width = 15, height = 4)
DimPlot(seu, order = cell.type.order, cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = FALSE,split.by = 'Ex.date', label.size = 6) +
  theme(legend.text = element_text(size=18), axis.title.y = element_text(size=18), 
        axis.title.x = element_text(size=18), axis.text.y = element_text(size =18),
        axis.text.x = element_text(size =18))
dev.off()

# ex date


```





```{r}
# test how clusters look with a different kn 

seu <- seu.q

#seu <- FindNeighbors(seu, dim = 1:12, k.param = 200)
seu <- FindClusters(seu, resolution = 1.5)

DimPlot(seu, group.by = 'cell.types')


# change the order of the cell types on the legend of the umap
cell.type.order <- c("Astrocytes 1", "Astrocytes 2","Radial Glia 1","Radial Glia 2",
                     "Epithelial","Endothelial","NPC","Neurons 1","Neurons 2",
                     "Oligodendrocytes","OPC","Stem cell like","Glia Lineage","Neural Lineage")
cell.type.order <- rev(cell.type.order)

# colour order to match cell type order
clust.colours <- c("chocolate1","orange","lightsalmon", "pink",
                   "steelblue3","deepskyblue","plum3","purple","orchid2",
                   "seagreen3","green","tomato3","burlywood3","grey90","lemonchiffon3")
 
Idents(seu) <- 'cell.types'
                  
pdf(paste(output_path,"UMAPlabelled.all.k200.12072022.pdf"), width = 11, height = 5)
DimPlot(seu, order = cell.type.order, cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = FALSE) +
  theme(legend.text = element_text(size=24), axis.title.y = element_text(size=24), 
        axis.title.x = element_text(size=24), axis.text.y = element_text(size =24),
        axis.text.x = element_text(size =24))
dev.off()

# the plot looks the same 

saveRDS(seu, (paste(outpath,"All9MOannaote.k200.12072022.Rds")))


```


Add different group data to compare

```{r}
unique(seu$Batch)

# add genotype 
Idents(seu) <- "Batch"
cluster.ids <- c("3450","3450","3450","AIW002","AIW002","AIW002","AJG001","AJG001","AJG001")
names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$Genotype <- Idents(seu)

# add experiment date
Idents(seu) <- "Batch"
cluster.ids <- c("0306","0317","0317","0306","0317","0317","0306","0317","0317")
names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$Ex.date <- Idents(seu)

# add batch information
Idents(seu) <- "Batch"
cluster.ids <- c("B","A","B","B","A","B","B","A","B")
names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$mo.batch <- Idents(seu)

# add organoid age
Idents(seu) <- "Batch"
cluster.ids <- c("273","263","284","273","263","284","273","263","284")
names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$age <- Idents(seu)


DimPlot(seu)


```




Quick check proportion of cell types across samples

```{r}

table(Idents(seu), seu$Batch)

# proportion across channels
prp.t <- as.data.frame(prop.table(table(Idents(seu))))





```




Proportions of cell types

```{r}

# genotype
sample.lables <- as.data.frame(table(seu$Genotype, seu$cell.types))

sample.lables$Freq <- as.double(sample.lables$Freq)

# bar chart of with percent 
ggplot(sample.lables, aes(x = Var1,y=Freq ,fill = Var2)) + geom_bar(position= "fill", stat = "identity") + scale_y_continuous(labels = scales::percent_format()) + theme_classic() + theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))  + xlab('Genotype') + ylab('Percent of Cell type') + RotatedAxis()



# change the order of the cell types on the legend of the umap
cell.type.order <- c("Astrocytes 1", "Astrocytes 2","Radial Glia 1","Radial Glia 2",
                     "Epithelial","Endothelial","NPC","Neurons 1","Neurons 2",
                     "Oligodendrocytes","OPC","Stem cell like","Glia Lineage","Neural Lineage")


# colour order to match cell type order
clust.colours <- c("chocolate1","orange","lightsalmon", "pink",
                   "steelblue3","deepskyblue","plum3","purple","orchid2",
                   "seagreen3","green","tomato3","burlywood3","grey90","lemonchiffon3")
 

### change cell type order


Genotype.order <- c("AIW002","AJG001","3450")


df <- sample.lables %>% mutate(Var2 = factor(Var2, levels = cell.type.order))
df <- df %>% mutate(Var1 = factor(Var1, levels = Genotype.order))



pdf(paste(output_path,"bar.pr.genotype.pdf"), width = 5, height = 3)
ggplot(df, aes(x = Var1,y=Freq ,fill = Var2)) + geom_bar(position= "fill", stat = "identity")  +
  RotatedAxis() +
  scale_y_continuous(labels = scales::percent_format()) + theme_classic() + theme(text = element_text(size=15),
        axis.text.x = element_text(angle=90, hjust=1))  + xlab('Genotype') + ylab('Percent of Cell type')  + scale_fill_manual(values = clust.colours)
dev.off()




```

Bar charts 


```{r}

# age



```


```{r}

# batch

```


```{r}

# ex date

```


Proportion tests

```{r}

library(scProportionTest)
# create a propotion test object
prop_test <- sc_utils(seu)

# set up the comparison

# need to compare separately
prop_test <- permutation_test(
	prop_test, cluster_identity = "cell.types",
	sample_1 = "AIW002", sample_2 = "AJG001",
	sample_identity = "Genotype"
)

# make the plot
permutation_plot(prop_test)

# save the plot 
png(paste(output_path,"prp.test.AIWvAJG.png"))
permutation_plot(prop_test)
dev.off()

prop_test <- permutation_test(
	prop_test, cluster_identity = "cell.types",
	sample_1 = "AIW002", sample_2 = "3450",
	sample_identity = "Genotype"
)

# make the plot
permutation_plot(prop_test)

# save the plot 
png(paste(output_path,"prp.test.AIWv3450.png"))
permutation_plot(prop_test)
dev.off()


prop_test <- permutation_test(
	prop_test, cluster_identity = "cell.types",
	sample_1 = "3450", sample_2 = "AJG001",
	sample_identity = "Genotype"
)

# make the plot
permutation_plot(prop_test)

# save the plot 
png(paste(output_path,"prp.test.3450vAJG.png"))
permutation_plot(prop_test)
dev.off()


# make a group of AIW vs other
Idents(seu) <- "Batch"
cluster.ids <- c("AJG-3450","AJG-3450","AJG-3450","AIW002","AIW002","AIW002",
                 "AJG-3450","AJG-3450","AJG-3450")
names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$ipsc <- Idents(seu)


DimPlot(seu, group.by = 'ipsc')

# join together AJG and 3450
prop_test <- permutation_test(
	prop_test, cluster_identity = "cell.types",
	sample_1 = "AIW002", sample_2 = "AJG-3450",
	sample_identity = "ipsc")

# make the plot
permutation_plot(prop_test)

# save the plot 
pdf(paste(output_path,"prp.test.AIWvs3450-AJG.pdf"),width = 9, height = 4)
permutation_plot(prop_test) + theme_bw() + 
  theme(axis.text.x=element_text(size=15),
                     axis.text.y=element_text(size=15))
dev.off()


seu



```


Heat map grouped by genotype

```{r}

#### now I need some why to compare expression between groups 
##### get the expression values 

exp.by.cell.geneotype <- as.data.frame(AverageExpression(seu, features = AB, group.by = c('cell.types','Genotype'), slot = 'scale.data'))
exp.by.cell.geneotype <- as.data.frame(scale(exp.by.cell.geneotype))
# add the rownames as a column

AB2 <- row.names(exp.by.cell.geneotype)
ex.data.c <- cbind(AB2,exp.by.cell.geneotype)


# reformat

longData<- melt(ex.data.c)

unique(longData$variable)

RNA.Glia.Lineage_AIW002     RNA.Glia.Lineage_AJG001     RNA.Glia.Lineage_3450      
 [4] RNA.Neural.Lineage_AIW002   RNA.Neural.Lineage_AJG001   RNA.Neural.Lineage_3450    
 [7] RNA.Radial.Glia.1_AIW002    RNA.Radial.Glia.1_AJG001    RNA.Radial.Glia.1_3450     
[10] RNA.Epithelial_AIW002       RNA.Epithelial_AJG001       RNA.Epithelial_3450        
[13] RNA.Neurons.1_AIW002        RNA.Neurons.1_AJG001        RNA.Neurons.1_3450         
[16] RNA.Radial.Glia.2_AIW002    RNA.Radial.Glia.2_AJG001    RNA.Radial.Glia.2_3450     
[19] RNA.Astrocytes.1_AIW002     RNA.Astrocytes.1_AJG001     RNA.Astrocytes.1_3450      
[22] RNA.Astrocytes.2_AIW002     RNA.Astrocytes.2_AJG001     RNA.Astrocytes.2_3450      
[25] RNA.Endothelial_AIW002      RNA.Endothelial_AJG001      RNA.Endothelial_3450       
[28] RNA.Stem.cell.like_AIW002   RNA.Stem.cell.like_AJG001   RNA.Stem.cell.like_3450    
[31] RNA.NPC_AIW002              RNA.NPC_AJG001              RNA.NPC_3450               
[34] RNA.Neurons.2_AIW002        RNA.Neurons.2_AJG001        RNA.Neurons.2_3450         
[37] RNA.Oligodendrocytes_AIW002 RNA.Oligodendrocytes_AJG001 RNA.Oligodendrocytes_3450  
[40] RNA.OPC_AIW002              RNA.OPC_AJG001              RNA.OPC_3450               
42 Levels: RNA.Glia.Lineage_AIW002 RNA.Glia.Lineage_AJG001 ... RNA.OPC_3450

#now we need to split the variable Cell type and Genotype to have less crazy labels - maybe use facet for each cell type.
# maybe filter out some cell types 


x.order <- c("RNA.Astrocytes.1_AIW002","RNA.Astrocytes.1_AJG001", "RNA.Astrocytes.1_3450",
             "RNA.Astrocytes.2_AIW002", "RNA.Astrocytes.2_AJG001", "RNA.Astrocytes.2_3450",
             "RNA.Radial.Glia.1_AIW002","RNA.Radial.Glia.1_AJG001", "RNA.Radial.Glia.1_3450",
             "RNA.Radial.Glia.2_AIW002","RNA.Radial.Glia.2_AJG001", "RNA.Radial.Glia.2_3450",
             "RNA.Epithelial_AIW002","RNA.Epithelial_AJG001","RNA.Epithelial_3450",
             "RNA.Endothelial_AIW002","RNA.Endothelial_AJG001","RNA.Endothelial_3450",
             "RNA.NPC_AIW002","RNA.NPC_AJG001", "RNA.NPC_3450",
             "RNA.Neurons.1_AIW002","RNA.Neurons.1_AJG001", "RNA.Neurons.1_3450",
            "RNA.Neurons.2_AIW002", "RNA.Neurons.2_AJG001", "RNA.Neurons.2_3450",
             "RNA.Oligodendrocytes_AIW002", "RNA.Oligodendrocytes_AJG001", "RNA.Oligodendrocytes_3450",
            "RNA.OPC_AIW002", "RNA.OPC_AJG001", "RNA.OPC_3450",
            "RNA.Stem.cell.like_AIW002","RNA.Stem.cell.like_AJG001", "RNA.Stem.cell.like_3450",
            "RNA.Glia.Lineage_AIW002","RNA.Glia.Lineage_AJG001","RNA.Glia.Lineage_3450",
            "RNA.Neural.Lineage_AIW002","RNA.Neural.Lineage_AJG001","RNA.Neural.Lineage_3450"
             )


df <- longData %>% mutate(variable = factor(variable, levels = x.order))


# heatmap
pdf(paste(output_path,"heatmap.ave.gene.cell.pdf"), width = 11, height = 5)
ggplot(df, aes(x = variable, y = AB2)) + 
  geom_raster(aes(fill=value)) + 
  scale_fill_gradient2(low = "#154c79", mid = "#eeeee4", high = "#e28743") +
  labs(x="Cell types", y="Antibodies") +
  theme_bw() + theme(axis.text.x=element_text(size=12, angle=90, vjust=0.3),
                     axis.text.y=element_text(size=14),
                     plot.title=element_text(size=12))
dev.off()


saveRDS(seu, paste(outpath,"seuAllcells.15072022.Rds"))


```






