---
title: "R Notebook"
output: html_notebook
---


Process each sorted sample

```{r}
pathway <- "/Users/rhalenathomas/Documents/Data/scRNAseq/PhenoID/scRNAseqSorted/objs/"

Neurons1 <- readRDS(paste(pathway,"seu1.rds",sep = ""))
Neurons2 <- readRDS(paste(pathway,"seu2.rds",sep = ""))
Glia1 <- readRDS(paste(pathway,"seu3.rds",sep = ""))
Glia2 <- readRDS(paste(pathway,"seu4.rds",sep = ""))

Neurons1
Neurons2
Glia1   #-Astrocytes
Glia2   #-Radial Glia


```

Check QC

```{r}
library(Seurat)

VlnPlot(Neurons1, pt.size = 0.10, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Neurons1, pt.size = 0.10, features = c("nFeature_RNA"), y.max = 1000)
VlnPlot(Neurons2, pt.size = 0.10, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Glia1, pt.size = 0.10, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Glia2, pt.size = 0.10, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)



```



Filter

```{r}

Neuron1.ft <- subset(Neurons1, subset = nFeature_RNA > 250 & nCount_RNA < 10000)


Neuron2.ft <- subset(Neurons2, subset = nFeature_RNA > 350 & nCount_RNA > 250 & nCount_RNA < 10000) 


Glia1.ft <- subset(Glia1, subset = nFeature_RNA > 500 & nCount_RNA > 500 & nCount_RNA < 10000) 


Glia2.ft <- subset(Glia2, subset = nFeature_RNA > 300 & nCount_RNA > 500 & nCount_RNA < 10000) 


Neuron1.ft
Neuron2.ft
Glia1.ft
Glia2.ft


```

Remove doublets from each Sorted population

```{r}
suppressMessages(require(DoubletFinder))

```




```{r}
# Neurons1
seu.d <- Neuron1.ft
seu.d <- NormalizeData(seu.d)
seu.d = FindVariableFeatures(seu.d, verbose = F)
seu.d = ScaleData(seu.d, vars.to.regress = c("nFeature_RNA", "percent.mt"),
    verbose = F)
seu.d = RunPCA(seu.d, verbose = F, npcs = 20)
seu.d = RunUMAP(seu.d, dims = 1:10, verbose = F)

# starts with 1833 cells 
nExp <- round(ncol(seu.d) * 0.013)  # expect close to 1.3doublets
seu.d <- doubletFinder_v3(seu.d, pN = 0.25, pK = 0.09, nExp = nExp, PCs = 1:10)


# name of the DF prediction can change, so extract the correct column name.
DF.name = colnames(seu.d@meta.data)[grepl("DF.classification", colnames(seu.d@meta.data))]
seu.d <- seu.d[, seu.d@meta.data[, DF.name]== "Singlet"]
dim(seu.d)

Neurons1 <- seu.d


```

```{r}
# Neurons2
seu.d <- Neuron2.ft  # starting 10162 cells
seu.d <- NormalizeData(seu.d)
seu.d = FindVariableFeatures(seu.d, verbose = F)
seu.d = ScaleData(seu.d, vars.to.regress = c("nFeature_RNA", "percent.mt"),
    verbose = F)
seu.d = RunPCA(seu.d, verbose = F, npcs = 20)
seu.d = RunUMAP(seu.d, dims = 1:10, verbose = F)

nExp <- round(ncol(seu.d) * 0.076)  # expect 7.6% doublets for 
seu.d <- doubletFinder_v3(seu.d, pN = 0.25, pK = 0.09, nExp = nExp, PCs = 1:10)


# name of the DF prediction can change, so extract the correct column name.
DF.name = colnames(seu.d@meta.data)[grepl("DF.classification", colnames(seu.d@meta.data))]
seu.d <- seu.d[, seu.d@meta.data[, DF.name]== "Singlet"]
dim(seu.d)

Neurons2 <- seu.d

```

Astrocytes has too many cells Demultiplex with hashtags

```{r}

DefaultAssay(Glia1.ft) <- 'HTO'
ht <- HTODemux(Glia1.ft, positive.quantile = 0.98)
table(ht$HTO_classification.global)
table(ht$HTO_classification)
table(ht$hash.ID)
RidgePlot(ht, assay = "HTO", group.by = "HTO_maxID", features = rownames(ht[["HTO"]]), log = TRUE)
# perhaps I put the wrong hashtags in the compute canada pipeline

VlnPlot(ht, assay = "RNA", group.by = 'hash.ID', features = "nFeature_RNA")

# one of the replicates isn't represented labelled or not recognize well
# I'll select the two reps with clear hashtags


Idents(ht) <- 'hash.ID'
glia1.temp <- subset(ht, idents = c("B0252-TotalSeqB","B0251-TotalSeqB"))
dim(glia1.temp)




```



In this case some doublets will already be removed

```{r}
# Glia 1 astrocytes  8678 cells

DefaultAssay(glia1.temp) <- "RNA"
seu.d <- glia1.temp
seu.d <- NormalizeData(seu.d)
seu.d = FindVariableFeatures(seu.d, verbose = F)
seu.d = ScaleData(seu.d, vars.to.regress = c("nFeature_RNA", "percent.mt"),
    verbose = F)
seu.d = RunPCA(seu.d, verbose = F, npcs = 20)
seu.d = RunUMAP(seu.d, dims = 1:10, verbose = F)

nExp <- round(ncol(seu.d) * 0.064)  # expect 6.4% doublets input of 13000 cells
seu.d <- doubletFinder_v3(seu.d, pN = 0.25, pK = 0.09, nExp = nExp, PCs = 1:10)


# name of the DF prediction can change, so extract the correct column name.
DF.name = colnames(seu.d@meta.data)[grepl("DF.classification", colnames(seu.d@meta.data))]
seu.d <- seu.d[, seu.d@meta.data[, DF.name]== "Singlet"]
dim(seu.d)

Astrocytes <- seu.d

```

Glia2 - Radial Glia, low CD44 in the FACS sort.

```{r}

seu.d <- Glia2.ft   # 5000 cells 
seu.d <- NormalizeData(seu.d)
seu.d = FindVariableFeatures(seu.d, verbose = F)
seu.d = ScaleData(seu.d, vars.to.regress = c("nFeature_RNA", "percent.mt"),
    verbose = F)
seu.d = RunPCA(seu.d, verbose = F, npcs = 20)
seu.d = RunUMAP(seu.d, dims = 1:10, verbose = F)

nExp <- round(ncol(seu.d) * 0.039)  # expect 6% doublets 5000 cells 
seu.d <- doubletFinder_v3(seu.d, pN = 0.25, pK = 0.09, nExp = nExp, PCs = 1:10)


# name of the DF prediction can change, so extract the correct column name.
DF.name = colnames(seu.d@meta.data)[grepl("DF.classification", colnames(seu.d@meta.data))]
seu.d <- seu.d[, seu.d@meta.data[, DF.name]== "Singlet"]
dim(seu.d)

RadialGlia <- seu.d

```


Annotate the cell types in each Flow population

```{r}

# Neurons 1
seu <- NormalizeData(Neurons1, normalization.method = "LogNormalize", scale.factor = 10000)
Idents(seu) <- 'orig.ident'
seu$orig.ident <- 'Neurons1'
dim(seu)

seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 2000)
seu <- ScaleData(seu)
seu <- RunPCA(seu)
seu <- RunUMAP(seu, reduction = "pca", n.neighbors = 43, dims = 1:30)
DimPlot(seu, reduction = "umap")




```







Rename Idents to prepare to merge


```{r}

Idents(Neurons1) <- 'orig.ident'
Neurons1$orig.ident <- 'Neurons1'


Idents(Neurons2) <- 'orig.ident'
Neurons2$orig.ident <- 'Neurons2'

Idents(Astrocytes) <- 'orig.ident'
Astrocytes$orig.ident <- 'Astroctyes'
Idents(RadialGlia) <- 'orig.ident'
RadialGlia$orig.ident <- 'RadialGlia'


# downsample the Neurons2, Astrocytes and Radial Glia
Neurons2.sub <- subset(Neurons2, downsample = 2000)
Neurons2.sub
Idents(Neurons2.sub) <- 'orig.ident'
Astro.sub <- subset(Astrocytes, downsample = 2000)
Astro.sub
Idents(Astro.sub) <- 'orig.ident'
RG.sub <- subset(RadialGlia, downsample = 2000)
RG.sub
Idents(RG.sub) <- 'orig.ident'
sorted.merge <- merge(Neurons1, y=c(Neurons2.sub,Astro.sub,RG.sub), add.cell.ids = c("Neurons1","Neurons2","Astrocytes","RadialGlia"), project = "SortedCells")
sorted.merge
unique(sorted.merge$orig.ident)

saveRDS(sorted.merge, "/Users/rhalenathomas/Documents/Data/scRNAseq/PhenoID/scRNAseqSorted/objs/mergedObjectNov2.Rds")



```


Cluster merge object 

```{r}

seu <- sorted.merge

seu <- NormalizeData(seu, normalization.method = "LogNormalize", scale.factor = 10000)
seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 2000)
seu <- ScaleData(seu)
seu <- RunPCA(seu)
ElbowPlot(seu, ndims = 30)
seu <- RunUMAP(seu, dim = 1:18, n.neighbors = 43)
DimPlot(seu, reduction = "umap")

#



```


Quick annotation from previously annotated


```{r}

anchors <- FindTransferAnchors(reference = seu.L ,query = seu, dims = 1:25)
predictions <- TransferData(anchorset = anchors, refdata = seu.L$Cell_Subtype_Markers)
seu <- AddMetaData(seu, metadata = predictions$predicted.id, col.name = 'predicted.id')

predictions <- TransferData(anchorset = anchors, refdata = seu.L$Cell_Types)
seu <- AddMetaData(seu, metadata = predictions$predicted.id, col.name = 'pred.Celltypes')
DimPlot(seu, group.by = 'predicted.id')
DimPlot(seu, group.by = 'predicted.id', split.by = 'orig.ident')

DimPlot(seu, group.by = 'pred.Celltypes')
DimPlot(seu, group.by = 'pred.Celltypes', split.by = 'orig.ident')


```

NOTE TO ME ******* 

I will need to look through the pre - filtering cells - neurons 1 is the same 

Possible just down sample astro so the numbers are the same. The other samples were randomly downsampled.  Maybe if I repeat the down sampling with different seeds I'll get results more similar.  Also if I cluster before down sampling like I did the first time.  





```{r}






```








Calculate proprotion in object in the figures:

```{r}

pathway <- "/Users/rhalenathomas/Documents/Data/scRNAseq/PhenoID/scRNAseqSorted/objs/"
seu.L <- readRDS(paste(pathway, "CombinedLabeledMarkers14102022.RDS"))

DimPlot(seu.L, group.by = 'orig.ident')

table(seu.L$orig.ident)

# switch back the switched Neurons
# rename
Idents(seu.L) <- 'orig.ident'
cluster.ids <- c("Neurons1", "Neurons2", "Astrocytes","RadialGlia"
                 )


names(cluster.ids) <- levels(seu.L)
seu.L <- RenameIdents(seu.L, cluster.ids)
seu.L$orig.ident <- Idents(seu.L)


saveRDS(seu.L, paste(pathway, "FigureObject02112022.RDS"))


```



```{r}

table(seu.L$orig.ident, seu.L$Cell_Types)
table(seu.L$orig.ident, seu.L$Cell_Subtype_Markers)

```

Figure 6 F

```{r}

# make the bar chart  with the correct colours 

sample.order <- c("Astrocytes-COL3A1","Astrocytes-FABP5", "Astrocytes-HPD",
                  "DANeurons-RAB3B","DANeurons-TPBG","DANeurons-TTR",
                  "Mix","NPC",
                  "Neurons-ASCL1",
                  "Neurons-CP","Neurons-GRIA2",
                  "Neurons-MGP",
                  "Neurons-SPARCL1",
                  "Neurons-TPH1", "Neurons-TFPI1",
                  "RadialGlia-CY1B1", "RadialGlia-NEAT1","RadialGlia-PTN",
                  "RadialGlia-RPL41", "RadialGlia-TOP2A", "RadialGlia-VCAN")
sample.order <- rev(sample.order)


clust.colours <- c("#F1C166","#FF9400","#E95901",          # oranges Astrocytes
                   "#43D59C","#7DA696","#15AE36",     # greens DA neurons
                   "steelblue","red2",   # mix and NPC
                   "#8F67FF",
                   "#6840DB","#9B8BC7",   # neurons 2  CP, GRIA2
                   "#7B22FB",    # MGP
                   "purple", # SPARCL1
                   "#9863E5","#8E36D2", # TPH1, TFPI1
                   "#F68D8D", # RG CY1B1
                   "#F68888","#F688BD",   # RG NEAT1, PTN 
                   "#F5288A","#DC57CF","#F3A6EC")




df <- as.data.frame(table(seu.L$orig.ident, seu.L$Cell_Subtype_Markers))

# the chart will default to alphabetical


# but not for the x axis - reorder with factor to match figure 6A
df$FACpop <- factor(df$Var1, levels = c("Astrocytes","RadialGlia","Neurons1","Neurons2"))
df$Cell_subtypes <- factor(df$Var2, levels = sample.order)

pdf(paste(output_path,"BarChartProportionCellTypesin4popsNov2.pdf"), height = 5, width = 8)
ggplot(df, aes(x = FACpop, y = Freq, fill = Cell_subtypes)) + 
  geom_col(position = "fill") + theme_classic() +
  scale_fill_manual(values = clust.colours) +
  #scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90, size = 16), axis.text.y = element_text(size = 16)) +
  labs(y = "Proportion of Cells", x = "") +
  theme(axis.title.y = element_text(size = 16), legend.text = element_text(size = 16), legend.title = element_text(size = 16)) 
dev.off()




```





